---
title: "Server-Side Data — Tests"
output:
  html_document:
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reactable)
```

## Lazy sub-row fetching (grouped expansion)

These examples test that expanding/collapsing groups triggers a re-query and
only fetches sub-rows for expanded groups (9E lazy sub-row fetching).

**What to verify:**
1. All groups start collapsed with expander arrows visible
2. Clicking a group expander shows sub-rows (triggers server re-query)
3. Collapsing a group hides sub-rows (triggers another re-query)
4. Multi-level: expanding a parent shows sub-groups (also collapsed initially)
5. Performance: with 100k rows, expanding a group should be fast (only fetches
   that group's sub-rows, not all groups)

### DuckDB server backend - single-level groupBy

```{r eval=FALSE}
library(shiny)
library(reactable)

ui <- fluidPage(
  h3("DuckDB server - lazy grouped expansion"),
  reactableOutput("tbl")
)

server <- function(input, output) {
  output$tbl <- renderReactable({
    reactable(
      MASS::Cars93[, c("Manufacturer", "Model", "Type", "Min.Price", "Max.Price", "Horsepower")],
      backend = backendDuckDB(),
      defaultPageSize = 10,
      groupBy = "Manufacturer",
      filterable = TRUE,
      searchable = TRUE,
      columns = list(
        Min.Price = colDef(aggregate = "sum"),
        Max.Price = colDef(aggregate = "mean"),
        Horsepower = colDef(aggregate = "max")
      )
    )
  })
}

shinyApp(ui, server)
```

### DuckDB server backend - multi-level groupBy

```{r eval=FALSE}
library(shiny)
library(reactable)

ui <- fluidPage(
  h3("DuckDB server - multi-level lazy expansion"),
  reactableOutput("tbl")
)

server <- function(input, output) {
  output$tbl <- renderReactable({
    reactable(
      MASS::Cars93[, c("Manufacturer", "Model", "Type", "Origin", "Min.Price", "Max.Price")],
      backend = backendDuckDB(),
      defaultPageSize = 10,
      groupBy = c("Origin", "Type"),
      filterable = TRUE,
      searchable = TRUE,
      columns = list(
        Min.Price = colDef(aggregate = "sum", format = colFormat(digits = 1)),
        Max.Price = colDef(aggregate = "mean", format = colFormat(digits = 1)),
        Manufacturer = colDef(aggregate = "unique")
      )
    )
  })
}

shinyApp(ui, server)
```

### df backend - single-level groupBy

```{r eval=FALSE}
library(shiny)
library(reactable)

ui <- fluidPage(
  h3("df backend - lazy grouped expansion"),
  reactableOutput("tbl")
)

server <- function(input, output) {
  output$tbl <- renderReactable({
    reactable(
      MASS::Cars93[, c("Manufacturer", "Model", "Type", "Min.Price", "Max.Price", "Horsepower")],
      backend = backendDf(),
      defaultPageSize = 10,
      groupBy = "Manufacturer",
      filterable = TRUE,
      searchable = TRUE,
      columns = list(
        Min.Price = colDef(aggregate = "sum"),
        Max.Price = colDef(aggregate = "mean"),
        Horsepower = colDef(aggregate = "max")
      )
    )
  })
}

shinyApp(ui, server)
```

### df backend - multi-level groupBy

```{r eval=FALSE}
library(shiny)
library(reactable)

ui <- fluidPage(
  h3("df backend - multi-level lazy expansion"),
  reactableOutput("tbl")
)

server <- function(input, output) {
  output$tbl <- renderReactable({
    reactable(
      MASS::Cars93[, c("Manufacturer", "Model", "Type", "Origin", "Min.Price", "Max.Price")],
      backend = backendDf(),
      defaultPageSize = 10,
      groupBy = c("Origin", "Type"),
      filterable = TRUE,
      searchable = TRUE,
      columns = list(
        Min.Price = colDef(aggregate = "sum", format = colFormat(digits = 1)),
        Max.Price = colDef(aggregate = "mean", format = colFormat(digits = 1)),
        Manufacturer = colDef(aggregate = "unique")
      )
    )
  })
}

shinyApp(ui, server)
```

### Large dataset - 1000 rows with DuckDB server

Tests lazy fetching with a moderate dataset. Without lazy fetching, expanding any group would
fetch all sub-rows for all visible groups. With lazy fetching, only the expanded
group's sub-rows are fetched. Kept to 1000 rows since non-paginateSubRows renders all
sub-rows in the DOM at once.

```{r eval=FALSE}
library(shiny)
library(reactable)

n <- 1000
data <- data.frame(
  region = sample(c("East", "West", "North", "South"), n, replace = TRUE),
  state = sample(state.name[1:10], n, replace = TRUE),
  category = sample(LETTERS[1:5], n, replace = TRUE),
  revenue = round(runif(n, 100, 50000), 2),
  quantity = sample(1:100, n, replace = TRUE),
  stringsAsFactors = FALSE
)

ui <- fluidPage(
  h3("DuckDB server - 100k rows lazy expansion"),
  p("Expand a group - only that group's sub-rows should be fetched"),
  reactableOutput("tbl")
)

server <- function(input, output) {
  output$tbl <- renderReactable({
    reactable(
      data,
      backend = backendDuckDB(),
      defaultPageSize = 20,
      groupBy = "region",
      filterable = TRUE,
      searchable = TRUE,
      sortable = TRUE,
      columns = list(
        revenue = colDef(aggregate = "sum", format = colFormat(currency = "USD", separators = TRUE, digits = 0)),
        quantity = colDef(aggregate = "sum"),
        state = colDef(aggregate = "unique"),
        category = colDef(aggregate = "count")
      )
    )
  })
}

shinyApp(ui, server)
```

These tests use the V8 server backend (`backendV8()`). All selection tests
mirror the DuckDB test doc to ensure consistent behavior across backends.

## Cross-page selection

### Select-all across pages

Select all, then navigate to page 2 -- all rows should still be selected.
Deselect a row on page 2, navigate back to page 1 -- all page 1 rows should
still be selected. Navigate back to page 2 -- the deselected row should still
be unchecked.

With search: type a search term (e.g., "Ford"), select all filtered rows,
clear the search -- only the previously filtered rows should remain selected.
The select-all checkbox should not be checked (only a subset is selected).

### Select-all + search filter reset

1. Click select-all checkbox (all rows become selected)
2. Type in the search box (e.g., "Ford")
3. Selection should be cleared -- the select-all should not carry over to the
   filtered result set

### Select-all + column filter reset

1. Click select-all checkbox
2. Type a value in a column filter (e.g., filter Manufacturer to "Ford")
3. Selection should be cleared

### Select-all + both search and column filter

### Per-row selection across pages

1. Select rows 1 and 3 on page 1
2. Navigate to page 2 -- no rows should be selected
3. Select row 2 on page 2
4. Navigate back to page 1 -- rows 1 and 3 should still be selected
5. Search for "Ford" -- selected Ford rows should still be checked
6. Clear search -- rows 1 and 3 (and row 2 from page 2) should still be selected

### Default selected rows

Rows 1, 3, and 5 should be pre-selected on load. Navigate to page 2 -- no rows
should be checked. Navigate back to page 1 -- rows 1, 3, and 5 should still be
selected. The select-all checkbox should not be checked (only 3 of 93 are selected).

### Default selected with single selection

Row 2 should be pre-selected on load. Clicking another row should deselect row 2.

### Default selected with sorted data

Rows 1 and 2 (by original data order) should be pre-selected. With Horsepower
sorted descending, the selected rows may not appear on the first page. Navigate
to find them -- they should be checked.

## Shiny selection tests

### Cross-page selection

Run this Shiny app to test cross-page selection with `getReactableState("selected")`.

Test steps:
1. Select all -- should select all 93 rows. `getReactableState` should show `1:93`.
2. Search for "Ford" -- selected Ford rows should still be checked.
   `getReactableState` should still show all 93 indices.
3. Clear search, filter Manufacturer to "Ford" -- only Ford rows are checked.
4. Click select-all on filtered view -- should select only the filtered rows.
   `getReactableState` should show non-sequential indices (e.g., `14, 15, 16, 17, 18`).
5. Clear filter -- only those Ford rows should remain selected, others unchecked.
6. Sort by a column -- selections should persist.

```r
library(shiny)
library(reactable)

ui <- fluidPage(
  h3("Cross-page selection test (V8 backend)"),
  actionButton("select_rows", "Select rows 1, 3, 5"),
  actionButton("clear_selection", "Clear selection"),
  actionButton("select_all", "Select all rows"),
  reactableOutput("table"),
  verbatimTextOutput("selected_output")
)

server <- function(input, output, session) {
  output$table <- renderReactable({
    reactable(
      MASS::Cars93[, 1:8],
      backend = backendV8(),
      selection = "multiple",
      defaultPageSize = 10,
      searchable = TRUE,
      filterable = TRUE
    )
  })

  observeEvent(input$select_rows, {
    updateReactable("table", selected = c(1, 3, 5))
  })
  observeEvent(input$clear_selection, {
    updateReactable("table", selected = NA)
  })
  observeEvent(input$select_all, {
    updateReactable("table", selected = seq_len(nrow(MASS::Cars93)))
  })

  output$selected_output <- renderPrint({
    selected <- getReactableState("table", "selected")
    cat("Count:", length(selected), "\n")
    cat("Selected:", paste(selected, collapse = ", "), "\n")
  })
}

shinyApp(ui, server)
```

### defaultSelected

Test that `defaultSelected` highlights the correct rows and that
`getReactableState("selected")` returns the right 1-based indices on load.

Test steps:
1. On load, rows 1, 3, 5 should be checked. Output should show `1, 3, 5`.
2. Navigate to page 2 -- no rows should be checked.
3. Navigate back to page 1 -- rows 1, 3, 5 still checked.
4. Deselect row 3, output should show `1, 5`.

```r
library(shiny)
library(reactable)

ui <- fluidPage(
  h3("defaultSelected test (V8 backend)"),
  actionButton("select_rows", "Select rows 2, 4, 6"),
  actionButton("clear_selection", "Clear selection"),
  reactableOutput("table"),
  verbatimTextOutput("selected_output")
)

server <- function(input, output, session) {
  output$table <- renderReactable({
    reactable(
      MASS::Cars93[, 1:8],
      backend = backendV8(),
      selection = "multiple",
      defaultSelected = c(1, 3, 5),
      defaultPageSize = 10,
      searchable = TRUE
    )
  })

  observeEvent(input$select_rows, {
    updateReactable("table", selected = c(2, 4, 6))
  })
  observeEvent(input$clear_selection, {
    updateReactable("table", selected = NA)
  })

  output$selected_output <- renderPrint({
    selected <- getReactableState("table", "selected")
    cat("Count:", length(selected), "\n")
    cat("Selected:", paste(selected, collapse = ", "), "\n")
  })
}

shinyApp(ui, server)
```

### defaultSelected + defaultSorted

Test that `defaultSelected` refers to original data row indices, not display
position, even when `defaultSorted` reorders the table.

Test steps:
1. On load, the table is sorted by Horsepower descending. Rows 1 and 2 from the
   original data (Acura Integra and Acura Legend) should be selected, NOT the
   first two rows on screen. Output should show `1, 2`.
2. The selected rows may not be visible on the first page. Navigate to find
   them -- they should be checked.
3. Sort by a different column -- the same rows should stay checked, output
   unchanged.

```r
library(shiny)
library(reactable)

ui <- fluidPage(
  h3("defaultSelected + defaultSorted test (V8 backend)"),
  actionButton("select_rows", "Select rows 3, 5, 7"),
  actionButton("clear_selection", "Clear selection"),
  reactableOutput("table"),
  verbatimTextOutput("selected_output")
)

server <- function(input, output, session) {
  output$table <- renderReactable({
    reactable(
      MASS::Cars93[, c("Manufacturer", "Model", "Type", "Min.Price", "Price", "Max.Price", "MPG.city", "Horsepower")],
      backend = backendV8(),
      selection = "multiple",
      defaultSelected = c(1, 2),
      defaultSorted = list(Horsepower = "desc"),
      defaultPageSize = 10,
      searchable = TRUE
    )
  })

  observeEvent(input$select_rows, {
    updateReactable("table", selected = c(3, 5, 7))
  })
  observeEvent(input$clear_selection, {
    updateReactable("table", selected = NA)
  })

  output$selected_output <- renderPrint({
    selected <- getReactableState("table", "selected")
    cat("Count:", length(selected), "\n")
    cat("Selected:", paste(selected, collapse = ", "), "\n")
  })
}

shinyApp(ui, server)
```
