---
title: "Conditional Styling"
output:
  html_document:
    toc: true
    toc_float:
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reactable)
```


You can conditionally style a table using functions that return inline styles
or CSS classes. Just like with [custom rendering](custom-rendering.html),
style functions can either be in R or JavaScript:

<table>
<thead><tr><th>R functions</th><th>JavaScript functions</th></tr></thead>
<tbody>
<tr><td>

```{r eval=FALSE}
reactable(
  iris,
  rowStyle = function(index) {
    if (iris[index, "Sepal.Width"] > 3.5) {
      list(fontWeight = "bold")
    }
  }
)
```

</td><td>

```{r eval=FALSE}
reactable(
  iris,
  rowStyle = JS("function(rowInfo) {
    if (rowInfo.row['Sepal.Width'] > 3.5) {
      return { fontWeight: 'bold' }
    }
  }")
)
```

</td></tr>
<tr><td>

- Easier to use but more static
- Style once, when the table is created

</td><td>

- Harder to use but more dynamic
- Style on the fly, based on client-side state

</td></tr>
</tbody></table>

Whichever one to use depends on the situation and personal preference.
You might prefer to use R functions except when you need more dynamic behavior
(e.g., style based on sorted state).

#### Example: color scales
We can use R's built-in
[color utilities](https://bookdown.org/rdpeng/exdata/plotting-and-color-in-r.html#color-utilities-in-r)
to apply a color scale to a column:
```{r}
data <- iris[1:5, ]
orange_pal <- function(x) rgb(colorRamp(c("#ffe4cc", "#ffb54d"))(x), maxColorValue = 255)

reactable(
  data,
  columns = list(
    Petal.Length = colDef(
      style = function(value) {
        normalized <- (value - min(data$Petal.Length)) / (max(data$Petal.Length) - min(data$Petal.Length))
        color <- orange_pal(normalized)
        list(background = color)
      }
    )
  )
)
```

#### Example: highlight sorted columns
To style sorted columns, we need to use a JavaScript function to determine
whether a column is currently being sorted:
```{r highlight_sorted, eval=FALSE}
reactable(
  iris[1:5, ],
  defaultSorted = "Petal.Length",
  defaultColDef = colDef(
    class = JS("function(rowInfo, colInfo, state) {
      // Highlight sorted columns
      for (var i = 0; i < state.sorted.length; i++) {
        if (state.sorted[i].id === colInfo.id) {
          return 'sorted'
        }
      }
    }")
  )
)
```

```{css}
.sorted {
  background: rgba(0, 0, 0, 0.03);
}
```

```{r ref.label="highlight_sorted", echo=FALSE}
```


## Cell Styling

### R functions
Both `style` and `class` take an R function with up to 3 optional arguments:
```{r, eval=FALSE}
colDef(
  style = function(value, index, name) {
    # input:
    #   - value, the cell value
    #   - index, the row index (optional)
    #   - name, the column name (optional)
    #
    # output:
    #   - a named list with camelCased property names
    list(color = "red", marginLeft = "30px")
    #   - or an inline style string
    else "color: red; margin-left: 30px;"
  },
  class = function(value, index, name) {
    # input:
    #   - value, the cell value
    #   - index, the row index (optional)
    #   - name, the column name (optional)
    #
    # output:
    #   - CSS class names
    "class1 class2"
  }
)
```

> NOTE: R functions cannot apply styles to aggregated cells.

### JavaScript functions
Or a JavaScript function, wrapped in `JS()`, with up to 3 optional arguments:
```{r, eval=FALSE}
colDef(
  style = JS("
    function(rowInfo, colInfo, state) {
      // input:
      //  - rowInfo, an object containing row info
      //  - colInfo, an object containing column info (optional)
      //  - state, an object containing the table state (optional)
      //
      // output:
      //  - a style object with camelCased property names
      return { backgroundColor: 'gray' }
    }
  "),
  class = JS("
    function(rowInfo, colInfo, state) {
      // input:
      //  - rowInfo, an object containing row info
      //  - colInfo, an object containing column info (optional)
      //  - state, an object containing the table state (optional)
      //
      // output:
      //  - CSS class names
      return 'class1 class2'
    }
  ")
)
```

#### `rowInfo` properties
```{r, echo=FALSE}
props <- tibble::tribble(
  ~Property, ~Example, ~Description,
  "row", '{Petal.Length: 1.7, Species: "setosa"}', "row data object",
  "index", "20", "row index (zero-based)",
  "viewIndex", "0", "row index within the page (zero-based)",
  "page", "2", "page index (zero-based)",
  "aggregated", "true", "whether the row is aggregated",
  "subRows", '[{Petal.Length: 1.7, Species: "setosa"}, ...]', "array of sub row data objects (aggregated rows only)"
)

reactable(props, sortable = FALSE, bordered = TRUE, columns = list(
  Property = colDef(cell = function(value) htmltools::code(value), width = 100),
  Example = colDef(cell = function(value) htmltools::code(value))
))
```

#### `colInfo` properties
```{r, echo=FALSE}
props <- tibble::tribble(
  ~Property, ~Example, ~Description,
  "id", '"Petal.Length"', "column ID",
  "name", '"Petal Length"', "column display name"
)

reactable(props, sortable = FALSE, bordered = TRUE, columns = list(
  Property = colDef(cell = function(value) htmltools::code(value), width = 100),
  Example = colDef(cell = function(value) htmltools::code(value))
))
```

#### `state` properties
```{r, echo=FALSE}
props <- tibble::tribble(
  ~Property, ~Example, ~Description,
  "sorted", '[{id: "Petal.Length", desc: true}, ...]', "columns being sorted in the table",
  "pageRows", '[{Petal.Length: 1.7, Species: "setosa"}, ...]', "array of row data objects in the page",
  "sortedData", '[{Petal.Length: 1.7, Species: "setosa"}, ...]', "sorted array of row data objects in the table",
  "data", '[{Petal.Length: 1.7, Species: "setosa"}, ...]', "original array of row data objects in the table",
  "expanded", '{2: "Species"} or {2: {0: "Species"}}', "expanded rows in the table (can be nested)"
)

reactable(props, sortable = FALSE, bordered = TRUE, columns = list(
  Property = colDef(cell = function(value) htmltools::code(value), width = 100),
  Example = colDef(cell = function(value) htmltools::code(value))
))
```


## Row Styling

### R functions
Both `rowStyle` and `rowClass` take an R function with a single argument:
```{r, eval=FALSE}
reactable(
  rowStyle = function(index) {
    # input:
    #   - index, the row index
    #
    # output:
    #   - a named list with camelCased property names
    list(color = "red", marginLeft = "30px")
    #   - or an inline style string
    "color: red; margin-left: 30px;"
  },
  rowClass = function(index) {
    # input:
    #   - index, the row index
    #
    # output:
    #   - CSS class names
    "class1 class2"
  }
)
```

> NOTE: R functions cannot apply styles to aggregated rows.

### JavaScript functions
Or a JavaScript function with up to 2 optional arguments:
```{r, eval=FALSE}
reactable(
  rowStyle = JS("
    function(rowInfo, state) {
      // input:
      //  - rowInfo, an object containing row info
      //  - state, an object containing the table state (optional)
      //
      // output:
      //  - a style object with camelCased properties
      return { backgroundColor: 'gray' }
    }
  "),
  rowClass = JS("
    function(rowInfo, state) {
      // input:
      //  - rowInfo, an object containing row info
      //  - state, an object containing the table state (optional)
      //
      // output:
      //  - CSS class names
      return 'class1 class2'
    }
  ")
)
```

#### `rowInfo` properties
```{r, echo=FALSE}
props <- tibble::tribble(
  ~Property, ~Example, ~Description,
  "row", '{Petal.Length: 1.7, Species: "setosa"}', "row data object",
  "index", "20", "row index (zero-based)",
  "viewIndex", "0", "row index within the page (zero-based)",
  "page", "2", "page index (zero-based)",
  "aggregated", "true", "whether the row is aggregated",
  "subRows", '[{Petal.Length: 1.7, Species: "setosa"}, ...]', "array of sub row data objects (aggregated rows only)"
)

reactable(props, sortable = FALSE, bordered = TRUE, columns = list(
  Property = colDef(cell = function(value) htmltools::code(value), width = 100),
  Example = colDef(cell = function(value) htmltools::code(value))
))
```

#### `state` properties
```{r, echo=FALSE}
props <- tibble::tribble(
  ~Property, ~Example, ~Description,
  "sorted", '[{id: "Petal.Length", desc: true}, ...]', "columns being sorted in the table",
  "pageRows", '[{Petal.Length: 1.7, Species: "setosa"}, ...]', "array of row data objects in the page",
  "sortedData", '[{Petal.Length: 1.7, Species: "setosa"}, ...]', "sorted array of row data objects in the table",
  "data", '[{Petal.Length: 1.7, Species: "setosa"}, ...]', "original array of row data objects in the table",
  "expanded", '{2: "Species"} or {2: {0: "Species"}}', "expanded rows in the table (can be nested)"
)

reactable(props, sortable = FALSE, bordered = TRUE, columns = list(
  Property = colDef(cell = function(value) htmltools::code(value), width = 100),
  Example = colDef(cell = function(value) htmltools::code(value))
))
```


```{css echo=FALSE}
/* rmarkdown html documents */
.main-container {
  font-family: -apple-system, BlinkMacSystemFont, "Segoe UI", Helvetica, Arial, sans-serif;
}

.main-container blockquote {
  font-size: inherit;
}

/* pkgdown articles */
.contents {
  font-family: -apple-system, BlinkMacSystemFont, "Segoe UI", Helvetica, Arial, sans-serif;
}

.contents blockquote {
  font-size: inherit;
}

.reactable {
  font-size: 14px;
}
```
