---
title: "Custom Rendering"
output:
  html_document:
    toc: true
    toc_float:
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reactable)
library(htmltools)

propsTable <- function(props) {
  tags$div(
    style = "overflow: auto;",
    tabindex = "0",
    tags$table(
      class = "props-tbl",
      tags$thead(
        tags$tr(
          lapply(colnames(props), function(name) tags$th(name))
        )
      ),
      tags$tbody(
        apply(props, 1, function(row) {
          tags$tr(
            tags$th(scope = "row", tags$code(row[["Property"]])),
            tags$td(tags$code(row[["Example"]])),
            tags$td(row[["Description"]])
          )
        })
      )
    )
  )
}
```


You may want to customize how your data is displayed beyond what the built-in
formatters provide. For example, inserting a link, combining data from multiple
columns, or showing a column total that updates on filtering.

In reactable, you can customize data rendering using either an R or JavaScript
function that returns custom content:

<table style="table-layout: fixed;">
<thead><tr><th>R render functions</th><th>JavaScript render functions</th></tr></thead>
<tbody>
<tr><td>

```{r eval=FALSE}
reactable(iris, columns = list(
  Species = colDef(
    cell = function(value) {
      htmltools::tags$b(value)
    }
  )
))
```

</td><td>

```{r eval=FALSE}
reactable(iris, columns = list(
  Species = colDef(
    cell = JS("function(cellInfo) {
      return `<b>${cellInfo.value}</b>'
    }")
  )
))
```

</td></tr>
<tr><td>

- Easier to use but more static
- Render once, when the table is created
- Supports [`htmltools`](https://shiny.rstudio.com/articles/tag-glossary.html) and [`htmlwidgets`](https://www.htmlwidgets.org/)

</td><td>

- Harder to use but more dynamic
- Render on the fly, based on client-side state
- Can be more efficient in large tables

</td></tr>
</tbody></table>

Whichever one to use depends on the situation and personal preference.
You might prefer to use R render functions except in cases where you need more
dynamic behavior (e.g., render based on filtered state) or have a very large table.

#### Example: column total with filtering

::: {.callout}
This example requires reactable v0.3.0 or above.
:::

For example, you can easily add a column total using an R render function:

```{r}
data <- MASS::Cars93[20:24, c("Manufacturer", "Model", "Type", "Price")]

reactable(
  data,
  searchable = TRUE,
  columns = list(
    Price = colDef(footer = function(values) {
      htmltools::tags$b(sprintf("$%.2f", sum(values)))
    }),
    Manufacturer = colDef(footer = htmltools::tags$b("Total"))
  )
)
```

However, the column total doesn't update with filtering. For that, you need a
JavaScript render function with access to the client-side filtered state:

```{r}
data <- MASS::Cars93[20:24, c("Manufacturer", "Model", "Type", "Price")]

reactable(
  data,
  searchable = TRUE,
  columns = list(
    Price = colDef(
      html = TRUE,
      footer = JS("function(column, state) {
        let total = 0
        state.sortedData.forEach(function(row) {
          total += row[column.id]
        })
        return '<b>$' + total.toFixed(2) + '</b>'
      }")
    ),
    Manufacturer = colDef(html = TRUE, footer = "<b>Total</b>")
  )
)
```


## Cells

### R render function {#cells-r-render-function}

To customize cell rendering, provide an R function with up to 3 optional arguments:

```{r, eval=FALSE}
colDef(
  cell = function(value, index, name) {
    # input:
    #   - value, the cell value
    #   - index, the row index (optional)
    #   - name, the column name (optional)
    #
    # output:
    #   - content to render (e.g. an HTML tag or widget)
    htmltools::div(style = "color: red", toupper(value))
  }
)
```

### JavaScript render function {#cells-js-render-function}

Or a JavaScript function, wrapped in `JS()`, with up to 2 optional arguments:

```{r, eval=FALSE}
colDef(
  cell = JS("
    function(cellInfo, state) {
      // input:
      //  - cellInfo, an object containing cell info
      //  - state, an object containing the table state (optional, new in v0.3.0)
      //
      // output:
      //  - content to render (e.g. an HTML string)
      return `<div>${cellInfo.value}</div>`
    }
  "),
  html = TRUE  # to render as HTML
)
```

With JavaScript functions, you can also customize rendering of grouped cells and
aggregated cells:

```{r, eval=FALSE}
colDef(
  grouped = JS("function(cellInfo, state) {
    return cellInfo.value
  }"),
  aggregated = JS("function(cellInfo, state) {
    return cellInfo.value
  }")
)
```

#### `cellInfo` properties

```{r, echo=FALSE, asis=TRUE}
cellInfoProps <- dplyr::tribble(
  ~Property, ~Example, ~Description,
  "value", '"setosa"', "cell value",
  "row", '{ Petal.Length: 1.7, Species: "setosa" }', "row data",
  "column", '{ id: "Petal.Length" }', "column info object",
  "index", "20", "row index (zero-based)",
  "viewIndex", "0", "row index within the page (zero-based)",
  "aggregated", "true", "whether the row is aggregated",
  "expanded", "true", "whether the row is expanded (new in v0.3.0)",
  "filterValue", '"petal"', "column filter value (new in v0.3.0)",
  "subRows", '[{ Petal.Length: 1.7, Species: "setosa" }, ...]', "sub rows data (aggregated cells only)",
  "level", "0", "row nesting depth (zero-based)",
  "selected", "true", "whether the row is selected (new in v0.3.0)"
)

propsTable(cellInfoProps)
```

#### `state` properties

```{r, echo=FALSE, asis=TRUE}
stateProps <- dplyr::tribble(
  ~Property, ~Example, ~Description,
  "sorted", '[{ id: "Petal.Length", desc: true }, ...]', "columns being sorted in the table",
  "page", "2", "page index (zero-based, new in v0.3.0)",
  "pageSize", "10", "page size (new in v0.3.0)",
  "pages", "5", "number of pages (new in v0.3.0)",
  "filters", '[{ id: "Species", value: "petal" }]', "column filter values (new in v0.3.0)",
  "searchValue", '"petal"', "table search value (new in v0.3.0)",
  "selected", '[0, 1, 4]', "selected row indices (zero-based, new in v0.3.0)",
  "pageRows", '[{ Petal.Length: 1.7, Species: "setosa" }, ...]', "current row data on the page",
  "sortedData", '[{ Petal.Length: 1.7, Species: "setosa" }, ...]', "current row data in the table (after sorting, filtering, grouping)",
  "data", '[{ Petal.Length: 1.7, Species: "setosa" }, ...]', "original row data in the table",
  "meta", '{ custom: 123 }', tagList("custom table metadata from", tags$code("reactable()"), "(new in v0.3.0.9000)"),
  "hiddenColumns", '["Petal.Length"]', "columns being hidden in the table"
)

propsTable(stateProps)
```

## Headers

### R render function {#headers-r-render-function}

To customize header rendering, provide an R function with up to 2 optional arguments:

```{r, eval=FALSE}
colDef(
  header = function(value, name) {
    # input:
    #   - value, the header value
    #   - name, the column name (optional)
    #
    # output:
    #   - content to render (e.g. an HTML tag or widget)
    htmltools::div(value)
  }
)
```

### JavaScript render function {#headers-js-render-function}

Or a JavaScript function with up to 2 optional arguments:

```{r, eval=FALSE}
colDef(
  header = JS("
    function(column, state) {
      // input:
      //  - column, an object containing column properties
      //  - state, an object containing the table state (optional, new in v0.3.0)
      //
      // output:
      //  - content to render (e.g. an HTML string)
      return `<div>${column.name}</div>`
    }
  "),
  html = TRUE  # to render as HTML
)
```

#### `column` properties

```{r, echo=FALSE, asis=TRUE}
columnProps <- dplyr::tribble(
  ~Property, ~Example, ~Description,
  "id", '"Petal.Length"', "column ID (new in v0.3.0)",
  "name", '"Petal Length"', "column display name (new in v0.3.0)",
  "filterValue", '"petal"', "column filter value (new in v0.3.0)",
  "setFilter", 'function setFilter(value: any)', tagList("function to set the column filter value",  "(set to ", tags$code("undefined"), " to clear the filter) (new in v0.3.0)"),
  "column", '{ id: "Petal.Length", name: "Petal Length", filterValue: "petal" }', "column info object (deprecated in v0.3.0)",
  "data", '[{ Petal.Length: 1.7, Petal.Width: 0.2, _subRows: [] }, ...]', "current row data in the table (deprecated in v0.3.0)"
)

propsTable(columnProps)
```

#### `state` properties

```{r, echo=FALSE, asis=TRUE}
propsTable(stateProps)
```

## Footers

### R render function {#footers-r-render-function}

To add footer content, provide an R function with up to 2 optional arguments:

```{r, eval=FALSE}
colDef(
  footer = function(values, name) {
    # input:
    #   - values, the column values
    #   - name, the column name (optional)
    #
    # output:
    #   - content to render (e.g. an HTML tag or widget)
    htmltools::div(paste("Total:", sum(values)))
  }
)
```

### JavaScript render function {#footers-js-render-function}

Or a JavaScript function with up to 2 optional arguments:

```{r, eval=FALSE}
colDef(
  footer = JS("
    function(column, state) {
      // input:
      //  - column, an object containing column properties
      //  - state, an object containing the table state (optional, new in v0.3.0)
      //
      // output:
      //  - content to render (e.g. an HTML string)
      return `<div>Rows: ${state.sortedData.length}</div>`
    }
  "),
  html = TRUE  # to render as HTML
)
```

#### `column` properties

```{r, echo=FALSE, asis=TRUE}
propsTable(columnProps)
```

#### `state` properties

```{r, echo=FALSE, asis=TRUE}
propsTable(stateProps)
```

## Expandable Row Details

### R render function {#details-r-render-function}

To add expandable row details, provide an R function with up to 2 optional arguments:

```{r, eval=FALSE}
reactable(
  details = function(index, name) {
    # input:
    #   - index, the row index
    #   - name, the column name (optional)
    #
    # output:
    #   - content to render (e.g., an HTML tag or nested table), or NULL to hide details
    htmltools::div(
      paste("Details for row:", index),
      reactable(data[index, ])
    )
  }
)
```

### JavaScript render function {#details-js-render-function}

Or a JavaScript function with up to 2 optional arguments:

```{r, eval=FALSE}
reactable(
  details = JS("
    function(rowInfo, state) {
      // input:
      //  - rowInfo, an object containing row info
      //  - state, an object containing the table state (optional, new in v0.3.0)
      //
      // output:
      //  - content to render (e.g. an HTML string)
      return `<div>Details for row: ${rowInfo.index}</div>`
    }
  ")
)
```

#### `rowInfo` properties

```{r, echo=FALSE, asis=TRUE}
rowInfoProps <- dplyr::tribble(
  ~Property, ~Example, ~Description,
  "values", '{ Petal.Length: 1.7, Species: "setosa" }', "row data values (new in v0.3.0)",
  "row", '{ Petal.Length: 1.7, Species: "setosa" }', tagList("same as ", tags$code("values"), " (deprecated in v0.3.0)"),
  "index", "20", "row index (zero-based)",
  "viewIndex", "0", "row index within the page (zero-based)",
  "expanded", "true", "whether the row is expanded (new in v0.3.0)",
  "level", "0", "row nesting depth (zero-based)",
  "selected", "true", "whether the row is selected"
)

propsTable(rowInfoProps)
```

#### `state` properties

```{r, echo=FALSE, asis=TRUE}
propsTable(stateProps)
```


```{css echo=FALSE}
/* rmarkdown html documents */
.main-container {
  font-family: -apple-system, BlinkMacSystemFont, "Segoe UI", Helvetica, Arial, sans-serif;
}

.main-container blockquote {
  font-size: inherit;
}
```
