---
title: "Custom Rendering"
output:
  html_document:
    toc: true
    toc_float:
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reactable)
```


Occasionally you'll want to customize how your data is displayed beyond what
the built-in formatters provide. For example, inserting a link, combining data
from multiple columns, or showing a column total that updates on filtering.

In reactable, you can customize data rendering by providing either an R or
JavaScript render function:

<table>
<thead><tr><th>R render functions</th><th>JavaScript render functions</th></tr></thead>
<tbody>
<tr><td>

```{r eval=FALSE}
reactable(iris, columns = list(
  Species = colDef(
    cell = function(value) {
      htmltools::tags$b(value)
    })
  )
)
```

</td><td>

```{r eval=FALSE}
reactable(iris, columns = list(
  Species = colDef(
    cell = JS("function(cellInfo) {
      return '<b>' + cellInfo.value + '</b>'
    }")
  )
))
```

</td></tr>
<tr><td>

- Easier to use but more static
- Render once, when the table is created
- Supports htmltools and htmlwidgets

</td><td>

- Harder to use but more dynamic
- Render on the fly, based on client-side state
- Can be more efficient in large tables

</td></tr>
</tbody></table>

Whichever one to use depends on the situation and personal preference.
You might prefer to use R render functions except in cases where you need more
dynamic behavior (e.g., render based on filtered state) or have a very large table.


## Cells

### R render function
To customize cell rendering, provide an R function with up to 3 optional arguments:
```{r, eval=FALSE}
colDef(
  cell = function(value, index, name) {
    # input:
    #   - value, the cell value
    #   - index, the row index (optional)
    #   - name, the column name (optional)
    #
    # output:
    #   - content to render (e.g. an HTML tag or widget)
    htmltools::div(style = "color: red", toupper(value))
  }
)
```

### JavaScript render function
Or a JavaScript function, wrapped in `JS()`, with a single argument:
```{r, eval=FALSE}
colDef(
  cell = JS("
    function(cellInfo) {
      // input:
      //  - cellInfo, an object containing cell and row info
      //
      // output:
      //  - content to render (e.g. an HTML string)
      return '<div>' + cellInfo.value + '</div>'
    }
  "),
  html = TRUE  # to render as HTML
)
```

With JavaScript functions, you can also customize rendering of aggregated cells:
```{r, eval=FALSE}
colDef(
  aggregated = JS("function(cellInfo) { return cellInfo.value }")
)
```

#### `cellInfo` properties
```{r, echo=FALSE}
props <- tibble::tribble(
  ~Property, ~Example, ~Description,
  "value", '"setosa"', "cell value",
  "row", '{Petal.Length: 1.7, Species: "setosa"}', "row data object",
  "column", '{id: "Species"}', "column info object",
  "index", "20", "row index (zero-based)",
  "viewIndex", "0", "row index within the page (zero-based)",
  "page", "2", "page index (zero-based)",
  "show", "true", "whether the column is visible"
)

reactable(props, sortable = FALSE, bordered = TRUE, columns = list(
  Property = colDef(cell = function(value) htmltools::code(value), width = 100),
  Example = colDef(cell = function(value) htmltools::code(value))
))
```


## Headers

### R render function
To customize header rendering, provide an R function with up to 2 optional arguments:
```{r, eval=FALSE}
colDef(
  header = function(value, name) {
    # input:
    #   - value, the header value
    #   - name, the column name (optional)
    #
    # output:
    #   - content to render (e.g. an HTML tag or widget)
    htmltools::div(value)
  }
)
```

### JavaScript render function
Or a JavaScript function with a single argument:
```{r, eval=FALSE}
colDef(
  header = JS("
    function(colInfo) {
      // input:
      //  - colInfo, an object containing column info
      //
      // output:
      //  - content to render (e.g. an HTML string)
      return '<div>' + colInfo.column.name + '</div>'
    }
  "),
  html = TRUE  # to render as HTML
)
```

#### `colInfo` properties
```{r, echo=FALSE}
props <- tibble::tribble(
  ~Property, ~Example, ~Description,
  "column", '{id: "Petal.Width", name: "Petal Width", show: true}', "column info object",
  "data", '[{Sepal.Length: 5.1, Petal.Width: 0.2}, ...]', "array of row data objects"
)

reactable(props, sortable = FALSE, bordered = TRUE, columns = list(
  Property = colDef(cell = function(value) htmltools::code(value), width = 100),
  Example = colDef(cell = function(value) htmltools::code(value), width = 450)
))
```


## Footers

### R render function
To add footer content, provide an R function with up to 2 optional arguments:
```{r, eval=FALSE}
colDef(
  footer = function(values, name) {
    # input:
    #   - values, the column values
    #   - name, the column name (optional)
    #
    # output:
    #   - content to render (e.g. an HTML tag or widget)
    htmltools::div(paste("Total:", sum(values)))
  }
)
```

### JavaScript render function
Or a JavaScript function with a single argument:
```{r, eval=FALSE}
colDef(
  footer = JS("
    function(colInfo) {
      // input:
      //  - colInfo, an object containing column info
      //
      // output:
      //  - content to render (e.g. an HTML string)
      return '<div>Rows: ' + colInfo.data.length + '</div>'
    }
  "),
  html = TRUE  # to render as HTML
)
```

#### `colInfo` properties
```{r, echo=FALSE}
props <- tibble::tribble(
  ~Property, ~Example, ~Description,
  "column", '{id: "Petal.Width", name: "Petal Width", show: true}', "column info object",
  "data", '[{Sepal.Length: 5.1, Petal.Width: 0.2}, ...]', "array of row data objects"
)

reactable(props, sortable = FALSE, bordered = TRUE, columns = list(
  Property = colDef(cell = function(value) htmltools::code(value), width = 100),
  Example = colDef(cell = function(value) htmltools::code(value), width = 450)
))
```


## Expandable Row Details

### R render function
To add expandable row details, provide an R function with a single argument:
```{r, eval=FALSE}
reactable(
  details = function(index) {
    # input:
    #   - index, the row index
    #
    # output:
    #   - content to render (e.g. an HTML tag or subtable), or NULL to hide details for the row
    htmltools::div(
      paste("Details for row:", index),
      reactable(data[index, ])
    )
  }
)
```

### JavaScript render function
Or a JavaScript function with a single argument:
```{r, eval=FALSE}
reactable(
  details = JS("
    function(rowInfo) {
      // input:
      //  - rowInfo, an object containing row info
      //
      // output:
      //  - content to render (e.g. an HTML string)
      return '<div>' + JSON.stringify(rowInfo) + '</div>'
    }
  ")
)
```

#### `rowInfo` properties
```{r, echo=FALSE}
props <- tibble::tribble(
  ~Property, ~Example, ~Description,
  "row", '{Petal.Length: 1.7, Species: "setosa"}', "row data object",
  "index", "20", "row index (zero-based)",
  "viewIndex", "0", "row index within the page (zero-based)",
  "page", "2", "page index (zero-based)"
)

reactable(props, sortable = FALSE, bordered = TRUE, columns = list(
  Property = colDef(cell = function(value) htmltools::code(value), width = 100),
  Example = colDef(cell = function(value) htmltools::code(value))
))
```


```{css echo=FALSE}
/* rmarkdown html documents */
.main-container {
  font-family: -apple-system, BlinkMacSystemFont, "Segoe UI", Helvetica, Arial, sans-serif;
}

.main-container blockquote {
  font-size: inherit;
}

/* pkgdown articles */
.contents {
  font-family: -apple-system, BlinkMacSystemFont, "Segoe UI", Helvetica, Arial, sans-serif;
}

.contents blockquote {
  font-size: inherit;
}
```
