---
title: 'Client-Side DuckDB'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

set.seed(5)

options(reactable.static = TRUE)
```

::: {.callout}
New in v0.4.5.9000
:::

The DuckDB backend enables sorting, filtering, searching, and pagination for large datasets entirely in the browser, with no server required. This works in static R Markdown and Quarto documents, where traditional server-side processing isn't available.

Instead of embedding all data as JSON and processing it in JavaScript, the DuckDB backend serializes data in a compact binary format at render time, then uses [DuckDB-WASM](https://duckdb.org/docs/api/wasm/overview) to execute SQL queries directly in the browser. For smaller datasets, data is embedded as [Arrow IPC](https://arrow.apache.org/docs/format/Columnar.html#ipc-streaming-format) directly in the HTML, keeping everything in a single file. For larger datasets, data is written as a separate [Parquet](https://parquet.apache.org/) file alongside the HTML, and DuckDB reads only the portions needed for the current page. Either way, only one page of results is rendered at a time, so even million-row tables stay responsive.

## 1,000,000 rows

Try sorting, filtering, and searching to see DuckDB handling one million rows entirely client-side.

```{r}
library(reactable)

rows <- 1000000
dates <- seq.Date(as.Date("2018-01-01"), as.Date("2018-12-01"), "day")
data <- data.frame(
  index = seq_len(rows),
  date = sample(dates, rows, replace = TRUE),
  city = sample(names(precip), rows, replace = TRUE),
  state = sample(rownames(USArrests), rows, replace = TRUE),
  temp = round(runif(rows, 0, 100), 1),
  stringsAsFactors = FALSE
)

reactable(
  data,
  backend = backendDuckDB(),
  filterable = TRUE,
  searchable = TRUE,
  highlight = TRUE,
  columns = list(
    state = colDef(
      html = TRUE,
      cell = JS('function(cell) {
        return `<a href="https://wikipedia.org/wiki/${cell.value}">${cell.value}</a>`
      }')
    )
  )
)
```

## Comparison with default client-side rendering

The default client-side mode loads all data as JSON into JavaScript memory and processes everything in JavaScript. This works well for small to medium tables but slows down at scale.

### Embedded Arrow IPC

Here's how embedded Arrow IPC mode compares with the default on the 1M row table above, measured in Chrome on Windows:

|                       | Default (client-side) | DuckDB (embedded Arrow IPC)                                                                                                           |
| :-------------------- | :-------------------- | :------------------------------------------------------------------------------------------------------------------------------------ |
| **Page load**         | ~2.4s                 | ~1s (first page pre-rendered)                                                                                                         |
| **Sorting**           | ~1s, blocks UI        | Near instant                                                                                                                          |
| **Search**            | <1s, blocks UI        | ~1s, non-blocking (Web Worker)                                                                                                        |
| **Pagination**        | Slightly slower       | Very fast                                                                                                                             |
| **Memory (after GC)** | 569 MB                | 428 MB                                                                                                                                |
| **Document size**     | 47 MB (8 MB gzipped)  | 55 MB (16 MB gzipped)                                                                                                                 |
| **Dependencies**      | None (built-in)       | [arrow](https://arrow.apache.org/docs/r/) R package, [DuckDB-WASM](https://duckdb.org/docs/api/wasm/overview) (~8 MB gzipped, cached) |

Sorting is the biggest difference. DuckDB sorts near-instantly, while the default backend freezes the page for about a second at 1M rows (about 4 seconds at 2M rows). Searching is a tradeoff: the default backend is slightly faster in raw time, but DuckDB runs queries in a background thread so the UI stays responsive. Page load is faster with DuckDB despite a larger document, because the default mode must parse and transform all JSON before rendering.

The DuckDB document is about 2x larger when gzip-compressed because base64-encoded binary data compresses less efficiently than JSON text.

### External Parquet

<!-- TODO: measure and fill in Parquet comparison table (page load, sorting, search, pagination, memory, document size) -->

With external Parquet files, the HTML document itself is much smaller since the data is stored separately. DuckDB fetches only the row groups needed for each query via HTTP range requests, which should reduce both page load time and memory usage compared to embedded Arrow IPC. Benchmarks have not been measured yet.

For small tables (under ~10,000 rows), the default client-side mode is simpler and has no extra dependencies. The DuckDB backend is most useful for large tables where interactive sorting and filtering speed matters.

## External Parquet files

For very large datasets, embedding all the data in the HTML document can make the page slow to download. External Parquet files solve this by writing the data as a separate `.parquet` file alongside the HTML document. DuckDB-WASM reads the Parquet file using [HTTP range requests](https://developer.mozilla.org/en-US/docs/Web/HTTP/Guides/Range_requests), downloading only the row groups needed for the current page.

### Automatic format selection

By default, `backendDuckDB()` automatically chooses the data format:

- **Small datasets** (Arrow IPC under ~20 MB): Embedded as base64-encoded Arrow IPC directly in the HTML, same as before.
- **Large datasets** (Arrow IPC over ~20 MB): Written as an external Parquet file with HTTP range request access.

You can override this with the `format` parameter:

```r
# Always use external Parquet files
reactable(data, backend = backendDuckDB(format = "parquet"))

# Always embed Arrow IPC (e.g., for self-contained output)
reactable(data, backend = backendDuckDB(format = "arrow"))
```

### Requirements and limitations

- **Not compatible with self-contained output**: External Parquet files cannot be embedded in a single self-contained HTML file. If the output format uses `self_contained: true`, the Parquet file will not be included. Use `self_contained: false` or `embed_resources: false` when using external Parquet files.
- **Web server required**: The Parquet file must be served over HTTP for range requests to work. This is the same requirement as the DuckDB-WASM files themselves. Most web servers (including GitHub Pages) support range requests.
- **File deployment**: The Parquet file is deployed as an htmlwidgets dependency alongside the HTML output (in the `*_files/lib/` directory). When deploying, include this directory along with the HTML file.

## Limitations

The DuckDB backend is currently **experimental**. The following features are not supported:

- **Custom search/filter methods**: `searchMethod` and column `filterMethod` are not supported and will produce a warning if used. DuckDB uses SQL-based filtering instead (case-insensitive substring match for text, prefix match for numbers).
- **Custom JavaScript aggregate functions**: Custom JS `aggregate` functions are not supported and will produce a warning. Use built-in aggregate names like `"sum"`, `"mean"`, `"max"`, `"min"`, `"median"`, `"count"`, `"unique"`, and `"frequency"`.
- **R function renderers**: R functions passed to `cell`, `details`, `style`, `class`, `rowClass`, and `rowStyle` will produce a warning because they're evaluated at render time against the original row order and don't work correctly when DuckDB re-sorts or filters data. Use `JS()` functions instead, which execute in the browser with the current page's data.
- **Row selection**: Not yet supported with the DuckDB backend.

## Browser requirements

The DuckDB backend requires a modern browser with:

- **WebAssembly** - supported in all modern browsers (Chrome 57+, Firefox 52+, Safari 11+, Edge 16+).
- **Web Workers** - used to run DuckDB in a background thread so the UI stays responsive. Supported in all modern browsers.

**DuckDB-WASM download size**: The DuckDB-WASM binary is approximately 33 MB uncompressed (~8 MB gzipped). Most web servers (including GitHub Pages) serve it gzip-compressed. The WASM file is downloaded once on first page load and cached by the browser for subsequent visits. This is in addition to the data embedded in the page.

**Memory limits**: WebAssembly has a maximum memory limit of 4 GB. With embedded Arrow IPC, the full dataset is loaded into DuckDB's in-memory database, so practical limits are roughly 1-2 GB of data depending on column types (string-heavy datasets use more memory). With external Parquet files, DuckDB reads only the needed row groups via range requests, so memory usage is much lower and larger datasets can be supported.

**Local file:// URLs**: DuckDB-WASM does not work when opening HTML files directly from the filesystem (`file://` URLs) because browsers block `fetch()` and Web Worker requests from `file://` origins for security reasons. The pre-rendered first page will still display, but sorting, filtering, and pagination won't work. To test locally, serve the file over HTTP, for example with `servr::httd()` in R.

## How it works

When you use `backend = backendDuckDB()`, reactable changes how data is processed:

1. **At render time**, R serializes the data frame to a compact binary format (via the [arrow](https://arrow.apache.org/docs/r/) package). For smaller datasets (Arrow IPC under ~20 MB), the data is base64-encoded and embedded directly in the HTML. For larger datasets, the data is written as an external Parquet file alongside the HTML. The first page of data is also pre-rendered as JSON so the table displays instantly.

2. **On page load**, the browser loads the data into a [DuckDB-WASM](https://duckdb.org/docs/api/wasm/overview) database. For embedded Arrow IPC, the base64 data is decoded and imported. For external Parquet files, DuckDB creates a SQL view over the remote file and reads only the needed portions via HTTP range requests. The pre-rendered first page displays immediately, without waiting for DuckDB to initialize.

3. **On each interaction** (sort, filter, search, page change), reactable constructs a SQL query and DuckDB executes it in the browser. Only the matching page of results is returned and rendered.

This means only one page of results is ever in JavaScript memory at a time, and sorting/filtering runs at database speed rather than JavaScript array speed. With external Parquet files, the full dataset isn't even loaded into the browser -- DuckDB fetches only the portions it needs for each query.

### Parquet internals

When using the Parquet format, reactable writes the data to a `.parquet` file alongside the HTML output (as an htmlwidgets dependency). In the browser, DuckDB-WASM creates a SQL view over the remote Parquet file:

```sql
CREATE VIEW reactable_data AS SELECT * FROM read_parquet('path/to/data.parquet')
```

Because this uses a view rather than loading the entire file, DuckDB only fetches the row groups it needs for each query. This means much less data is downloaded on page load, and sorting/filtering only fetches the relevant parts of the file.

### Data size

Arrow IPC is a compact binary format that's typically smaller than the equivalent JSON:

```{r}
# Measure Arrow IPC size (the binary data embedded in the page)
arrow_ipc_size <- function(df) {
  tf <- tempfile(fileext = ".arrows")
  on.exit(unlink(tf), add = TRUE)
  arrow::write_ipc_stream(df, tf)
  file.size(tf)
}

format_mb <- function(bytes) sprintf("%.1f MB", bytes / 1024^2)

arrow_size <- arrow_ipc_size(data)

# Base64 encoding adds ~33% overhead
arrow_size_b64 <- ceiling(arrow_size * 4 / 3)

# Measure equivalent JSON size (what the default client-side mode would embed)
json_size <- function(df) {
  nchar(jsonlite::toJSON(df, dataframe = "columns"), type = "bytes")
}

json <- json_size(data)
```

|                            |                       1M rows |
| :------------------------- | ----------------------------: |
| Arrow IPC (raw binary)     |     `r format_mb(arrow_size)` |
| Arrow IPC (base64 in HTML) | `r format_mb(arrow_size_b64)` |
| JSON (default mode)        |           `r format_mb(json)` |

The Arrow IPC data is base64-encoded when embedded in the HTML, which adds ~33% to the raw binary size. Even with base64 overhead, the embedded data is comparable in size to the equivalent JSON. The key difference is what happens in the browser: with JSON, all rows are parsed and loaded into JavaScript memory. With Arrow IPC, the data is loaded into DuckDB and only one page of rows is ever in JavaScript memory at a time.

With gzip compression (standard on most web servers including GitHub Pages), the actual download sizes are significantly smaller, roughly 3x to 4x smaller than the uncompressed sizes above.

```{css echo=FALSE}
/* rmarkdown html documents */
.main-container {
  font-family: -apple-system, BlinkMacSystemFont, "Segoe UI", Helvetica, Arial, sans-serif;
}
```
