---
title: "Examples"
output:
  html_document:
    toc: true
    toc_float:
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reactable)
```

## Basic Usage
To create a data table, use `reactable()` on a data frame or matrix.
The table will be sortable and paginated by default:
```{r}
library(reactable)

reactable(iris)
```

### Column definitions
Columns can be customized by providing a named list of column definitions
created by `colDef()` to `columns`:
```{r}
reactable(
  iris[1:5, ],
  columns = list(
    Sepal.Length = colDef(name = "Sepal Length"),
    Sepal.Width = colDef(name = "Sepal Width"),
    Species = colDef(align = "center")
  )
)
```

For convenience, you can also specify a default `colDef()` to use for all
columns in `defaultColDef`:
```{r}
reactable(
  iris[1:5, ],
  defaultColDef = colDef(
    header = function(value) gsub(".", " ", value, fixed = TRUE),
    cell = function(value) format(value, nsmall = 1),
    align = "center",
    minWidth = 70,
    headerStyle = list(background = "#f7f7f8")
  ),
  columns = list(
    Species = colDef(minWidth = 140)  # overrides the default
  ),
  bordered = TRUE,
  highlight = TRUE
)
```

## Sorting

Tables are sortable by default. You can sort a column by clicking on its
header, or sort multiple columns by holding the shift key while sorting.

Sorting toggles between ascending and descending order by default. To clear
the sort, hold the shift key while sorting, and the sorting will additionally
toggle between ascending, descending, and unsorted order.

> Ascending order means the lowest, first, or earliest values will appear first.
> Descending order means the largest, last, or latest values will appear first.

### Default sorted columns

You can set the default sorted columns by providing a vector of column names
to `defaultSorted`:

```{r}
reactable(iris[48:52, ], defaultSorted = c("Species", "Petal.Length"))
```

You can also provide a named list to customize the default sort orders.
Use `"asc"` for ascending order, or `"desc"` for descending order:

```{r}
reactable(iris[48:52, ], defaultSorted = list(Species = "asc", Petal.Length = "desc"))
```

### Default sort order

Columns are sorted in ascending order first by default. To change the default sort
order for all columns in the table, set `defaultSortOrder` in `reactable()` to
`"asc"` for ascending order, or `"desc"` for descending order.

To change the sort order of an individual column, set `defaultSortOrder` in
its `colDef()` to `"asc"` or `"desc"`. The default sort order of the column takes
precedence over the table.

```{r}
reactable(
  iris[48:52, ],
  defaultSortOrder = "desc",
  columns = list(
    Species = colDef(defaultSortOrder = "asc")
  ),
  defaultSorted = c("Species", "Petal.Length")
)
```

### Sort missing values last

You can ignore missing values when sorting by setting `sortNALast` on a column:

```{r}
reactable(
  data.frame(
    n = c(1, 2, 3, -Inf, Inf),
    x = c(2, 3, 1, NA, NaN),
    y = c("aa", "cc", "bb", NA, NA)
  ),
  defaultColDef = colDef(sortNALast = TRUE),
  defaultSorted = "x"
)
```

### No sorting

You can disable sorting by setting `sortable` to `FALSE` on the table or column.
When only some columns are sortable, it can help to indicate sortable
columns using `showSortable`:

```{r}
reactable(
  iris[1:5, ],
  sortable = FALSE,
  showSortable = TRUE,
  columns = list(
    Petal.Width = colDef(sortable = TRUE),
    Petal.Length = colDef(sortable = TRUE)
  )
)
```

### Hide sort icons

You can hide sort icons by setting `showSortIcon` to `FALSE`. This is only
recommended when you want to use a [custom sort indicator](cookbook/cookbook.html#custom-sort-indicators).
```{r}
reactable(iris[1:5, ], showSortIcon = FALSE)
```


## Filtering
You can make columns filterable using `filterable`:
```{r}
data <- MASS::Cars93[1:20, c("Manufacturer", "Model", "Type", "AirBags", "Price")]

reactable(data, filterable = TRUE, minRows = 10)
```

To make specific columns filterable (or not), set `filterable` in the column definition:
```{r}
reactable(
  data,
  filterable = TRUE,
  columns = list(
    Price = colDef(filterable = FALSE)
  )
)
```


## Searching
You can make the entire table searchable using `searchable`:
```{r}
data <- MASS::Cars93[1:20, c("Manufacturer", "Model", "Type", "AirBags", "Price")]

reactable(data, searchable = TRUE, minRows = 10)
```


## Pagination
You can change the default page size by configuring `defaultPageSize`:
```{r}
reactable(iris[1:6, ], defaultPageSize = 4)
```

You can also set the minimum rows per page using `minRows`. This may be
useful when rows don't completely fill the page, or if the table has filtering:
```{r}
reactable(iris[1:6, ], defaultPageSize = 4, minRows = 4, searchable = TRUE)
```

### Page size options
You can show a dropdown of page sizes for users to choose from using `showPageSizeOptions`.
The page size options can be customized through `pageSizeOptions`:
```{r}
reactable(iris[1:12, ], showPageSizeOptions = TRUE, pageSizeOptions = c(4, 8, 12), defaultPageSize = 4)
```

### Alternative pagination types
You can use an alternative pagination type by setting `paginationType` to:

- `"jump"` to show a page jump
- `"simple"` to show previous/next buttons only

#### Page jump
```{r}
reactable(iris[1:50, ], paginationType = "jump", defaultPageSize = 4)
```

#### Simple
```{r}
reactable(iris[1:50, ], paginationType = "simple", defaultPageSize = 4)
```

### Hide page info
You can hide page info by setting `showPageInfo` to `FALSE`:
```{r}
reactable(iris[1:12, ], showPageInfo = FALSE, defaultPageSize = 4)

reactable(iris[1:12, ], showPageInfo = FALSE, showPageSizeOptions = TRUE, defaultPageSize = 4)
```

### Always show pagination
By default, pagination is hidden if the table only has one page. To keep the
pagination shown, set `showPagination` to `TRUE`. This is especially useful if
you want to keep the page info showing the number of rows in the table.
```{r}
reactable(iris[1:5, ], showPagination = TRUE)
```

### No pagination

Tables are paginated by default, but you can disable pagination by setting `pagination` to `FALSE`:

```{r}
reactable(iris[1:20, ], pagination = FALSE, highlight = TRUE, height = 250)
```

> **NOTE:** Disabling pagination is not recommended for large tables with many
> interactive elements (such as links, expand buttons, or selection checkboxes),
> as that can make it difficult for keyboard users to navigate the page.

## Grouping and Aggregation
You can group rows in a table by specifying one or more columns in `groupBy`:
```{r}
data <- MASS::Cars93[10:22, c("Manufacturer", "Model", "Type", "Price", "MPG.city")]

reactable(data, groupBy = "Manufacturer")
```

When rows are grouped, you can aggregate data in a column using an `aggregate` function:
```{r}
data <- MASS::Cars93[14:38, c("Type", "Price", "MPG.city", "DriveTrain", "Man.trans.avail")]

reactable(
  data,
  groupBy = "Type",
  columns = list(
    Price = colDef(aggregate = "max"),
    MPG.city = colDef(aggregate = "mean", format = colFormat(digits = 1)),
    DriveTrain = colDef(aggregate = "unique"),
    Man.trans.avail = colDef(aggregate = "frequency")
  )
)
```

You can use one of the built-in aggregate functions:
```r
colDef(aggregate = "sum")        # Sum of numbers
colDef(aggregate = "mean")       # Mean of numbers
colDef(aggregate = "max")        # Maximum of numbers
colDef(aggregate = "min")        # Minimum of numbers
colDef(aggregate = "median")     # Median of numbers
colDef(aggregate = "count")      # Count of values
colDef(aggregate = "unique")     # Comma-separated list of unique values
colDef(aggregate = "frequency")  # Comma-separated counts of unique values
```

Or a custom aggregate function in JavaScript:
```r
colDef(
  aggregate = JS("
    function(values, rows) {
      // input:
      //  - values: an array of all values in the group
      //  - rows: an array of row data values for all rows in the group (optional)
      //
      // output:
      //  - an aggregated value, e.g. a comma-separated list
      return values.join(', ')
    }
  ")
)
```

### Multiple groups
```{r}
data <- data.frame(
  State = state.name,
  Region = state.region,
  Division = state.division,
  Area = state.area
)

reactable(
  data,
  groupBy = c("Region", "Division"),
  columns = list(
    Division = colDef(aggregate = "unique"),
    Area = colDef(aggregate = "sum", format = colFormat(separators = TRUE))
  ),
  bordered = TRUE
)
```

### Custom aggregate function
Custom aggregate functions are useful when none of the built-in aggregate functions
apply, or when you want to aggregate values from multiple columns. For example,
when calculating aggregate averages or percentages.

Within a custom aggregate function, you can access the values in the column using
the `values` argument, and the values in other columns using the `rows` argument:
```js
columns = list(
  Price = colDef(
    aggregate = JS("function(values, rows) {
      values
      // [46.8, 27.6, 57]

      rows
      // [
      //   { "Model": "Dynasty", "Manufacturer": "Dodge", "Price": 46.8, "Units": 2 },
      //   { "Model": "Colt", "Manufacturer": "Dodge", "Price": 27.6, "Units": 5 },
      //   { "Model": "Caravan", "Manufacturer": "Dodge", "Price": 57, "Units": 5 }
      // ]
    }")
  )
)
```

Here's an example that calculates an aggregate average price by dividing the
the sum of two columns, `Price` and `Units`:

```{r, message=FALSE}
library(dplyr)

set.seed(10)

data <- sample_n(MASS::Cars93[23:40, ], 30, replace = TRUE) %>%
  mutate(Price = Price * 3, Units = sample(1:5, 30, replace = TRUE)) %>%
  mutate(Avg.Price = Price / Units) %>%
  select(Model, Manufacturer, Price, Units, Avg.Price)

reactable(
  data,
  groupBy = "Manufacturer",
  columns = list(
    Price = colDef(aggregate = "sum", format = colFormat(currency = "USD")),
    Units = colDef(aggregate = "sum"),
    Avg.Price = colDef(
      # Calculate the aggregate Avg.Price as `sum(Price) / sum(Units)`
      aggregate = JS("function(values, rows) {
        let totalPrice = 0
        let totalUnits = 0
        rows.forEach(function(row) {
          totalPrice += row['Price']
          totalUnits += row['Units']
        })
        return totalPrice / totalUnits
      }"),
      format = colFormat(currency = "USD")
    )
  )
)
```

### Include sub rows in pagination

> New in v0.2.3.9000 (unreleased)

By default, sub rows are excluded from pagination and always shown on the same
page when expanded. To include sub rows in pagination, you can set `paginateSubRows`
to `TRUE`. This is recommended for larger tables with row groups that may not all
fit on one page.

```{r}
data <- MASS::Cars93[, c("Manufacturer", "Model", "Type", "Price", "MPG.city")]

reactable(data, groupBy = "Type", paginateSubRows = TRUE)
```


## Column Formatting
You can format data in a column by providing `colFormat()` options to `format`:
```{r}
data <- data.frame(
  price_USD = c(123456.56, 132, 5650.12),
  price_INR = c(350, 23208.552, 1773156.4),
  number_FR = c(123456.56, 132, 5650.12),
  temp = c(22, NA, 31),
  percent = c(0.9525556, 0.5, 0.112),
  date = as.Date(c("2019-01-02", "2019-03-15", "2019-09-22"))
)

reactable(data, columns = list(
  price_USD = colDef(format = colFormat(prefix = "$", separators = TRUE, digits = 2)),
  price_INR = colDef(format = colFormat(currency = "INR", separators = TRUE, locales = "hi-IN")),
  number_FR = colDef(format = colFormat(locales = "fr-FR")),
  temp = colDef(format = colFormat(suffix = " °C")),
  percent = colDef(format = colFormat(percent = TRUE, digits = 1)),
  date = colDef(format = colFormat(date = TRUE, locales = "en-GB"))
))
```

By default, numbers, dates, times, and currencies are formatted in the locale of
the browser. This means users may see different results, depending on the language
preferences for their browser.

To use a specific locale for data formatting, provide a vector of BCP 47 language
tags in the `locales` argument. See a list of
[common BCP 47 language tags](https://docs.microsoft.com/en-us/openspecs/office_standards/ms-oe376/6c085406-a698-4e12-9d4d-c3b0ee3dbc4a)
for reference.

### Date formatting
```{r}
datetimes <- as.POSIXct(c("2019-01-02 3:22:15", "2019-03-15 09:15:55", "2019-09-22 14:20:00"),
                        tz = "America/New_York")
data <- data.frame(
  datetime = datetimes,
  date = datetimes,
  time = datetimes,
  time_24h = datetimes,
  datetime_pt_BR = datetimes
)

reactable(data, columns = list(
  datetime = colDef(format = colFormat(datetime = TRUE)),
  date = colDef(format = colFormat(date = TRUE)),
  time = colDef(format = colFormat(time = TRUE)),
  time_24h = colDef(format = colFormat(time = TRUE, hour12 = FALSE)),
  datetime_pt_BR = colDef(format = colFormat(datetime = TRUE, locales = "pt-BR"))
))
```

### Currency formatting
```{r}
data <- data.frame(
  USD = c(12.12, 2141.213, 0.42, 1.55, 34414),
  EUR = c(10.68, 1884.27, 0.37, 1.36, 30284.32),
  INR = c(861.07, 152122.48, 29.84, 110, 2444942.63),
  JPY = c(1280, 226144, 44.36, 164, 3634634.61),
  MAD = c(115.78, 20453.94, 4.01, 15, 328739.73)
)

reactable(data, columns = list(
  USD = colDef(
    format = colFormat(currency = "USD", separators = TRUE, locales = "en-US")
  ),
  EUR = colDef(
    format = colFormat(currency = "EUR", separators = TRUE, locales = "de-DE")
  ),
  INR = colDef(
    format = colFormat(currency = "INR", separators = TRUE, locales = "hi-IN")
  ),
  JPY = colDef(
    format = colFormat(currency = "JPY", separators = TRUE, locales = "ja-JP")
  ),
  MAD = colDef(
    format = colFormat(currency = "MAD", separators = TRUE, locales = "ar-MA")
  )
))
```

### Formatting aggregated cells
Column formatters apply to both standard and aggregated cells by default. If you
want to format aggregated cells separately, provide a named list of `cell` and
`aggregated` options:
```{r, eval=FALSE}
colDef(
  format = list(
    cell = colFormat(...),       # Standard cells
    aggregated = colFormat(...)  # Aggregated cells
  )
)
```

For example, only the aggregated `States` are formatted here:
```{r}
data <- data.frame(
  States = state.name,
  Region = state.region,
  Area = state.area
)

reactable(
  data,
  groupBy = "Region",
  columns = list(
    States = colDef(
      aggregate = "count",
      format = list(
        aggregated = colFormat(suffix = " states")
      )
    ),
    Area = colDef(
      aggregate = "sum",
      format = colFormat(suffix = " mi²", separators = TRUE)
    )
  )
)
```

### Displaying missing values
Missing values are ignored by formatters and shown as empty cells by default.
You can customize their display text by setting `na` on a column:
```{r}
reactable(
  data.frame(
    n = c(1, 2, NA, 4, 5),
    x = c(55, 27, NA, NaN, 19),
    y = c(1, NA, 0.25, 0.55, NA)
  ),
  columns = list(
    x = colDef(na = "–", format = colFormat(prefix = "$")),
    y = colDef(na = "NA", format = colFormat(percent = TRUE))
  )
)
```

### Custom data formatting
If none of the built-in formatters apply to your data, you can use a
[custom cell renderer](#cell-rendering) instead.


## Custom Rendering
You can customize how data is displayed using an R or JavaScript function
that returns custom content. R render functions support
[Shiny HTML tags](https://shiny.rstudio.com/articles/tag-glossary.html) (or
[`htmltools`](https://mastering-shiny.org/advanced-ui.html#generating-html)) and
[`htmlwidgets`](https://www.htmlwidgets.org/),
while JavaScript render functions allow for more dynamic behavior.

You can also render content as HTML using `colDef(html = TRUE)`. Note that all
raw HTML is escaped by default.

See [Custom Rendering](custom-rendering.html) for details on how to
use render functions, and the [Demo Cookbook](cookbook/cookbook.html)
for even more examples of custom rendering.

### Cell rendering

#### R render function
```{r}
data <- MASS::Cars93[1:5, c("Manufacturer", "Model", "Type", "AirBags", "Price")]

reactable(data, columns = list(
  Model = colDef(cell = function(value, index) {
    # Render as a link
    url <- sprintf("https://wikipedia.org/wiki/%s_%s", data[index, "Manufacturer"], value)
    htmltools::tags$a(href = url, target = "_blank", as.character(value))
  }),
  AirBags = colDef(cell = function(value) {
    # Render as an X mark or check mark
    if (value == "None") "\u274c No" else "\u2714\ufe0f Yes"
  }),
  Price = colDef(cell = function(value) {
    # Render as currency
    paste0("$", format(value * 1000, big.mark = ","))
  })
))
```

#### JavaScript render function
```{r}
data <- MASS::Cars93[1:5, c("Manufacturer", "Model", "Type", "AirBags", "Price")]

reactable(data, columns = list(
  Model = colDef(html = TRUE, cell = JS("
    function(cellInfo) {
      // Render as a link
      const url = 'https://wikipedia.org/wiki/' + cellInfo.row['Manufacturer'] + '_' + cellInfo.value
      return '<a href=\"' + url + '\" target=\"_blank\">' + cellInfo.value + '</a>'
    }
  ")),
  AirBags = colDef(cell = JS("
    function(cellInfo) {
      // Render as an X mark or check mark
      return cellInfo.value === 'None' ? '\u274c No' : '\u2714\ufe0f Yes'
    }
  ")),
  Price = colDef(cell = JS("
    function(cellInfo) {
      // Render as currency
      return '$' + (cellInfo.value * 1000).toLocaleString()
    }
  "))
))
```

#### Embedding HTML widgets
```{r message=FALSE}
library(dplyr)
library(sparkline)

data <- chickwts %>%
  group_by(feed) %>%
  summarise(weight = list(weight)) %>%
  mutate(boxplot = NA, sparkline = NA)

reactable(data, columns = list(
  weight = colDef(cell = function(values) {
    sparkline(values, type = "bar", chartRangeMin = 0, chartRangeMax = max(chickwts$weight))
  }),
  boxplot = colDef(cell = function(value, index) {
    sparkline(data$weight[[index]], type = "box")
  }),
  sparkline = colDef(cell = function(value, index) {
    sparkline(data$weight[[index]])
  })
))
```

### Grouped cell rendering

> New in v0.2.3.9000 (unreleased)

```{r}
data <- MASS::Cars93[10:22, c("Manufacturer", "Model", "Type", "Price", "MPG.city")]

reactable(
  data,
  groupBy = c("Manufacturer", "Type"),
  columns = list(
    Manufacturer = colDef(
      # Render grouped cells without the row count
      grouped = JS("function(cellInfo) {
        return cellInfo.value
      }")
    ),
    Type = colDef(
      # Render grouped cells with the row count, only if there are multiple sub rows
      grouped = JS("function(cellInfo) {
        if (cellInfo.subRows.length > 1) {
          return cellInfo.value + ' (' + cellInfo.subRows.length + ')'
        }
        return cellInfo.value
      }")
    )
  )
)
```

### Aggregated cell rendering
```{r, message=FALSE}
library(dplyr)

set.seed(10)

data <- sample_n(tail(MASS::Cars93, 9), 30, replace = TRUE) %>%
  select(Manufacturer, Model, Type, Sales = Price)

reactable(
  data,
  groupBy = "Manufacturer",
  searchable = TRUE,
  columns = list(
    Model = colDef(aggregate = "unique"),
    Type = colDef(
      # Render aggregated value as a comma-separated list of unique values
      aggregated = JS("function(cellInfo) {
        const values = cellInfo.subRows.map(function(row) { return row['Type'] })
        const unique = values.reduce(function(obj, v) { obj[v] = true; return obj }, {})
        return Object.keys(unique).join(', ')
      }")
    ),
    Sales = colDef(
      aggregate = "sum",
      # Render aggregated cell as currency
      aggregated = JS("function(cellInfo) {
        return '$' + cellInfo.value.toFixed(2)
      }")
    )
  )
)
```

### Header rendering
```{r}
library(htmltools)

reactable(
  iris[1:5, ],
  defaultColDef = colDef(header = function(value) {
    units <- div(style = "color: #999", "cm")
    div(title = value, value, units)
  }),
  columns = list(
    Petal.Width = colDef(name = "Petal Width", html = TRUE, align = "left", header = JS("
      function(colInfo) {
        return colInfo.column.name + '<div style=\"color: #999\">cm</div>'
      }
    ")),
    Species = colDef(header = function(value) {
      tags$a(href = "https://wikipedia.org/wiki/List_of_Iris_species", title = "Iris species", value)
    })
  )
)
```


## Footers
You can add column footers by specifying content to render in `footer`.

`footer` can also be an R or JavaScript render function — see
[Custom Rendering](custom-rendering.html) for details on how to
use render functions.

#### R render function
```{r, message=FALSE}
library(dplyr)
library(htmltools)

data <- MASS::Cars93[18:47, ] %>%
  select(Manufacturer, Model, Type, Sales = Price)

reactable(
  data,
  defaultPageSize = 5,
  columns = list(
    Manufacturer = colDef(footer = "Total"),
    Sales = colDef(footer = function(values) sprintf("$%.2f", sum(values)))
  ),
  defaultColDef = colDef(footerStyle = list(fontWeight = "bold"))
)
```

#### JavaScript render function
```{r}
reactable(
  data,
  searchable = TRUE,
  defaultPageSize = 5,
  minRows = 5,
  columns = list(
    Manufacturer = colDef(footer = "Total"),
    Sales = colDef(
      footer = JS("function(colInfo) {
        let total = 0
        colInfo.data.forEach(function(row) {
          total += row[colInfo.column.id]
        })
        return '$' + total.toFixed(2)
      }")
    )
  ),
  defaultColDef = colDef(footerStyle = list(fontWeight = "bold"))
)
```

#### Embedding HTML widgets {#embedding-html-widgets-footers}
```{r}
library(sparkline)

reactable(
  iris[1:20, ],
  defaultPageSize = 5,
  bordered = TRUE,
  defaultColDef = colDef(footer = function(values) {
    if (!is.numeric(values)) return()
    sparkline(values, type = "box", width = 100, height = 30)
  })
)
```


## Expandable Row Details
You can make rows expandable with additional content through `details`, which
takes an R or JavaScript render function. See [Custom Rendering](custom-rendering.html)
for details on how to use render functions.

```{r}
reactable(iris[1:5, ], details = function(index) {
  htmltools::div(
    "Details for row: ", index,
    htmltools::tags$pre(paste(capture.output(iris[index, ]), collapse = "\n"))
  )
})
```

The details column can be customized by providing a `colDef()` instead. This can
be used to add a column name, render HTML content, or change the column width:
```{r}
reactable(iris[1:5, ], details = colDef(
  name = "More",
  details = JS("function(rowInfo) {
    return 'Details for row: ' + rowInfo.index +
      '<pre>' + JSON.stringify(rowInfo.row, null, 2) + '</pre>'
  }"),
  html = TRUE,
  width = 60
))
```

### Nested tables
With R render functions, you can render HTML tags, HTML widgets, and even nested tables:
```{r}
data <- unique(CO2[, c("Plant", "Type")])

reactable(data, details = function(index) {
  plant_data <- CO2[CO2$Plant == data$Plant[index], ]
  htmltools::div(style = "padding: 16px",
    reactable(plant_data, outlined = TRUE)
  )
})
```

### Conditional row details
R render functions support conditional rendering. If a render function
returns `NULL`, the row won't be expandable:
```{r}
reactable(iris[1:5, ], details = function(index) {
  if (index %in% c(3, 5)) {
    reactable(data.frame(x = c(1, 2, 3), y = c("a", "b", "c")), fullWidth = FALSE)
  }
})
```

### Multiple row details
You can add `details` to individual columns, and even show multiple details for a row:
```{r}
reactable(iris[1:5, ],
  details = function(index) {
    if (index %in% c(3, 5)) {
      reactable(data.frame(x = c(1, 2, 3), y = c("a", "b", "c")), fullWidth = FALSE)
    }
  },
  columns = list(
    Petal.Length = colDef(details = function(index) {
      paste("Petal.Length: ", iris[index, "Petal.Length"])
    }),
    Sepal.Length = colDef(format = colFormat(digits = 1), details = JS("
      function(rowInfo) {
        return 'Sepal.Length: ' + rowInfo.row['Sepal.Length']
      }
    "))
  )
)
```

### Default expanded rows

You can expand all rows by default by setting `defaultExpanded` to `TRUE`:
```{r}
reactable(
  iris[1:12, ],
  defaultPageSize = 4,
  details = function(index) paste("Details for row:", index),
  defaultExpanded = TRUE
)
```


## Conditional Styling
You can conditionally style a table using functions that return inline styles
or CSS classes. Just like with custom rendering, style functions can either be
in R or JavaScript.

See [Conditional Styling](conditional-styling.html) for details on how to
use style functions, and the [Demo Cookbook](cookbook/cookbook.html)
for even more examples of conditional styling.

### Cell styling

#### R style function
```{r}
reactable(sleep[1:6, ], columns = list(
  extra = colDef(
    style = function(value) {
      if (value > 0) {
        color <- "#008000"
      } else if (value < 0) {
        color <- "#e00000"
      } else {
        color <- "#777"
      }
      list(color = color, fontWeight = "bold")
    }
  )
))
```

#### JavaScript style function
```{r}
reactable(sleep[1:6, ], columns = list(
  extra = colDef(
    style = JS("function(rowInfo) {
      const value = rowInfo.row['extra']
      let color
      if (value > 0) {
        color = '#008000'
      } else if (value < 0) {
        color = '#e00000'
      } else {
        color = '#777'
      }
      return { color: color, fontWeight: 'bold' }
    }")
  )
))
```

### Row styling

#### R style function
```{r row_styling, eval=FALSE}
reactable(sleep[1:6, ], 
  rowStyle = function(index) {
    if (sleep[index, "extra"] < -1) {
      list(background = "rgba(0, 0, 0, 0.05)")
    }
  },
  rowClass = function(index) {
    if (sleep[index, "extra"] < -1) {
      "bold"
    }
  }
)
```

```{css}
.bold {
  font-weight: bold;
}
```

```{r ref.label="row_styling", echo=FALSE}
```

#### JavaScript style function
```{r}
reactable(sleep[1:6, ],
  rowStyle = JS("function(rowInfo) {
    if (rowInfo.row['extra'] < -1) {
      return { background: 'rgba(0, 0, 0, 0.05)' }
    }
  }"),
  rowClass = JS("function(rowInfo) {
    if (rowInfo.row['extra'] < -1) {
      return 'bold'
    }
  }")
)
```


## Table Styling

You can customize table styling using several options, which can all be combined:

### Highlight rows on hover

```{r}
reactable(iris[1:5, ], highlight = TRUE)
```

### Bordered

```{r}
reactable(iris[1:5, ], bordered = TRUE)
```

### Borderless

```{r}
reactable(iris[1:5, ], borderless = TRUE)
```

### Outlined

```{r}
reactable(iris[1:5, ], outlined = TRUE)
```

### Striped

```{r}
reactable(iris[1:5, ], striped = TRUE)
```

### Bordered + striped + highlighting

```{r}
reactable(iris[1:5, ], bordered = TRUE, striped = TRUE, highlight = TRUE)
```

### Outlined + borderless

```{r}
reactable(iris[1:5, ], outlined = TRUE, borderless = TRUE)
```

### Compact

```{r}
reactable(iris[1:5, ], compact = TRUE)
```

### No text wrapping
Long text is wrapped by default, but you can force text to fit on a single line
by setting `wrap` to `FALSE`:
```{r}
data <- aggregate(. ~ Species, iris, toString)

reactable(
  data,
  wrap = FALSE,
  resizable = TRUE,
  bordered = TRUE,
  columns = list(Petal.Length = colDef(name = "Petal Length (cm)", minWidth = 50))
)
```

### Fixed height + sticky header/footer
You can make tables scrollable by setting a fixed height or width.
Headers and footers are sticky by default, so they stay in place when scrolling:
```{r}
reactable(
  iris[1:20, ],
  height = 270,
  striped = TRUE,
  defaultColDef = colDef(
    footer = function(values, name) htmltools::div(name, style = list(fontWeight = 600))
  )
)
```

### Column widths
By default, columns have a minimum width of 100px and stretch to fill the table.
You can control the width of a column using the following arguments in `colDef()`:

- `minWidth` - minimum width of the column in pixels (defaults to `100`)
- `maxWidth` - maximum width of the column in pixels
- `width` - fixed width of the column in pixels (overrides `minWidth` and `maxWidth`)

When columns stretch, `minWidth` also controls the ratio at which columns grow.
For example, if a table consists of 3 columns having `minWidth = 100` each,
the columns will stretch at a ratio of `100:100:100`. Each column will take up 1/3
of the table's width and not shrink below 100px.

Another example: if a table consists of three columns having minimum widths of
200px, 100px, and 100px, the columns will take up 50%, 25%, and 25% of the table's
width respectively:

```{r}
reactable(
  MASS::Cars93[1:6, c("Make", "Type", "Weight")],
  columns = list(
    Make = colDef(minWidth = 200),   # 50% width, 200px minimum
    Type = colDef(minWidth = 100),   # 25% width, 100px minimum
    Weight = colDef(minWidth = 100)  # 25% width, 100px minimum
  ),
  bordered = TRUE
)
```

### No full width
Tables are full width by default, but you can shrink the table to fit its
contents by setting `fullWidth` to `FALSE`:
```{r}
reactable(
  MASS::Cars93[1:5, 1:5],
  fullWidth = FALSE,
  bordered = TRUE,
  defaultColDef = colDef(minWidth = 120)
)
```

You can also set a maximum or fixed width on the table:
```{r}
reactable(
  MASS::Cars93[1:5, 1:5],
  bordered = TRUE,
  defaultColDef = colDef(minWidth = 120),
  # Set a maximum width on the table:
  style = list(maxWidth = 650),
  # Or a fixed width:
  width = 650
)
```

### Vertical alignment

> New in v0.2.3.9000 (unreleased)

You can change the vertical alignment of cell content using the `vAlign` or
`headerVAlign` arguments in `colDef()` and `colGroup()`. `vAlign` controls the
alignment of data cells, while `headerVAlign` controls the alignment of header cells.
Possible options are `"top"` (the default), `"center"`, and `"bottom"`.

```{r}
library(dplyr)
library(htmltools)

data <- starwars[1:6, ] %>%
  select(character = name, height, mass, gender, homeworld, species)

reactable(
  data,
  columns = list(
    character = colDef(
      name = "Character / Species",
      # Show species under character names
      cell = function(value, index) {
        species <- data$species[index]
        species <- if (!is.na(species)) species else "Unknown"
        div(
          div(style = list(fontWeight = 600), value),
          div(style = list(fontSize = 12), species)
        )
      }
    ),
    species = colDef(show = FALSE)
  ),
  # Vertically center cells and bottom-align headers
  defaultColDef = colDef(vAlign = "center", headerVAlign = "bottom"),
  bordered = TRUE,
  width = 600
)
```

### Custom CSS
For more control over styling, you can add custom class names to the table and
apply your own CSS:
```{r}
reactable(
  iris[1:18, ],
  defaultPageSize = 6,
  borderless = TRUE,
  class = "my-tbl",
  defaultColDef = colDef(headerClass = "my-header"),
  columns = list(
    Sepal.Width = colDef(class = "my-col"),
    Petal.Width = colDef(class = "my-col")
  ),
  rowClass = "my-row"
)
```

```{css, echo=FALSE}
.my-tbl {
  border: 1px solid rgba(0, 0, 0, 0.1);
}

.my-header {
  border-width: 1px;
}

.my-col {
  border-right: 1px solid rgba(0, 0, 0, 0.05);
}

.my-row:hover {
  background-color: #f5f8ff;
}
```

In R Markdown documents, you can embed CSS using a `css` language chunk:
```{r, echo=FALSE, comment=NA}
cat('```{css, echo=FALSE}
.my-tbl {
  border: 1px solid rgba(0, 0, 0, 0.1);
}

.my-header {
  border-width: 1px;
}

.my-col {
  border-right: 1px solid rgba(0, 0, 0, 0.05);
}

.my-row:hover {
  background-color: #f5f8ff;
}
```')
```

The examples here embed CSS for demonstration, but it's sometimes better to add CSS
through an external style sheet. To learn more about adding custom CSS through an
external style sheet:

- [Using custom CSS in R Markdown documents](https://bookdown.org/yihui/rmarkdown/html-document.html#custom-css)
- [Using custom CSS in Shiny apps](https://shiny.rstudio.com/articles/css.html)


## Theming

Themes provide a powerful way to customize table styling that can be reused
across tables. You can either set theme variables to change the default styles
(e.g., row stripe color), or add your own custom CSS to specific elements of the table.

To apply a theme, provide a `reactableTheme()` to `theme`:
```{r}
reactable(
  iris[1:30, ],
  searchable = TRUE,
  striped = TRUE,
  highlight = TRUE,
  bordered = TRUE,
  theme = reactableTheme(
    borderColor = "#dfe2e5",
    stripedColor = "#f6f8fa",
    highlightColor = "#f0f5f9",
    cellPadding = "8px 12px",
    style = list(fontFamily = "-apple-system, BlinkMacSystemFont, Segoe UI, Helvetica, Arial, sans-serif"),
    searchInputStyle = list(width = "100%")
  )
)
```

### Global theme
To set the default theme for all tables, use the global `reactable.theme` option:
```{r}
options(reactable.theme = reactableTheme(
  color = "hsl(233, 9%, 87%)",
  backgroundColor = "hsl(233, 9%, 19%)",
  borderColor = "hsl(233, 9%, 22%)",
  stripedColor = "hsl(233, 12%, 22%)",
  highlightColor = "hsl(233, 12%, 24%)",
  inputStyle = list(backgroundColor = "hsl(233, 9%, 25%)"),
  selectStyle = list(backgroundColor = "hsl(233, 9%, 25%)"),
  pageButtonHoverStyle = list(backgroundColor = "hsl(233, 9%, 25%)"),
  pageButtonActiveStyle = list(backgroundColor = "hsl(233, 9%, 28%)")
))

reactable(
  iris[1:30, ],
  filterable = TRUE,
  showPageSizeOptions = TRUE,
  striped = TRUE,
  highlight = TRUE,
  details = function(index) paste("Details for row", index)
)
```

```{r, echo=FALSE}
options(reactable.theme = NULL)
```

### Nested selectors
You can use nested CSS selectors in theme styles to target the current element,
using `&` as the selector, or other child elements (just like in Sass).
This is useful for adding pseudo-classes like `&:hover`, or adding styles in a
certain context like `.outer-container &`.

For example, to highlight headers when sorting:
```{r}
reactable(
  iris[1:5, ],
  columns = list(Sepal.Length = colDef(sortable = FALSE)),
  showSortable = TRUE,
  theme = reactableTheme(
    headerStyle = list(
      "&:hover[aria-sort]" = list(background = "hsl(0, 0%, 96%)"),
      "&[aria-sort='ascending'], &[aria-sort='descending']" = list(background = "hsl(0, 0%, 96%)"),
      borderColor = "#555"
    )
  )
)
```

Or to apply a dark theme when a parent element has a certain class, like `.dark`:
```{r theme_toggle, eval=FALSE}
theme <- reactableTheme(
  style = list(".dark &" = list(color = "#fff", background = "#282a36")),
  cellStyle = list(".dark &" = list(borderColor = "rgba(255, 255, 255, 0.15)")),
  headerStyle = list(".dark &" = list(borderColor = "rgba(255, 255, 255, 0.15)")),
  paginationStyle = list(".dark &" = list(borderColor = "rgba(255, 255, 255, 0.15)")),
  rowHighlightStyle = list(".dark &" = list(background = "rgba(255, 255, 255, 0.04)")),
  pageButtonHoverStyle = list(".dark &" = list(background = "rgba(255, 255, 255, 0.08)")),
  pageButtonActiveStyle = list(".dark &" = list(background = "rgba(255, 255, 255, 0.1)"))
)

tbl <- reactable(iris[1:12, ], highlight = TRUE, defaultPageSize = 6, theme = theme)

# Simple theme toggle button
tags$button(onclick = "document.querySelector('.themeable-tbl').classList.toggle('dark')",
            "Toggle light/dark")

# Start with the dark theme enabled
div(class = "themeable-tbl dark", tbl)
```

```{r ref.label="theme_toggle", echo=FALSE}
```

### Dynamic theming
Themes can also be functions that return a `reactableTheme()` for
context-specific styling.

For example, to style tables in RStudio R Notebooks only when a dark
editor theme is active:
```{r, eval=FALSE}
options(reactable.theme = function() {
  theme <- reactableTheme(
    color = "hsl(233, 9%, 85%)",
    backgroundColor = "hsl(233, 9%, 19%)",
    borderColor = "hsl(233, 9%, 22%)",
    stripedColor = "hsl(233, 12%, 22%)",
    highlightColor = "hsl(233, 12%, 24%)",
    inputStyle = list(backgroundColor = "hsl(233, 9%, 25%)"),
    selectStyle = list(backgroundColor = "hsl(233, 9%, 25%)"),
    pageButtonHoverStyle = list(backgroundColor = "hsl(233, 9%, 25%)"),
    pageButtonActiveStyle = list(backgroundColor = "hsl(233, 9%, 28%)")
  )

  if (isTRUE(getOption("rstudio.notebook.executing"))) {
    if (requireNamespace("rstudioapi", quietly = TRUE) && rstudioapi::getThemeInfo()$dark) {
      return(theme)
    }
  }
})
```


## Column Groups
You can create column groups by passing a list of `colGroup()` definitions to `columnGroups`:
```{r}
reactable(
  iris[1:5, ],
  columns = list(
    Sepal.Length = colDef(name = "Length"),
    Sepal.Width = colDef(name = "Width"),
    Petal.Length = colDef(name = "Length"),
    Petal.Width = colDef(name = "Width")
  ),
  columnGroups = list(
    colGroup(name = "Sepal", columns = c("Sepal.Length", "Sepal.Width")),
    colGroup(name = "Petal", columns = c("Petal.Length", "Petal.Width"))
  )
)
```


## Column Resizing
You can make columns resizable by setting `resizable` to `TRUE`:
```{r}
reactable(MASS::Cars93[1:5, ], resizable = TRUE, wrap = FALSE, bordered = TRUE)
```


## Sticky Columns

> New in v0.2.3.9000 (unreleased)

You can make columns sticky when scrolling horizontally using the `sticky`
argument in `colDef()` or `colGroup()`. Set `sticky` to either `"left"`
or `"right"` to make the column stick to the left or right side.

```{r}
reactable(
  MASS::Cars93[1:5, ],
  columns = list(
    Manufacturer = colDef(
      sticky = "left",
      # Add a right border style to visually distinguish the sticky column
      style = list(borderRight = "1px solid #eee"),
      headerStyle = list(borderRight = "1px solid #eee")
    ),
    Make = colDef(
      sticky = "right",
      # Add a left border style to visually distinguish the sticky column
      style = list(borderLeft = "1px solid #eee"),
      headerStyle = list(borderLeft = "1px solid #eee")
    )
  ),
  defaultColDef = colDef(minWidth = 150)
)
```

> **NOTE:** Sticky columns are not supported in IE11.

Before v0.2.3.9000, columns could be made sticky by applying custom styles.
This method was harder to use and did not work well with column resizing,
row highlighting, or row striping. See [Sticky Columns](sticky-columns.html)
for examples of the previous method.

### Multiple sticky columns

```{r}
# Background style to visually distinguish sticky columns
sticky_style <- list(backgroundColor = "#f7f7f7")

reactable(
  MASS::Cars93[1:5, ],
  columns = list(
    Manufacturer = colDef(
      sticky = "left",
      style = sticky_style,
      headerStyle = sticky_style
    ),
    Model = colDef(
      sticky = "left",
      style = sticky_style,
      headerStyle = sticky_style
    ),
    Type = colDef(
      sticky = "left",
      style = sticky_style,
      headerStyle = sticky_style
    )
  ),
  resizable = TRUE,
  wrap = FALSE,
  bordered = TRUE
)
```

### Sticky column groups

If a column group is sticky, all columns in the group will automatically be
made sticky.

```{r}
reactable(
  MASS::Cars93[1:5, ],
  columnGroups = list(
    colGroup("Make", columns = c("Manufacturer", "Model"), sticky = "left"),
    colGroup("Price", columns = c("Min.Price", "Price", "Max.Price"), sticky = "left")
  ),
  defaultColDef = colDef(footer = "Footer"),
  resizable = TRUE,
  wrap = FALSE,
  bordered = TRUE
)
```


## Row Names and Row Headers

### Row names

Row names are shown by default if present. You can customize the row names column
using `".rownames"` as the column name:

```{r}
reactable(
  USPersonalExpenditure,
  columns = list(
    .rownames = colDef(name = "Category", sortable = TRUE)
  )
)
```

If row names haven't been set explicitly, you can force them to show by setting
`rownames` to `TRUE`:

```{r}
reactable(iris[1:5, ], rownames = TRUE)
```

### Row headers

> New in v0.2.3.9000 (unreleased)

You can mark up cells in a column as row headers by setting `rowHeader` to `TRUE`
in `colDef()`.

Use this to help users navigate the table using assistive technologies. When cells
are marked up as row headers, assistive technologies will read them aloud while
navigating through cells in the table.

Cells in the row names column are automatically marked up as row headers.

```{r}
data <- MASS::Cars93[1:5, c("Make", "Type", "Price", "MPG.city", "AirBags")]

reactable(
  data,
  columns = list(
    Make = colDef(rowHeader = TRUE, style = list(fontWeight = 600))
  ),
  bordered = TRUE
)
```


## Cell Click Actions
You can add cell click actions using the `onClick` argument, which accepts the
following values:

- `"expand"` to expand the row
- `"select"` to select the row
- A JavaScript function for a custom action, e.g., sending the click event to Shiny

### Expand on click
```{r}
reactable(
  iris[48:52, ],
  groupBy = "Species",
  details = function(index) paste("Details for row:", index),
  onClick = "expand",
  # Give rows a pointer cursor to indicate that they're clickable
  rowStyle = list(cursor = "pointer")
)
```

### Select on click
```{r}
reactable(iris[1:5, ], selection = "multiple", onClick = "select")
```

### Custom action

This example uses a custom click action to create custom "show details" action
buttons in each row of the table:

```{r}
data <- cbind(
  MASS::Cars93[1:5, c("Manufacturer", "Model", "Type", "Price")],
  details = NA
)

reactable(
  data,
  columns = list(
    # Render a "show details" button in the last column of the table.
    # This button won't do anything by itself, but will trigger the custom
    # click action on the column.
    details = colDef(
      name = "",
      sortable = FALSE,
      cell = function() htmltools::tags$button("Show details")
    )
  ),
  onClick = JS("function(rowInfo, colInfo) {
    // Only handle click events on the 'details' column
    if (colInfo.id !== 'details') {
      return
    }

    // Display an alert dialog with details for the row
    window.alert('Details for row ' + rowInfo.index + ':\\n' + JSON.stringify(rowInfo.row, null, 2))

    // Send the click event to Shiny, which will be available in input$show_details
    // Note that the row index starts at 0 in JavaScript, so we add 1
    if (window.Shiny) {
      Shiny.setInputValue('show_details', { index: rowInfo.index + 1 }, { priority: 'event' })
    }
  }")
)
```

> **WARNING:** Custom click actions are currently not accessible to keyboard
> users, and are generally not recommended. If they must be used, ensure that
> they can be triggered by a keyboard through other means, such as a button
> in the example above.


## Language Options

You can customize the language in the table by providing a set of `reactableLang()`
options to `language`:
```{r}
reactable(
  iris[1:30, ],
  searchable = TRUE,
  paginationType = "simple",
  language = reactableLang(
    searchPlaceholder = "Search...",
    noData = "No entries found",
    pageInfo = "{rowStart} to {rowEnd} of {rows} entries",
    pagePrevious = "\u276e",
    pageNext = "\u276f",

    # Accessible labels for assistive technologies such as screen readers.
    # These are already set by default, but don't forget to update them when
    # changing visible text.
    pagePreviousLabel = "Previous page",
    pageNextLabel = "Next page"
  )
)
```

### Global language options
To set the default language strings for all tables, use the global
`reactable.language` option:
```{r}
options(reactable.language = reactableLang(
  pageSizeOptions = "\u663e\u793a {rows}",
  pageInfo = "{rowStart} \u81f3 {rowEnd} \u9879\u7ed3\u679c,\u5171 {rows} \u9879",
  pagePrevious = "\u4e0a\u9875",
  pageNext = "\u4e0b\u9875"
))

reactable(iris[1:12, ], defaultPageSize = 4, showPageSizeOptions = TRUE)
```

```{r, echo=FALSE}
options(reactable.language = NULL)
```

## Shiny
To use reactable in Shiny apps, use `renderReactable()` and `reactableOutput()`:
```{r, eval=FALSE}
library(shiny)
library(reactable)

ui <- fluidPage(
  titlePanel("reactable example"),
  reactableOutput("table")
)

server <- function(input, output, session) {
  output$table <- renderReactable({
    reactable(iris)
  })
}

shinyApp(ui, server)
```

### Row selection
You can enable row selection by setting `selection` to `"single"`
for single selection, or `"multiple"` for multiple selection.

To get the selected rows in Shiny, use `getReactableState()`.
The selected rows are given as a vector of row indices
(e.g. `c(1, 6, 4)`) or `NULL` if no rows are selected.

```{r, eval=FALSE}
library(shiny)
library(reactable)

ui <- fluidPage(
  titlePanel("row selection example"),
  reactableOutput("table"),
  verbatimTextOutput("selected")
)

server <- function(input, output, session) {
  selected <- reactive(getReactableState("table", "selected"))

  output$table <- renderReactable({
    reactable(iris, selection = "multiple", onClick = "select")
  })

  output$selected <- renderPrint({
    print(selected())
  })

  observe({
    print(iris[selected(), ])
  })
}

shinyApp(ui, server)
```

> **NOTE:** `selectionId` can also be used to get the selected rows in Shiny,
> but this will be deprecated in a future release.

#### Default selected rows

You can preselect rows by specifying a vector of row indices in `defaultSelected`:
```{r}
reactable(iris[1:4, ], selection = "multiple", defaultSelected = c(1, 3))
```

#### Style selected rows

You can style selected rows using `rowSelectedStyle` in `reactableTheme()`:
```{r}
reactable(
  iris[1:4, ],
  selection = "multiple",
  defaultSelected = c(1, 3),
  borderless = TRUE,
  onClick = "select",
  theme = reactableTheme(
    rowSelectedStyle = list(backgroundColor = "#eee", boxShadow = "inset 2px 0 0 0 #ffa62d")
  )
)
```

Or using a `rowStyle` or `rowClass` JavaScript function:
```{r}
reactable(
  MASS::Cars93[10:22, c("Manufacturer", "Model", "Type", "Price", "MPG.city")],
  groupBy = "Manufacturer",
  selection = "multiple",
  defaultSelected = c(1, 2),
  borderless = TRUE,
  onClick = "select",
  rowStyle = JS("function(rowInfo) {
    if (rowInfo && rowInfo.selected) {
      return { backgroundColor: '#eee', boxShadow: 'inset 2px 0 0 0 #ffa62d' }
    }
  }")
)
```

#### Customize the selection column

You can customize the selection column using `".selection"` as the column name:
```{r}
reactable(
  MASS::Cars93[1:4, ],
  columns = list(
    .selection = colDef(
      width = 80,
      sticky = "left",
      style = list(cursor = "pointer"),
      headerStyle = list(cursor = "pointer")
    )
  ),
  selection = "multiple",
  onClick = "select",
  resizable = TRUE,
  wrap = FALSE,
  bordered = TRUE
)
```

### Update a reactable instance

You can update the selected rows, expanded rows, current page, or data using
`updateReactable()`:
```{r, eval=FALSE}
library(shiny)
library(reactable)

data <- MASS::Cars93[, 1:7]

ui <- fluidPage(
  actionButton("select_btn", "Select rows"),
  actionButton("clear_btn", "Clear selection"),
  actionButton("expand_btn", "Expand rows"),
  actionButton("collapse_btn", "Collapse rows"),
  actionButton("page_btn", "Change page"),
  selectInput("filter_type", "Filter type", unique(data$Type), multiple = TRUE),
  reactableOutput("table")
)

server <- function(input, output) {
  output$table <- renderReactable({
    reactable(
      data,
      filterable = TRUE,
      searchable = TRUE,
      selection = "multiple",
      details = function(index) paste("Details for row:", index)
    )
  })

  observeEvent(input$select_btn, {
    # Select rows
    updateReactable("table", selected = c(1, 3, 5))
  })

  observeEvent(input$clear_btn, {
    # Clear row selection
    updateReactable("table", selected = NA)
  })

  observeEvent(input$expand_btn, {
    # Expand all rows
    updateReactable("table", expanded = TRUE)
  })

  observeEvent(input$collapse_btn, {
    # Collapse all rows
    updateReactable("table", expanded = FALSE)
  })

  observeEvent(input$page_btn, {
    # Change current page
    updateReactable("table", page = 3)
  })

  observe({
    # Filter data
    filtered <- if (length(input$filter_type) > 0) {
      data[data$Type %in% input$filter_type, ]
    } else {
      data
    }
    updateReactable("table", data = filtered)
  })
}

shinyApp(ui, server)
```

### Get the state of a reactable instance

You can get the current page, page size, number of pages, or selected rows
using `getReactableState()`:
```{r, eval=FALSE}
library(shiny)
library(reactable)

ui <- fluidPage(
  actionButton("prev_page_btn", "Previous page"),
  actionButton("next_page_btn", "Next page"),
  reactableOutput("table"),
  verbatimTextOutput("table_state")
)

server <- function(input, output) {
  output$table <- renderReactable({
    reactable(
      iris,
      showPageSizeOptions = TRUE,
      selection = "multiple"
    )
  })

  output$table_state <- renderPrint({
    state <- req(getReactableState("table"))
    print(state)
  })

  observeEvent(input$prev_page_btn, {
    # Change to the previous page
    page <- getReactableState("table", "page")
    if (page > 1) {
      updateReactable("table", page = page - 1)
    }
  })

  observeEvent(input$next_page_btn, {
    # Change to the next page
    state <- getReactableState("table")
    if (state$page < state$pages) {
      updateReactable("table", page = state$page + 1)
    }
  })
}

shinyApp(ui, server)
```


## Cross-Widget Interactions

You can link selection and filtering with other HTML widgets in an R Markdown
document or Shiny app using [Crosstalk](https://rstudio.github.io/crosstalk).
To get started, install the `crosstalk` package and wrap your data frame in a
`crosstalk::SharedData` object:

```{r, eval=FALSE}
install.packages("crosstalk")

library(crosstalk)

data <- SharedData$new(iris)
```

Then, pass the shared data to `reactable()` and any other Crosstalk-compatible
HTML widget or filter input:

```{r, eval=FALSE}
reactable(data)

filter_slider("sepal_length", "Sepal Length", data, ~Sepal.Length)
```

For more examples and a list of Crosstalk-compatible widgets, check out
[Using Crosstalk](https://rstudio.github.io/crosstalk/using.html)
in the Crosstalk package documentation.

### Filtering

Tables can be filtered by widgets that support Crosstalk's filtering API,
such as Crosstalk's `filter_checkbox()`, `filter_slider()`, and `filter_select()` inputs:

```{r}
library(crosstalk)

cars <- MASS::Cars93[1:20, c("Manufacturer", "Model", "Type", "Price")]
data <- SharedData$new(cars)

bscols(
  widths = c(3, 9),
  list(
    filter_checkbox("type", "Type", data, ~Type),
    filter_slider("price", "Price", data, ~Price, width = "100%"),
    filter_select("mfr", "Manufacturer", data, ~Manufacturer)
  ),
  reactable(data, minRows = 10)
)
```

### Linked selection

Table selection state can be linked with other widgets that support Crosstalk's
linked selection (or linked brushing) API.

In this example, you can select rows to highlight points on the map, or select areas
on the map to highlight rows in the table.

```{r, message=FALSE}
library(crosstalk)
library(leaflet)
library(dplyr)

# A SpatialPointsDataFrame for the map.
# Set a group name to share data points with the table.
brew_sp <- SharedData$new(breweries91, group = "breweries")

# A regular data frame (without coordinates) for the table.
# Use the same group name as the map data.
brew_data <- as_tibble(breweries91) %>%
  select(brewery, address, village, founded) %>%
  SharedData$new(group = "breweries")

map <- leaflet(brew_sp) %>%
  addTiles() %>%
  addMarkers()

tbl <- reactable(
  brew_data,
  selection = "multiple",
  onClick = "select",
  rowStyle = list(cursor = "pointer"),
  minRows = 10
)

htmltools::browsable(
  htmltools::tagList(map, tbl)
)
```

## JavaScript API

> New in v0.2.3.9000 (unreleased)

You can use the JavaScript API to create custom interactive controls for your table
without the use of Shiny, or add cross-widget interactions beyond what Crosstalk provides.

See the [JavaScript API guide](javascript-api.html) for details on how to use the
JavaScript API in R Markdown documents or Shiny apps, and the full API reference.

### CSV download button

```{r}
library(htmltools)
library(fontawesome)

data <- MASS::Cars93[1:15, c("Manufacturer", "Model", "Type", "Price")]

htmltools::browsable(
  tagList(
    tags$button(
      tagList(fontawesome::fa("download"), "Download as CSV"),
      onclick = "Reactable.downloadDataCSV('cars-download-table', 'cars.csv')"
    ),

    reactable(
      data,
      searchable = TRUE,
      defaultPageSize = 5,
      elementId = "cars-download-table"
    )
  )
)
```

### Custom column filter

```{r}
library(htmltools)

data <- MASS::Cars93[1:15, c("Manufacturer", "Model", "Type", "Price")]

htmltools::browsable(
  tagList(
    div(
      div(tags$label("Filter Type", `for` = "cars-type-filter")),
      tags$select(
        id = "cars-type-filter",
        onchange = "Reactable.setFilter('cars-filter-table', 'Type', this.value)",
        tags$option("All", value = ""),
        lapply(unique(data$Type), tags$option)
      )
    ),
    
    tags$hr("aria-hidden" = "false"),

    reactable(data, defaultPageSize = 5, elementId = "cars-filter-table")
  )
)
```

### Custom search input

```{r}
library(htmltools)

data <- MASS::Cars93[1:15, c("Manufacturer", "Model", "Type", "Price")]

htmltools::browsable(
  tagList(
    div(
      style = "margin-bottom: 12px",
      tags$input(
        type = "text",
        placeholder = "Search for cars...",
        style = "padding: 4px 8px; width: 100%",
        oninput = "Reactable.setSearch('cars-search-table', this.value)"
      )
    ),

    reactable(data, defaultPageSize = 5, elementId = "cars-search-table")
  )
)
```

### Column grouping select

```{r}
library(dplyr)
library(htmltools)

set.seed(10)

data <- sample_n(tail(MASS::Cars93, 9), 30, replace = TRUE) %>%
  select(Manufacturer, Model, Type, Sales = Price)

htmltools::browsable(
  tagList(
    div(tags$label("Group by", `for` = "cars-grouping-select")),
    tags$select(
      id = "cars-grouping-select",
      onchange = "Reactable.setGroupBy('cars-grouping-table', this.value ? [this.value] : [])",
      tags$option("None", value = ""),
      lapply(c("Manufacturer", "Model", "Type"), tags$option)
    ),

    tags$hr("aria-hidden" = "false"),

    reactable(
      data,
      columns = list(
        Manufacturer = colDef(aggregate = "unique"),
        Model = colDef(aggregate = "unique"),
        Type = colDef(aggregate = "unique"),
        Sales = colDef(aggregate = "sum", format = colFormat(currency = "USD"))
      ),
      defaultPageSize = 5,
      minRows = 5,
      elementId = "cars-grouping-table"
    )
  )
)
```

### Row expansion toggle button

```{r}
library(htmltools)

data <- MASS::Cars93[1:5, c("Manufacturer", "Model", "Type", "Price")]

htmltools::browsable(
  tagList(
    tags$button(
      "Expand/collapse all",
      onclick = "Reactable.toggleAllRowsExpanded('cars-expansion-table')"
    ),

    reactable(
      data,
      groupBy = "Manufacturer",
      defaultPageSize = 5,
      elementId = "cars-expansion-table"
    )
  )
)
```


```{css echo=FALSE}
/* rmarkdown html documents */
.main-container {
  font-family: -apple-system, BlinkMacSystemFont, "Segoe UI", Helvetica, Arial, sans-serif;
}

.main-container blockquote {
  font-size: inherit;
}

/* pkgdown articles */
.contents {
  font-family: -apple-system, BlinkMacSystemFont, "Segoe UI", Helvetica, Arial, sans-serif;
}

.contents blockquote {
  font-size: inherit;
}

.reactable {
  font-size: 14px;
}
```
