---
title: 'JavaScript API'
output:
  html_document:
    toc: true
    toc_float:
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reactable)
library(htmltools)

propsTable <- function(props) {
  tags$div(
    style = "overflow: auto;",
    tabindex = "0",
    tags$table(
      class = "props-tbl",
      tags$thead(
        tags$tr(
          lapply(colnames(props), function(name) tags$th(name))
        )
      ),
      tags$tbody(
        apply(props, 1, function(row) {
          tags$tr(
            tags$th(scope = "row", tags$code(row[["Property"]])),
            tags$td(tags$code(row[["Example"]])),
            tags$td(row[["Description"]])
          )
        })
      )
    )
  )
}
```

## Introduction

The JavaScript API lets you manipulate and access tables from JavaScript.
You can use this to create custom interactive controls for your table without
the use of Shiny, or add cross-widget interactions beyond what Crosstalk provides.

Common use cases for the JavaScript API include:

- Export buttons to download table data to a CSV file
- Custom filter inputs located outside of the table
- Toggle buttons for row grouping or row expansion

#### Example: CSV download button

```{r}
library(htmltools)

htmltools::browsable(
  tagList(
    tags$button("Download as CSV", onclick = "Reactable.downloadDataCSV('cars-table')"),

    reactable(
      MASS::Cars93[, 1:5],
      defaultPageSize = 5,
      elementId = "cars-table"
    )
  )
)
```

## Using the JavaScript API

To use the JavaScript API, your table must first have a unique ID that distinguishes
it from other tables:

- For tables in static documents (e.g., R Markdown), add an ID through the
  `elementId` argument in `reactable()`:

  ```r
  reactable(MASS::Cars93, elementId = "cars-table")
  ```

- For tables in Shiny apps, the ID will be the Shiny output ID specified in
  `reactableOutput()`. For example, the table in this app uses `cars_table` as its ID:

  ```r
  library(shiny)

  ui <- fluidPage(
    reactableOutput("cars_table")
  )

  server <- function(input, output) {
    output$cars_table <- renderReactable({
      reactable(MASS::Cars93)
    })
  }

  shinyApp(ui, server)
  ```

Once your table has an ID, you can use any of the `Reactable` JavaScript functions
with that table ID. For example, to download data from the `cars-table` table to a
CSV file, the JavaScript code would look like this:

```js
// Download the "cars-table" data to a CSV file named 'cars.csv'
Reactable.downloadDataCSV('cars-table', 'cars.csv')
```

To try this out interactively, you can open your
[browser's developer tools](https://developer.mozilla.org/en-US/docs/Learn/Common_questions/What_are_browser_developer_tools)
and run this function in the
[JavaScript console](https://developer.mozilla.org/en-US/docs/Learn/Common_questions/What_are_browser_developer_tools#the_javascript_console).

### Creating custom interactive controls

Most users will likely want to use the JavaScript API through an interactive control,
such as a button, so they could decide when to download the table data. Using HTML,
you can create a `<button>` element with an `onclick` action that calls
the `Reactable` JavaScript function.

This example uses the `htmltools` package to render a CSV download button.
You can copy this code into an R console to view the output:

```{r, eval=FALSE}
library(htmltools)

htmltools::browsable(
  tagList(
    tags$button("Download as CSV", onclick = "Reactable.downloadDataCSV('cars-table', 'cars.csv')"),

    reactable(MASS::Cars93[, 1:5], elementId = "cars-table")
  )
)
```

::: {.callout-note}
**Note:** `htmltools::browsable()` is a convenient way to view the rendered HTML when
copying code into the console. It isn't required to render HTML in R Markdown documents
or Shiny apps.
:::

To reuse this button in other tables, you can also convert it into a function that
generates download buttons:

```{r, eval=FALSE}
library(htmltools)

csvDownloadButton <- function(tableId, label = "Download as CSV", filename = "data.csv") {
  htmltools::tags$button(
    label,
    onclick = sprintf("Reactable.downloadDataCSV('%s', '%s')", tableId, filename)
  )
}

htmltools::browsable(
  tagList(
    csvDownloadButton("cars-table", "Download as CSV", filename = "cars.csv"),

    reactable(MASS::Cars93[, 1:5], elementId = "cars-table")
  )
)
```

For more examples of custom controls that use the JavaScript API, check out the
[JavaScript API examples](examples.html#javascript-api).

## JavaScript API Reference

### `Reactable.downloadDataCSV()`

Downloads the table data to a CSV file. The downloaded file will be named `data.csv`
by default, but you can customize this using the optional `filename` argument.

The downloaded data will include any filters that have been applied, and
exclude any sorting or grouping. Hidden columns will also be included, but
this may be customizable in the future.

For further customization, you can use the `options` argument to only include specific columns,
exclude column headers, change the field separator, or change the decimal separator (new in v0.4.0).

```ts
Reactable.downloadDataCSV(
  tableId: string,
  filename = "data.csv",
  // New in v0.4.0
  options?: {
    columnIds: string[],
    headers: true,
    sep: ',',
    dec: '.'
  }
)
```

#### Examples

```js
// Download table data to a file named data.csv
Reactable.downloadDataCSV('cars-table')

// Download table data to a file named cars93.csv
Reactable.downloadDataCSV('cars-table', 'cars93.csv')

// Download table data to a tab-separated values file
Reactable.downloadDataCSV('cars-table', 'cars93.tsv', { sep: '\t' })

// Download table data with custom columns and headers excluded
Reactable.downloadDataCSV('cars-table', 'cars93.csv', {
  columnIds: ['Model', 'Type'],
  headers: false
})
```

### `Reactable.getDataCSV()`

::: {.callout}
New in v0.4.0
:::

Gets the table data as a CSV string. Same as [`Reactable.downloadDataCSV()`](#reactable-downloaddatacsv)
but does not download the data.

```ts
Reactable.getDataCSV(
  tableId: string,
  options?: {
    columnIds: string[],
    headers: true,
    sep: ',',
    dec: '.'
  }
)
```

#### Value

A string with the table data in CSV format.

#### Examples

```js
// Get table data as a CSV string
const csv = Reactable.getDataCSV('cars-table')
console.log(csv)

// Get table data as a tab-separated values string
const tsv = Reactable.getDataCSV('cars-table', { sep: '\t' })
console.log(tsv)
```

### `Reactable.setSearch()`

Sets the search value of a table. To clear the search, set the value to `undefined`.

```ts
Reactable.setSearch(
  tableId: string,
  value: any
)
```

#### Examples

```js
// Set the search value to "midsize"
Reactable.setSearch('cars-table', 'midsize')

// Clear the search value
Reactable.setSearch('cars-table', undefined)
```

### `Reactable.setFilter()`

Sets the filter value of a column. To clear the column filter, set the value
to `undefined`.

```ts
Reactable.setFilter(
  tableId: string,
  columnId: string,
  value: any
)
```

#### Examples

```js
// Set the filter value of the "Type" column to "midsize"
Reactable.setFilter('cars-table', 'Type', 'midsize')

// Clear the filter value of the "Type" column
Reactable.setFilter('cars-table', 'Type', undefined)
```

### `Reactable.setAllFilters()`

Sets all column filter values in the table. To clear the column filters,
set `filters` to an empty array, `[]`.

```ts
Reactable.setAllFilters(
  tableId: string,
  filters: Array<{ id: string, value: any }>
)
```

#### Examples

```js
// Set the column filters for the "Type" column
Reactable.setAllFilters('cars-table', [{ id: 'Type', value: 'midsize' }])

// Set the column filters for the "Type" and "Model" columns
Reactable.setAllFilters('cars-table', [
  { id: 'Type', value: 'midsize' },
  { id: 'Model', value: 'legend' }
])

// Clear all column filters
Reactable.setAllFilters([])
```

### `Reactable.toggleGroupBy()`

Toggles the `groupBy` state for a column between grouped and ungrouped.
To enable or disable grouping explicitly, set the optional `isGrouped` argument
to `true` or `false`.

```ts
Reactable.toggleGroupBy(
  tableId: string,
  columnId: string,
  isGrouped?: boolean
)
```

#### Examples

```js
// Toggle groupBy state for the "Type" column
Reactable.toggleGroupBy('cars-table', 'Type')

// Enable grouping for the "Type" column
Reactable.toggleGroupBy('cars-table', 'Type', true)

// Disable grouping for the "Type" column
Reactable.toggleGroupBy('cars-table', 'Type', false)
```

### `Reactable.setGroupBy()`

Sets the `groupBy` columns for the table. To clear the `groupBy` columns,
set `columnIds` to an empty array, `[]`.

```ts
Reactable.setGroupBy(
  tableId: string,
  columnIds: string[]
)
```

#### Examples

```js
// Set the groupBy columns to "Type" and "Manufacturer"
Reactable.setGroupBy('cars-table', ['Type', 'Manufacturer'])

// Clear the groupBy columns
Reactable.setGroupBy('cars-table', [])
```

### `Reactable.toggleAllRowsExpanded()`

Toggles the expanded state of all rows in the table between expanded and collapsed.
To expand or collapse rows explicitly, set the optional `isExpanded` argument to
`true` or `false`.

```ts
Reactable.toggleAllRowsExpanded(
  tableId: string,
  isExpanded?: boolean
)
```

#### Examples

```js
// Toggle expanded state for all rows
Reactable.toggleAllRowsExpanded('cars-table')

// Expand all rows
Reactable.toggleAllRowsExpanded('cars-table', true)

// Collapse all rows
Reactable.toggleAllRowsExpanded('cars-table', false)
```

### `Reactable.setMeta()`

::: {.callout}
New in v0.4.0
:::

Sets the custom metadata for the table. `meta` can either be an object with new
values, or a function that takes the previous metadata object and returns new values.

New values are merged into the current metadata, so only the values specified in
`meta` will be updated. To clear all metadata, set `meta` to `undefined`.

```ts
Reactable.setMeta(
  tableId: string,
  meta?: object | Function
)
```

#### Examples

```js
// Set metadata, updating the values for the `count`, `enabled`, and `formatter` properties
Reactable.setMeta('cars-table', { count: 123, enabled: true, formatter: value => '$' + value })

// Set metadata using a function that increments `count` and toggles an `enabled` boolean
Reactable.setMeta('cars-table', prevMeta => {
  return { count: prevMeta.count + 1, enabled: !prevMeta.enabled }
})

// Clear metadata
Reactable.setMeta('cars-table', undefined)
```

### `Reactable.toggleHideColumn()`

::: {.callout}
New in v0.4.0
:::

Toggles the hidden state for a column between hidden and shown.
To hide or show a column explicitly, set the optional `isHidden` argument to `true` or `false`.

```ts
Reactable.toggleHideColumn(
  tableId: string,
  columnId: string,
  isHidden?: boolean
)
```

#### Examples

```js
// Toggle hidden state for the "Type" column
Reactable.toggleHideColumn('cars-table', 'Type')

// Hide the "Type" column
Reactable.toggleHideColumn('cars-table', 'Type', true)

// Show the "Type" column
Reactable.toggleHideColumn('cars-table', 'Type', false)
```

### `Reactable.setHiddenColumns()`

::: {.callout}
New in v0.4.0
:::

Sets the hidden columns for the table. `columnIds` can either be an array of column IDs, or a
function that takes the previous hidden column IDs and returns new column IDs.

To clear the hidden columns, set `columnIds` to an empty array, `[]`.

```ts
Reactable.setHiddenColumns(
  tableId: string,
  columnIds: string[] | Function
)
```

#### Examples

```js
// Set the hidden columns to "Type" and "Manufacturer"
Reactable.setHiddenColumns('cars-table', ['Type', 'Manufacturer'])

// Set hidden columns using a function that adds "Type" to the existing hidden columns
Reactable.setHiddenColumns('cars-table', prevHiddenColumns => {
  return prevHiddenColumns.concat('Type')
})

// Clear the hidden columns
Reactable.setHiddenColumns('cars-table', [])
```

### `Reactable.setData()`

::: {.callout}
New in v0.4.0
:::

Updates the table data. `data` can either be in row or column format. In row format, `data` is an array
of row objects. In column format, `data` is an object containing arrays of column values. `data` should
have all the same columns as the original table data.

When updating data, the selected rows and current page will reset by default. All other state will persist,
including sorting, filtering, and grouping state.

```ts
Reactable.setData(
  tableId: string,
  data: Array<object> | object
)
```

#### Examples

```js
// Update data in row format
Reactable.setData('cars-table', [
  { Model: 'Legend', Type: 'Midsize' },
  { Model: 'Integra', Type: 'Small' }
])

// Update data in column format
Reactable.setData('cars-table', {
  Model: ['Legend', 'Integra'],
  Type: ['Midsize', 'Small']
})
```

### `Reactable.getState()`

Gets the current state of a table.

```ts
Reactable.getState(tableId: string)
```

#### Value {#state-properties}

An object with the following properties:

```{r, echo=FALSE, asis=TRUE}
stateProps <- dplyr::tribble(
  ~Property, ~Example, ~Description,
  "sorted", '[{ id: "Petal.Length", desc: true }, ...]', "columns being sorted in the table",
  "page", "2", "page index (zero-based)",
  "pageSize", "10", "page size",
  "pages", "5", "number of pages",
  "filters", '[{ id: "Species", value: "petal" }]', "column filter values",
  "searchValue", '"petal"', "table search value",
  "selected", '[0, 1, 4]', "selected row indices (zero-based)",
  "pageRows", '[{ Petal.Length: 1.7, Species: "setosa" }, ...]', "current row data on the page",
  "sortedData", '[{ Petal.Length: 1.7, Species: "setosa" }, ...]', "current row data in the table (after sorting, filtering, grouping)",
  "data", '[{ Petal.Length: 1.7, Species: "setosa" }, ...]', "original row data in the table",
  "meta", '{ custom: 123 }', tagList("custom table metadata from", tags$code("reactable()"), "(new in v0.4.0)"),
  "hiddenColumns", '["Petal.Length"]', "columns being hidden in the table"
)

propsTable(stateProps)
```

#### Examples

```js
Reactable.getState('cars-table')
// { page: 2, searchValue: 'petal', ... }
```

### `Reactable.onStateChange()`

::: {.callout}
New in v0.4.0
:::

Sets up a function that will be called whenever the table state changes.

`listenerFn` should be a function that takes the current table state, an object with
properties described at [state properties](#state-properties).

`Reactable.onStateChange()` can only be run when the table has finished rendering,
which isn't guaranteed to happen at page load time. To ensure that the table is ready
before running `Reactable.onStateChange()`, you can either use `htmlwidgets::onRender()`
for static widgets and Shiny outputs, or `htmlwidgets::onStaticRenderComplete()` for
static widgets only.

```ts
Reactable.onStateChange(
  tableId: string,
  listenerFn: Function
)
```

#### Value

`Reactable.onStateChange` returns a function that can be called to cancel the listener function.

#### Examples

```js
Reactable.onStateChange('cars-table', state => {
  console.log('selected rows:', state.selected)
})

// Cancel the listener function
const cancel = Reactable.onStateChange('cars-table', listenerFn)
cancel()
```

##### Static widgets

```{r, eval=FALSE}
library(htmltools)

data <- MASS::Cars93[, c("Manufacturer", "Model", "Type", "Price")]

browsable(
  tagList(
    "Table state:",
    tags$pre(id = "tbl-state", "{}"),
    reactable(
      data,
      searchable = TRUE,
      selection = "multiple",
      onClick = "select",
      elementId = "tbl"
    ),
    htmlwidgets::onStaticRenderComplete("
      Reactable.onStateChange('tbl', state => {
        const { selected, sorted, searchValue, pageIndex } = state
        document.getElementById('tbl-state').textContent = JSON.stringify({
          selected,
          sorted,
          searchValue,
          pageIndex
        })
      })
    ")
  )
)
```

##### Shiny outputs

```{r, eval=FALSE}
library(shiny)

data <- MASS::Cars93[, c("Manufacturer", "Model", "Type", "Price")]

ui <- fluidPage(
  reactableOutput("tbl"),
  verbatimTextOutput("tbl_state")
)

server <- function(input, output) {
  output$tbl <- renderReactable({
    tbl <- reactable(data, searchable = TRUE)

    htmlwidgets::onRender(tbl, "() => {
      Reactable.onStateChange('tbl', state => {
        const { sorted, searchValue } = state
        Shiny.setInputValue('tbl_state', { sorted, searchValue })
      })
    }")
  })

  output$tbl_state <- renderPrint({
    writeLines("Table state:\n")
    print(input$tbl_state)
  })
}

shinyApp(ui, server)
```

### `Reactable.gotoPage()`

::: {.callout}
New in v0.4.4.9000
:::

Changes the current page. `pageIndex` can either be a number or a function that takes
the previous `pageIndex` and returns the new value.

`pageIndex` is zero-based, so the first page would have a `pageIndex` of `0`.
If `pageIndex` is outside the valid `pageIndex` range, `Reactable.gotoPage()` will do nothing.

```ts
Reactable.gotoPage(
  tableId: string,
  pageIndex?: number | Function
)
```

#### Examples

```js
// Go to page index 0 (the first page)
Reactable.gotoPage('cars-table', 0)

// Go to page index 2 (the third page)
Reactable.gotoPage('cars-table', 2)

// Go to the next page
Reactable.gotoPage('cars-table', prevPage => prevPage + 1)

// Go to the previous page
Reactable.gotoPage('cars-table', prevPage => prevPage - 1)
```

### `Reactable.setPageSize()`

::: {.callout}
New in v0.4.4.9000
:::

Sets the current page size.

```ts
Reactable.gotoPage(
  tableId: string,
  pageSize?: number
)
```

#### Examples

```js
// Set the page size to 10
Reactable.setPageSize('cars-table', 10)
```


```{css echo=FALSE}
/* rmarkdown html documents */
.main-container {
  font-family: -apple-system, BlinkMacSystemFont, "Segoe UI", Helvetica, Arial, sans-serif;
}

.main-container blockquote {
  font-size: inherit;
}

h3 {
  font-weight: 600;
}
```
