---
title: "JavaScript API"
output:
  html_document:
    toc: true
    toc_float:
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reactable)
library(htmltools)

propsTable <- function(props) {
  tags$table(
    class = "props-tbl",
    tags$thead(
      tags$tr(
        lapply(colnames(props), function(name) tags$th(name))
      )
    ),
    tags$tbody(
      apply(props, 1, function(row) {
        tags$tr(
          tags$th(scope = "row", tags$code(row[["Property"]])),
          tags$td(tags$code(row[["Example"]])),
          tags$td(row[["Description"]])
        )
      })
    )
  )
}
```

> New in v0.2.3.9000 (unreleased)

## Introduction

The JavaScript API lets you manipulate and access tables from JavaScript.
You can use this to create custom interactive controls for your table without
the use of Shiny, or add cross-widget interactions beyond what Crosstalk provides.

Common use cases for the JavaScript API include:

- Export buttons to download table data to a CSV file
- Custom filter inputs located outside of the table
- Toggle buttons for row grouping or row expansion

#### Example: CSV download button

```{r}
library(htmltools)

htmltools::browsable(
  tagList(
    tags$button("Download as CSV", onclick = "Reactable.downloadDataCSV('cars-table')"),

    reactable(
      MASS::Cars93[, 1:5],
      defaultPageSize = 5,
      elementId = "cars-table"
    )
  )
)
```

## Using the JavaScript API

To use the JavaScript API, your table must first have a unique ID that distinguishes
it from other tables:

- For tables in static documents (e.g., R Markdown), add an ID through the
  `elementId` argument in `reactable()`:

  ```r
  reactable(MASS::Cars93, elementId = "cars-table")
  ```

- For tables in Shiny apps, the ID will be the Shiny output ID specified in
  `reactableOutput()`. For example, the table in this app uses `cars_table` as its ID:

  ```r
  library(shiny)

  ui <- fluidPage(
    reactableOutput("cars_table")
  )

  server <- function(input, output) {
    output$cars_table <- renderReactable({
      reactable(MASS::Cars93)
    })
  }

  shinyApp(ui, server)
  ```

Once your table has an ID, you can use any of the `Reactable` JavaScript functions
with that table ID. For example, to download data from the `cars-table` table to a
CSV file, the JavaScript code would look like this:

```js
// Download the "cars-table" data to a CSV file named 'cars.csv'
Reactable.downloadDataCSV('cars-table', 'cars.csv')
```

To try this out interactively, you can open your
[browser's developer tools](https://developer.mozilla.org/en-US/docs/Learn/Common_questions/What_are_browser_developer_tools)
and run this function in the
[JavaScript console](https://developer.mozilla.org/en-US/docs/Learn/Common_questions/What_are_browser_developer_tools#the_javascript_console).

### Creating custom interactive controls

Most users will likely want to use the JavaScript API through an interactive control,
such as a button, so they could decide when to download the table data. Using HTML,
you can create a `<button>` element with an `onclick` action that calls
the `Reactable` JavaScript function.

This example uses the `htmltools` package to render a CSV download button.
You can copy this code into an R console to view the output:

```{r, eval=FALSE}
library(htmltools)

htmltools::browsable(
  tagList(
    tags$button("Download as CSV", onclick = "Reactable.downloadDataCSV('cars-table', 'cars.csv')"),

    reactable(MASS::Cars93[, 1:5], elementId = "cars-table")
  )
)
```

> **NOTE**: `htmltools::browsable()` is a convenient way to view the rendered HTML when
> copying code into the console. It isn't required to render HTML in R Markdown documents
> or Shiny apps.

To reuse this button in other tables, you can also convert it into a function that
generates download buttons:

```{r, eval=FALSE}
library(htmltools)

csvDownloadButton <- function(tableId, label = "Download as CSV", filename = "data.csv") {
  htmltools::tags$button(
    label,
    onclick = sprintf("Reactable.downloadDataCSV('%s', '%s')", tableId, filename)
  )
}

htmltools::browsable(
  tagList(
    csvDownloadButton("cars-table", "Download as CSV", filename = "cars.csv"),

    reactable(MASS::Cars93[, 1:5], elementId = "cars-table")
  )
)
```

For more examples of custom controls that use the JavaScript API, check out the
[JavaScript API examples](examples.html#javascript-api).

## JavaScript API Reference

### `Reactable.downloadDataCSV()`

Downloads the table data to a CSV file. The downloaded file is named `data.csv`
by default, but you can customize this using the optional `filename` argument.

The downloaded data will include any filters that have been applied, and
exclude any sorting or grouping. Hidden columns will also be included, but
this may be customizable in the future.

```js
Reactable.downloadDataCSV(
  tableId: string,
  filename = "data.csv"
)
```

#### Examples

```js
// Download table data to a file named data.csv
Reactable.downloadDataCSV('cars-table')

// Download table data to a file named cars93.csv
Reactable.downloadDataCSV('cars-table', 'cars93.csv')
```

### `Reactable.setSearch()`

Sets the search value of a table. To clear the search, set the value to `undefined`.

```js
Reactable.setSearch(
  tableId: string,
  value: any
)
```

#### Examples

```js
// Set the search value to "midsize"
Reactable.setSearch('cars-table', 'midsize')

// Clear the search value
Reactable.setSearch('cars-table', undefined)
```

### `Reactable.setFilter()`

Sets the filter value of a column. To clear the column filter, set the value
to `undefined`.

```js
Reactable.setFilter(
  tableId: string,
  columnId: string,
  value: any
)
```

#### Examples

```js
// Set the filter value of the "Type" column to "midsize"
Reactable.setFilter('cars-table', 'Type', 'midsize')

// Clear the filter value of the "Type" column
Reactable.setFilter('cars-table', 'Type', undefined)
```

### `Reactable.setAllFilters()`

Sets all column filter values in the table. To clear the column filters,
set `filters` to an empty array, `[]`.

```js
Reactable.setAllFilters(
  tableId: string,
  filters: Array<{ id: string, value: any }>
)
```

#### Examples

```js
// Set the column filters for the "Type" column
Reactable.setAllFilters('cars-table', [{ id: 'Type', value: 'midsize' }])

// Set the column filters for the "Type" and "Model" columns
Reactable.setAllFilters('cars-table', [{ id: 'Type', value: 'midsize' }, { id: 'Model', value: 'legend' }])

// Clear all column filters
Reactable.setAllFilters([])
```

### `Reactable.toggleGroupBy()`

Toggles the `groupBy` state for a column between grouped and ungrouped.
To enable or disable grouping explicitly, set the optional `isGrouped` argument
to `true` or `false`.

```js
Reactable.toggleGroupBy(
  tableId: string,
  columnId: string,
  isGrouped?: boolean
)
```

#### Examples

```js
// Toggle groupBy state for the "Type" column
Reactable.toggleGroupBy('cars-table', 'Type')

// Enable grouping for the "Type" column
Reactable.toggleGroupBy('cars-table', 'Type', true)

// Disable grouping for the "Type" column
Reactable.toggleGroupBy('cars-table', 'Type', false)
```

### `Reactable.setGroupBy()`

Sets the `groupBy` columns for the table. To clear the `groupBy` columns,
set `columnIds` to an empty array, `[]`.

```js
Reactable.setGroupBy(
  tableId: string,
  columnIds: Array<string>
)
```

#### Examples

```js
// Set the groupBy columns to "Type" and "Manufacturer"
Reactable.setGroupBy('cars-table', ['Type', 'Manufacturer'])

// Clear the groupBy columns
Reactable.setGroupBy('cars-table', [])
```

### `Reactable.toggleAllRowsExpanded()`

Toggles the expanded state of all rows in the table between expanded and collapsed.
To expand or collapse rows explicitly, set the optional `isExpanded` argument to
`true` or `false`.

```js
Reactable.toggleAllRowsExpanded(
  tableId: string,
  isExpanded?: boolean
)
```

#### Examples

```js
// Toggle expanded state for all rows
Reactable.toggleAllRowsExpanded('cars-table')

// Expand all rows
Reactable.toggleAllRowsExpanded('cars-table', true)

// Collapse all rows
Reactable.toggleAllRowsExpanded('cars-table', false)
```

### `Reactable.getState()`

Gets the current state of a table.

```js
Reactable.getState(tableId: string)
```

#### Value

An object with the following properties:

```{r, echo=FALSE, asis=TRUE}
stateProps <- dplyr::tribble(
  ~Property, ~Example, ~Description,
  "sorted", '[{ id: "Petal.Length", desc: true }, ...]', "array of columns being sorted in the table",
  "page", "2", "page index (zero-based)",
  "pageSize", "10", "page size",
  "pages", "5", "number of pages",
  "pageRows", '[{ Petal.Length: 1.7, Species: "setosa" }, ...]', "array of row data values in the page",
  "filters", '[{ id: "Species", value: "petal" }]', "array of column filter values",
  "searchValue", '"petal"', "table search value",
  "selected", '[0, 1, 4]', "array of selected row indices (zero-based)",
  "sortedData", '[{ Petal.Length: 1.7, Species: "setosa" }, ...]', "sorted array of row data values in the table",
  "data", '[{ Petal.Length: 1.7, Species: "setosa" }, ...]', "original array of row data values in the table"
)

propsTable(stateProps)
```

#### Examples

```js
Reactable.getState('cars-table')
// { page: 2, searchValue: 'petal', ... }
```

```{css echo=FALSE}
/* rmarkdown html documents */
.main-container {
  font-family: -apple-system, BlinkMacSystemFont, "Segoe UI", Helvetica, Arial, sans-serif;
}

.main-container blockquote {
  font-size: inherit;
}

/* pkgdown articles */
.contents {
  font-family: -apple-system, BlinkMacSystemFont, "Segoe UI", Helvetica, Arial, sans-serif;
}

.contents blockquote {
  font-size: inherit;
}

.reactable {
  font-size: 14px;
}

.props-tbl {
  display: table; /* Override pkgdown styles */
  font-size: 14px;
}

.props-tbl th {
  font-weight: 600;
}

.props-tbl th[scope="row"] {
  font-weight: 400;
}

h3 {
  font-weight: 600;
}
```
