---
title: "Static Rendering"
output:
  html_document:
    toc: true
    toc_float:
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reactable)
```

::: {.callout}
New in v0.3.0.9000
:::

## Introduction

Static rendering is an optional feature that allows you to render tables to static HTML in R.

By default, tables are rendered completely in the user's browser by JavaScript, and
there's no actual HTML being generated when you run `reactable()` in R.

This means when you load a page, tables will be blank at first, and then appear a short time
later as they get rendered. In extreme cases, it can take a noticeably long time for tables to appear,
like with the heavy [Examples without static rendering](./examples-no-static.html) page.

With static rendering, tables are pre-rendered to their initial HTML so they appear
immediately without any flash of content. Tables are then made interactive and subsequently
rendered by JavaScript as needed. You can try refreshing on the
[Examples with static rendering](./examples.html) page for a comparison.

**Advantages**:

- Better user experience in many cases, as tables appear immediately and don't cause
  any layout shift on page load.
- Better support for indexing of table content by search engines. Even crawlers that
  support JavaScript like Googlebot may have a crawl budget that limits crawling of
  JavaScript-heavy pages.
- Non-interactive tables (without pagination, sorting, filtering) no longer require
  JavaScript to view, which may be useful when JavaScript is unavailable, like in
  emails.

Static rendering uses the V8 JavaScript engine provided by the [V8 package](https://github.com/jeroen/V8).
V8 is not installed with reactable by default, so it must be installed before
using static rendering.

::: {.callout-note}
**Note:** Static rendering is currently experimental, and is not supported for tables
rendered via `reactableOutput()` in Shiny. There are also other [limitations and tradeoffs](#limitations-and-tradeoffs)
to consider, and static rendering may not make sense for every table.
:::

## Usage

To use static rendering, ensure that the [V8 package](https://github.com/jeroen/V8)
is installed first. V8 is not installed with reactable by default.

```{r, eval=FALSE}
install.packages("V8")
```

Then, use `reactable(static = TRUE)` to render a single table to static HTML:

```{r}
data <- MASS::Cars93[, c("Manufacturer", "Model", "Type", "Price")]

reactable(data, defaultPageSize = 5, static = TRUE)
```

Or use `options(reactable.static = TRUE)` to enable static rendering for all tables:

```{r}
options(reactable.static = TRUE)

reactable(data, defaultPageSize = 5)
```

```{r, echo=FALSE}
options(reactable.static = FALSE)
```

## Examples

The reactable package documentation uses static rendering in several articles, and
provides alternate versions of these pages without static rendering for comparison:

- [Examples with static rendering](./examples.html) vs. [Examples without static rendering](./examples-no-static.html)
- [Demo Cookbook with static rendering](./cookbook/cookbook.html) vs. [Demo Cookbook without static rendering](./cookbook/cookbook-no-static.html)

## Limitations and Tradeoffs

Although static rendering might seem useful to enable for all tables, there are some
limitations and tradeoffs to consider, and it may not make sense to use for every table.

- Static rendering isn't supported for tables rendered via `reactableOutput()` in Shiny.
  It is still supported for `reactable()` tables specified in a Shiny app's `ui` definition, however.
- Some features in reactable do not fully support static rendering yet:
  - `colFormat()` only supports `en` locales. Formatting with other locales will still
    work in the browser, but tables may initially show data formatted in an `en` locale.
  - Default expanded rows and row details with `reactable(defaultExpanded = TRUE)`
    are not statically rendered.
- There may be a noticeable period where tables are visible but not yet interactable,
  as the client-side JavaScript that makes tables interactive is still running. In these
  cases, the user experience may be better with fully client-side rendered tables.
- HTML documents will be larger, which could affect users with slower download speeds.
  Static rendering is especially not recommended for unpaginated tables with a large number of rows.
- Custom JavaScript render functions and style functions won't have access to the DOM,
  as they won't be run in a browser environment. If your custom JavaScript functions
  require the DOM (e.g., `window` or `document`), you can defer rendering until the DOM
  is available to prevent rendering failures:
  ```{r, eval=FALSE}
  colDef(
    cell = JS("
      function(cellInfo) {
        if (typeof document !== 'undefined') {
          // Not in a browser environment, defer rendering until running in a browser
          return null
        }
      }
    ")
  )
  ```
- Custom JavaScript render functions and style functions won't have access to external
  JavaScript code. Any external JavaScript has to be included inline for now, but this
  may change in the future.
- Tables using `reactableTheme()` may initially appear unstyled when rendered in pkgdown
  websites. Statically rendered theme styles are usually placed in the document `<head>`
  to prevent problem, but pkgdown renders all user-specified `<head>` content into the
  document body, which means tables will appear before browsers process the theme styles.


```{css echo=FALSE}
/* rmarkdown html documents */
.main-container {
  font-family: -apple-system, BlinkMacSystemFont, "Segoe UI", Helvetica, Arial, sans-serif;
}

.main-container blockquote {
  font-size: inherit;
}
```
