---
title: "Sticky Columns (Old Method)"
output:
  html_document:
    toc: true
    toc_float:
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reactable)
```

> Sticky columns are now supported using a `sticky` argument in `colDef()`
> and `colGroup()`, so use that if possible.
> See [Sticky Columns](./examples.html#sticky-columns) for examples.

### Basic sticky columns

You can make a single left or right column sticky using the [`position: sticky`](https://developer.mozilla.org/en-US/docs/Web/CSS/position) CSS property.

Note that `position: sticky` is supported by Chrome, Firefox, Safari, and Edge, but **not** IE11.
IE11 does fail gracefully though -- the columns just don't stick.

`background: #fff` and `z-index: 1` are added to properly hide scrolling content.

```{r}
reactable(
  MASS::Cars93,
  columns = list(
    Manufacturer = colDef(
      style = list(position = "sticky", left = 0, background = "#fff", zIndex = 1),
      headerStyle = list(position = "sticky", left = 0, background = "#fff", zIndex = 1)
    ),
    Make = colDef(
      style = list(position = "sticky", right = 0, background = "#fff", zIndex = 1),
      headerStyle = list(position = "sticky", right = 0, background = "#fff", zIndex = 1)
    )
  ),
  defaultColDef = colDef(minWidth = 150)
)
```

Adding a border style and reducing some of the repetition:

```{r}
sticky_style <- function(left = TRUE) {
  style <- list(position = "sticky", background = "#fff", zIndex = 1)
  if (left) {
    style <- c(style, list(left = 0, borderRight = "1px solid #eee"))
  } else {
    style <- c(style, list(right = 0, borderLeft = "1px solid #eee"))
  }
  style
}

reactable(
  MASS::Cars93,
  columns = list(
    Manufacturer = colDef(
      style = sticky_style(),
      headerStyle = sticky_style()
    ),
    Make = colDef(
      style = sticky_style(left = FALSE),
      headerStyle = sticky_style(left = FALSE)
    )
  ),
  defaultColDef = colDef(minWidth = 150)
)
```

### Multiple sticky columns

To make multiple columns sticky, you have to set the [`left`](https://developer.mozilla.org/en-US/docs/Web/CSS/left)
or `right` positions of the columns to where they'll stick. In this case, you'll
need to know the column widths -- columns are `100px` wide by default, so
the 2nd column from left needs `left: 100px;`, the 3rd column from left needs
`left: 200px;`, etc.

```{r sticky_multi, eval=FALSE}
reactable(
  MASS::Cars93,
  columns = list(
    Manufacturer = colDef(
      class = "sticky left-col-1",
      headerClass = "sticky left-col-1"
    ),
    Model = colDef(
      class = "sticky left-col-2",
      headerClass = "sticky left-col-2"
    ),
    Type = colDef(
      class = "sticky left-col-3",
      headerClass = "sticky left-col-3"
    ),
    Make = colDef(
      class = "sticky right-col-1",
      headerClass = "sticky right-col-1"
    )
  ),
  wrap = FALSE
)
```

```{css}
.sticky {
  position: sticky !important;
  background: #fff;
  z-index: 1;
}

.left-col-1 {
  left: 0;
}

.left-col-2 {
  left: 100px;
}

.left-col-3 {
  left: 200px;
  border-right: 1px solid #eee !important;
}

.right-col-1 {
  right: 0;
  border-left: 1px solid #eee !important;
}
```

```{r ref.label="sticky_multi", echo=FALSE}
```

### Sticky headers and footers

Sticky columns should work fine with sticky headers and footers:

```{r}
reactable(
  MASS::Cars93[1:20, ],
  pagination = FALSE,
  height = 400,
  wrap = FALSE,
  columns = list(
    Manufacturer = colDef(
      class = "sticky left-col-1",
      headerClass = "sticky left-col-1",
      footerClass = "sticky left-col-1"
    ),
    Model = colDef(
      class = "sticky left-col-2",
      headerClass = "sticky left-col-2",
      footerClass = "sticky left-col-2"
    ),
    Type = colDef(
      class = "sticky left-col-3",
      headerClass = "sticky left-col-3",
      footerClass = "sticky left-col-3"
    ),
    Make = colDef(
      class = "sticky right-col-1",
      headerClass = "sticky right-col-1",
      footerClass = "sticky right-col-1"
    )
  ),
  defaultColDef = colDef(footer = "Footer")
)
```

```{r eval=FALSE, echo=FALSE}
# Proof-of-concept of using row highlighting with sticky columns.
# WARNING: This relies on undocumented CSS classes and is not guaranteed
# to work in a future release. Last tested with reactable 0.2.3.
sticky_style <- function(left = TRUE) {
  style <- list(position = "sticky", background = "#fff", zIndex = 1)
  if (left) {
    style <- c(style, list(left = 0, borderRight = "1px solid #eee"))
  } else {
    style <- c(style, list(right = 0, borderLeft = "1px solid #eee"))
  }
  style
}

sticky_theme <- reactableTheme(
  cellStyle = list(
    ".rt-tr-highlight:hover &" = list(backgroundColor = "hsl(0, 0%, 95%) !important")
  )
)

reactable(
  MASS::Cars93,
  columns = list(
    Manufacturer = colDef(
      style = sticky_style(),
      headerStyle = sticky_style()
    ),
    Make = colDef(
      style = sticky_style(left = FALSE),
      headerStyle = sticky_style(left = FALSE)
    )
  ),
  defaultColDef = colDef(minWidth = 150),
  highlight = TRUE,
  theme = sticky_theme
)
```

```{css echo=FALSE}
/* rmarkdown html documents */
.main-container {
  font-family: -apple-system, BlinkMacSystemFont, "Segoe UI", Helvetica, Arial, sans-serif;
}

.main-container blockquote {
  font-size: inherit;
}

/* pkgdown articles */
.contents {
  font-family: -apple-system, BlinkMacSystemFont, "Segoe UI", Helvetica, Arial, sans-serif;
}

.contents blockquote {
  font-size: inherit;
}

.reactable {
  font-size: 14px;
}
```
