---
title: "DuckDB-WASM Engine — Basic Test"
output:
  html_document:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reactable)
```

## Basic pagination with DuckDB backend

This test renders a table with the DuckDB backend, which serializes data as Arrow IPC and
queries it client-side via DuckDB-WASM for pagination.

### MASS::Cars93 (93 rows)

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  showPageSizeOptions = TRUE,
  pageSizeOptions = c(10, 25, 50, 100)
)
```

### mtcars (32 rows)

```{r}
reactable(
  mtcars,
  backend = backendDuckDB(),
  defaultPageSize = 10
)
```

### iris (150 rows)

```{r}
reactable(
  iris,
  backend = backendDuckDB(),
  defaultPageSize = 25
)
```

### Small dataset (less than one page)

```{r}
reactable(
  head(mtcars, 5),
  backend = backendDuckDB(),
  defaultPageSize = 10
)
```

### Pre-rendered first page with defaultSorted

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  defaultSorted = list(Horsepower = "desc")
)
```

### Large dataset: 100k rows

```{r}
n <- 100000
large_data <- data.frame(
  id = seq_len(n),
  value = rnorm(n),
  category = sample(LETTERS, n, replace = TRUE),
  score = round(runif(n, 0, 100), 2)
)

reactable(
  large_data,
  backend = backendDuckDB(),
  defaultPageSize = 20,
  showPageSizeOptions = TRUE,
  pageSizeOptions = c(20, 50, 100)
)
```

### Large dataset: 1M rows

```{r}
n <- 1000000
big_data <- data.frame(
  id = seq_len(n),
  value = rnorm(n),
  category = sample(LETTERS, n, replace = TRUE),
  score = round(runif(n, 0, 100), 2)
)

reactable(
  big_data,
  backend = backendDuckDB(),
  defaultPageSize = 20,
  showPageSizeOptions = TRUE,
  searchable = TRUE
)
```

### Large dataset: 1M rows (no DuckDB)

```{r}
n <- 1000000
big_data <- data.frame(
  id = seq_len(n),
  value = rnorm(n),
  category = sample(LETTERS, n, replace = TRUE),
  score = round(runif(n, 0, 100), 2)
)

reactable(
  big_data,
  defaultPageSize = 20,
  showPageSizeOptions = TRUE,
  searchable = TRUE
)
```

### Large dataset: 1M rows (virtualized)

```{r}
n <- 1000000
big_data <- data.frame(
  id = seq_len(n),
  value = rnorm(n),
  category = sample(LETTERS, n, replace = TRUE),
  score = round(runif(n, 0, 100), 2)
)

reactable(
  big_data,
  backend = backendDuckDB(),
  pagination = FALSE,
  virtual = TRUE,
  height = 500,
  sortable = TRUE,
  filterable = TRUE,
  searchable = TRUE
)
```

### Unpaginated with virtual scrolling (MASS::Cars93)

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  pagination = FALSE,
  virtual = TRUE,
  height = 400,
  sortable = TRUE,
  filterable = TRUE,
  searchable = TRUE,
  striped = TRUE
)
```

### Comparison: same data without backend (normal mode)

```{r}
reactable(
  MASS::Cars93,
  defaultPageSize = 10
)
```

## Sorting

### Sortable columns

All columns should be sortable. Click column headers to sort. Sorting should happen server-side
via DuckDB and show correct results across pages.

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  sortable = TRUE
)
```

### Default sorted with sortable

Start sorted by Horsepower descending. Clicking other headers should re-sort via DuckDB.

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  sortable = TRUE,
  defaultSorted = list(Horsepower = "desc")
)
```

## Column Filters

### Filterable columns

All columns should have filter inputs. Typing in a filter should query DuckDB and update results.
Numeric columns should use "starts with" matching, text columns should use substring matching.

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  filterable = TRUE
)
```

### Filterable with sorting

Combining filters and sorting. Filter first, then sort the filtered results.

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  filterable = TRUE,
  sortable = TRUE
)
```

## Global Search

### Searchable table

Global search should search across all columns via DuckDB. Page count should update as results
are filtered.

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  searchable = TRUE
)
```

### Search with sort and filter combined

All three features working together: search, column filters, and sorting.

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  searchable = TRUE,
  filterable = TRUE,
  sortable = TRUE
)
```

## Large Dataset with Sort/Filter/Search

### 100k rows with all features

```{r}
n <- 100000
large_data <- data.frame(
  id = seq_len(n),
  value = rnorm(n),
  category = sample(LETTERS, n, replace = TRUE),
  score = round(runif(n, 0, 100), 2)
)

reactable(
  large_data,
  backend = backendDuckDB(),
  defaultPageSize = 20,
  searchable = TRUE,
  filterable = TRUE,
  sortable = TRUE,
  showPageSizeOptions = TRUE,
  pageSizeOptions = c(20, 50, 100)
)
```

## DuckDB R Server Backend (Shiny only)

The following examples use `backend = backendDuckDB()` which automatically uses server mode
in a Shiny session. They won't render in static HTML -- run as a Shiny app with
`shiny::runApp()` or use the example app at `inst/examples/shiny-server-data-duckdb.R`.

### Full Shiny app with DuckDB server backend

```{r eval=FALSE}
library(shiny)
library(reactable)

rows <- 100000
dates <- seq.Date(as.Date("2018-01-01"), as.Date("2018-12-01"), "day")
data <- data.frame(
  index = seq_len(rows),
  date = sample(dates, rows, replace = TRUE),
  city = sample(names(precip), rows, replace = TRUE),
  state = sample(rownames(USArrests), rows, replace = TRUE),
  temp = round(runif(rows, 0, 100), 1),
  stringsAsFactors = FALSE
)

ui <- fluidPage(
  titlePanel("DuckDB Server Backend — 100k rows"),
  reactableOutput("tbl")
)

server <- function(input, output) {
  output$tbl <- renderReactable({
    reactable(
      data,
      backend = backendDuckDB(),
      defaultPageSize = 20,
      filterable = TRUE,
      searchable = TRUE,
      sortable = TRUE,
      highlight = TRUE,
      showPageSizeOptions = TRUE,
      pageSizeOptions = c(10, 20, 50, 100),
      columns = list(
        index = colDef(name = "ID", width = 80),
        date = colDef(name = "Date"),
        city = colDef(name = "City"),
        state = colDef(
          name = "State",
          html = TRUE,
          cell = JS("function(cell) {
            return '<a href=\"https://wikipedia.org/wiki/' + cell.value + '\">' + cell.value + '</a>'
          }")
        ),
        temp = colDef(
          name = "Temperature",
          format = colFormat(digits = 1)
        )
      )
    )
  })
}

shinyApp(ui, server)
```

### Full Shiny app with 1M rows

```{r eval=FALSE}
library(shiny)
library(reactable)

rows <- 1000000
dates <- seq.Date(as.Date("2018-01-01"), as.Date("2023-12-31"), "day")
data <- data.frame(
  index = seq_len(rows),
  date = sample(dates, rows, replace = TRUE),
  city = sample(names(precip), rows, replace = TRUE),
  state = sample(rownames(USArrests), rows, replace = TRUE),
  temp = round(runif(rows, 0, 100), 1),
  amount = round(runif(rows, 1, 10000), 2),
  stringsAsFactors = FALSE
)

ui <- fluidPage(
  titlePanel("DuckDB Server Backend — 1M rows"),
  reactableOutput("tbl")
)

server <- function(input, output) {
  output$tbl <- renderReactable({
    reactable(
      data,
      backend = backendDuckDB(),
      defaultPageSize = 25,
      filterable = TRUE,
      searchable = TRUE,
      sortable = TRUE,
      highlight = TRUE,
      showPageSizeOptions = TRUE,
      pageSizeOptions = c(25, 50, 100, 500),
      columns = list(
        index = colDef(name = "ID", width = 80),
        date = colDef(name = "Date"),
        city = colDef(name = "City"),
        state = colDef(name = "State"),
        temp = colDef(name = "Temperature", format = colFormat(digits = 1)),
        amount = colDef(name = "Amount", format = colFormat(currency = "USD", separators = TRUE))
      )
    )
  })
}

shinyApp(ui, server)
```

### Comparison: DuckDB server vs V8 server vs df server

```{r eval=FALSE}
# All three backends produce the same results for basic sort/filter/search.
# DuckDB is ~100x faster than df for large datasets.

# DuckDB backend (fastest for large data)
reactable(MASS::Cars93, backend = backendDuckDB(), filterable = TRUE, searchable = TRUE)

# V8 backend (default, matches client-side behavior exactly)
reactable(MASS::Cars93, backend = backendV8(), filterable = TRUE, searchable = TRUE)

# Data frame backend (simplest, no external dependencies)
reactable(MASS::Cars93, backend = backendDf(), filterable = TRUE, searchable = TRUE)
```

## Grouping and Aggregation

### Basic groupBy with sum aggregate

Group by Manufacturer, show the sum of Min.Price and Max.Price per group. Clicking a group row
should expand to show the individual car rows within that group.

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  groupBy = "Manufacturer",
  columns = list(
    Min.Price = colDef(aggregate = "sum"),
    Max.Price = colDef(aggregate = "sum"),
    Horsepower = colDef(aggregate = "mean")
  )
)
```

### Different aggregate functions

Demonstrating all SQL-backed aggregate functions on a single grouped table.

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  groupBy = "Type",
  columns = list(
    Min.Price = colDef(aggregate = "sum", format = colFormat(digits = 1)),
    Max.Price = colDef(aggregate = "max", format = colFormat(digits = 1)),
    Horsepower = colDef(aggregate = "mean", format = colFormat(digits = 0)),
    MPG.city = colDef(aggregate = "median", format = colFormat(digits = 0)),
    RPM = colDef(aggregate = "min"),
    Cylinders = colDef(aggregate = "count"),
    Manufacturer = colDef(aggregate = "unique"),
    Origin = colDef(aggregate = "frequency")
  )
)
```

### Grouped with sort and filter

Combine grouping with sorting and filtering. Filter should narrow down rows before grouping,
and sort should sort the group rows by the group column or aggregated values.

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  groupBy = "Origin",
  filterable = TRUE,
  sortable = TRUE,
  columns = list(
    Min.Price = colDef(aggregate = "sum", format = colFormat(digits = 1)),
    Max.Price = colDef(aggregate = "sum", format = colFormat(digits = 1)),
    Horsepower = colDef(aggregate = "mean", format = colFormat(digits = 0)),
    Manufacturer = colDef(aggregate = "unique")
  )
)
```

### Grouped with search

Global search should filter rows before grouping.

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  groupBy = "Type",
  searchable = TRUE,
  columns = list(
    Min.Price = colDef(aggregate = "sum"),
    Horsepower = colDef(aggregate = "mean"),
    Manufacturer = colDef(aggregate = "count")
  )
)
```

### Multi-level grouping

Group by Origin first, then by Type within each origin. Two levels of expand/collapse.

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  groupBy = c("Origin", "Type"),
  columns = list(
    Min.Price = colDef(aggregate = "sum", format = colFormat(digits = 1)),
    Max.Price = colDef(aggregate = "sum", format = colFormat(digits = 1)),
    Horsepower = colDef(aggregate = "mean", format = colFormat(digits = 0)),
    Manufacturer = colDef(aggregate = "unique")
  )
)
```

### Large grouped dataset: 100k rows grouped by category

```{r}
n <- 100000
large_data <- data.frame(
  id = seq_len(n),
  category = sample(LETTERS, n, replace = TRUE),
  region = sample(c("East", "West", "North", "South"), n, replace = TRUE),
  value = round(rnorm(n, mean = 100, sd = 50), 2),
  score = round(runif(n, 0, 100), 2)
)

reactable(
  large_data,
  backend = backendDuckDB(),
  defaultPageSize = 20,
  groupBy = "category",
  sortable = TRUE,
  columns = list(
    value = colDef(aggregate = "sum", format = colFormat(digits = 0)),
    score = colDef(aggregate = "mean", format = colFormat(digits = 1)),
    id = colDef(aggregate = "count"),
    region = colDef(aggregate = "unique")
  )
)
```

### Large multi-level grouping: 100k rows grouped by category and region

```{r}
n <- 100000
large_data <- data.frame(
  id = seq_len(n),
  category = sample(LETTERS, n, replace = TRUE),
  region = sample(c("East", "West", "North", "South"), n, replace = TRUE),
  value = round(rnorm(n, mean = 100, sd = 50), 2),
  score = round(runif(n, 0, 100), 2)
)

reactable(
  large_data,
  backend = backendDuckDB(),
  defaultPageSize = 20,
  groupBy = c("category", "region"),
  sortable = TRUE,
  filterable = TRUE,
  searchable = TRUE,
  columns = list(
    value = colDef(aggregate = "sum", format = colFormat(digits = 0)),
    score = colDef(aggregate = "mean", format = colFormat(digits = 1)),
    id = colDef(aggregate = "count")
  )
)
```

### paginateSubRows with DuckDB backend

With `paginateSubRows = TRUE`, expanded sub-rows count toward the page size. This prevents a
group with many rows from overwhelming the page. Compare with the non-DuckDB version below.

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  groupBy = "Manufacturer",
  paginateSubRows = TRUE,
  columns = list(
    Min.Price = colDef(aggregate = "sum"),
    Max.Price = colDef(aggregate = "sum"),
    Horsepower = colDef(aggregate = "mean")
  )
)
```

### paginateSubRows with multi-level grouping

Multi-level grouping with paginateSubRows. Expanding groups at both levels should count
toward the page size.

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  groupBy = c("Origin", "Type"),
  paginateSubRows = TRUE,
  columns = list(
    Min.Price = colDef(aggregate = "sum", format = colFormat(digits = 1)),
    Max.Price = colDef(aggregate = "sum", format = colFormat(digits = 1)),
    Horsepower = colDef(aggregate = "mean", format = colFormat(digits = 0)),
    Manufacturer = colDef(aggregate = "unique")
  )
)
```

### paginateSubRows with large dataset

100k rows with 4 groups of ~25k rows each. Without paginateSubRows, expanding a group
would show all 25k sub-rows. With paginateSubRows, only pageSize rows are shown at a time.

```{r}
n <- 100000
large_grouped <- data.frame(
  region = sample(c("East", "West", "North", "South"), n, replace = TRUE),
  category = sample(LETTERS[1:10], n, replace = TRUE),
  value = round(rnorm(n, mean = 100, sd = 50), 2),
  score = round(runif(n, 0, 100), 2),
  stringsAsFactors = FALSE
)

reactable(
  large_grouped,
  backend = backendDuckDB(),
  defaultPageSize = 20,
  groupBy = "region",
  paginateSubRows = TRUE,
  sortable = TRUE,
  filterable = TRUE,
  columns = list(
    value = colDef(aggregate = "sum", format = colFormat(digits = 0)),
    score = colDef(aggregate = "mean", format = colFormat(digits = 1)),
    category = colDef(aggregate = "count")
  )
)
```

### paginateSubRows with Parquet sidecar

100k rows forced to Parquet format. Tests that paginateSubRows works with the Parquet
code path (HTTP range requests via DuckDB-WASM) rather than embedded Arrow IPC.

```{r}
n <- 100000
parquet_grouped <- data.frame(
  region = sample(c("East", "West", "North", "South"), n, replace = TRUE),
  category = sample(LETTERS[1:10], n, replace = TRUE),
  value = round(rnorm(n, mean = 100, sd = 50), 2),
  score = round(runif(n, 0, 100), 2),
  stringsAsFactors = FALSE
)

reactable(
  parquet_grouped,
  backend = backendDuckDB(format = "parquet"),
  defaultPageSize = 20,
  groupBy = "region",
  paginateSubRows = TRUE,
  sortable = TRUE,
  filterable = TRUE,
  columns = list(
    value = colDef(aggregate = "sum", format = colFormat(digits = 0)),
    score = colDef(aggregate = "mean", format = colFormat(digits = 1)),
    category = colDef(aggregate = "count")
  )
)
```

### paginateSubRows with Parquet sidecar and multi-level grouping

100k rows with Parquet format and two-level grouping. Expanding groups at both levels
should count toward the page size.

```{r}
n <- 100000
parquet_multi <- data.frame(
  region = sample(c("East", "West", "North", "South"), n, replace = TRUE),
  category = sample(LETTERS[1:10], n, replace = TRUE),
  value = round(rnorm(n, mean = 100, sd = 50), 2),
  score = round(runif(n, 0, 100), 2),
  stringsAsFactors = FALSE
)

reactable(
  parquet_multi,
  backend = backendDuckDB(format = "parquet"),
  defaultPageSize = 20,
  groupBy = c("region", "category"),
  paginateSubRows = TRUE,
  sortable = TRUE,
  filterable = TRUE,
  columns = list(
    value = colDef(aggregate = "sum", format = colFormat(digits = 0)),
    score = colDef(aggregate = "mean", format = colFormat(digits = 1))
  )
)
```

### Comparison: paginateSubRows without backend (normal client-side mode)

```{r}
reactable(
  MASS::Cars93,
  defaultPageSize = 10,
  groupBy = "Manufacturer",
  paginateSubRows = TRUE,
  columns = list(
    Min.Price = colDef(aggregate = "sum"),
    Max.Price = colDef(aggregate = "sum"),
    Horsepower = colDef(aggregate = "mean")
  )
)
```

### Comparison: groupBy without backend (normal client-side mode)

Same table without the DuckDB backend for behavior comparison.

```{r}
reactable(
  MASS::Cars93,
  defaultPageSize = 10,
  groupBy = "Manufacturer",
  columns = list(
    Min.Price = colDef(aggregate = "sum"),
    Max.Price = colDef(aggregate = "sum"),
    Horsepower = colDef(aggregate = "mean")
  )
)
```

### DuckDB server backend with grouping (Shiny only)

```{r eval=FALSE}
library(shiny)
library(reactable)

rows <- 100000
data <- data.frame(
  region = sample(c("East", "West", "North", "South"), rows, replace = TRUE),
  state = sample(rownames(USArrests), rows, replace = TRUE),
  category = sample(LETTERS[1:10], rows, replace = TRUE),
  revenue = round(runif(rows, 100, 50000), 2),
  quantity = sample(1:100, rows, replace = TRUE),
  stringsAsFactors = FALSE
)

ui <- fluidPage(
  titlePanel("DuckDB Server Backend — Grouped 100k rows"),
  reactableOutput("tbl")
)

server <- function(input, output) {
  output$tbl <- renderReactable({
    reactable(
      data,
      backend = backendDuckDB(),
      defaultPageSize = 20,
      groupBy = "region",
      filterable = TRUE,
      searchable = TRUE,
      sortable = TRUE,
      highlight = TRUE,
      columns = list(
        revenue = colDef(
          name = "Revenue",
          aggregate = "sum",
          format = colFormat(currency = "USD", separators = TRUE, digits = 0)
        ),
        quantity = colDef(name = "Quantity", aggregate = "sum"),
        state = colDef(name = "State", aggregate = "unique"),
        category = colDef(name = "Category", aggregate = "count")
      )
    )
  })
}

shinyApp(ui, server)
```

### DuckDB server backend with paginateSubRows (Shiny only)

```{r eval=FALSE}
library(shiny)
library(reactable)

data <- MASS::Cars93[, c("Manufacturer", "Model", "Type", "Min.Price", "Max.Price")]

ui <- fluidPage(
  h3("DuckDB Server-Side paginateSubRows Test"),
  checkboxInput("paginateSubRows", "paginateSubRows", value = TRUE),
  checkboxInput("multiGroup", "Multi-level groupBy (Manufacturer + Type)", value = FALSE),
  htmltools::tags$button("Expand/collapse all", onclick = "Reactable.toggleAllRowsExpanded('tbl')"),
  reactableOutput("tbl")
)

server <- function(input, output) {
  output$tbl <- renderReactable({
    groupBy <- if (input$multiGroup) c("Manufacturer", "Type") else "Manufacturer"
    reactable(
      data,
      elementId = "tbl",
      groupBy = groupBy,
      paginateSubRows = input$paginateSubRows,
      columns = list(
        Min.Price = colDef(aggregate = "sum"),
        Max.Price = colDef(aggregate = "mean")
      ),
      filterable = TRUE,
      searchable = TRUE,
      showPageSizeOptions = TRUE,
      defaultPageSize = 10,
      backend = backendDuckDB()
    )
  })
}

shinyApp(ui, server)
```

## Column Formatting

### colFormat with numbers, currencies, and percentages

`colFormat()` is applied client-side via JavaScript's `Intl.NumberFormat`, so it works correctly
with DuckDB query results. The data types (numbers, strings) are preserved through Arrow IPC.

```{r}
sales_data <- data.frame(
  product = c("Widget A", "Widget B", "Gadget X", "Gadget Y", "Gizmo Z"),
  units_sold = c(15234, 8921, 42010, 3150, 99120),
  revenue = c(152340.50, 89210.75, 420100.00, 31500.25, 991200.99),
  margin = c(0.234, 0.157, 0.412, 0.089, 0.301),
  avg_price = c(9.999, 10.0, 10.001, 10.0001, 10.00001),
  stringsAsFactors = FALSE
)

reactable(
  sales_data,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  columns = list(
    units_sold = colDef(
      name = "Units Sold",
      format = colFormat(separators = TRUE)
    ),
    revenue = colDef(
      name = "Revenue",
      format = colFormat(currency = "USD", separators = TRUE, digits = 2)
    ),
    margin = colDef(
      name = "Margin",
      format = colFormat(percent = TRUE, digits = 1)
    ),
    avg_price = colDef(
      name = "Avg Price",
      format = colFormat(digits = 2, prefix = "$")
    )
  )
)
```

### colFormat with dates

Date formatting via `colFormat()` works on DuckDB query results.

```{r}
date_data <- data.frame(
  event = c("Launch", "Update 1", "Update 2", "Sunset"),
  date = as.Date(c("2024-01-15", "2024-06-01", "2024-09-30", "2025-03-01")),
  datetime = as.POSIXct(c(
    "2024-01-15 09:30:00", "2024-06-01 14:00:00",
    "2024-09-30 23:59:59", "2025-03-01 00:00:01"
  )),
  stringsAsFactors = FALSE
)

reactable(
  date_data,
  backend = backendDuckDB(),
  columns = list(
    date = colDef(format = colFormat(date = TRUE)),
    datetime = colDef(format = colFormat(datetime = TRUE))
  )
)
```

### colFormat on grouped aggregates

`colFormat()` also applies to aggregated values in grouped tables.

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  groupBy = "Type",
  columns = list(
    Min.Price = colDef(
      aggregate = "sum",
      format = colFormat(currency = "USD", separators = TRUE, digits = 0)
    ),
    Max.Price = colDef(
      aggregate = "mean",
      format = colFormat(currency = "USD", separators = TRUE, digits = 2)
    ),
    Horsepower = colDef(
      aggregate = "mean",
      format = colFormat(digits = 0, suffix = " hp")
    ),
    MPG.city = colDef(
      aggregate = "median",
      format = colFormat(digits = 1, suffix = " mpg")
    ),
    Weight = colDef(
      aggregate = "sum",
      format = colFormat(separators = TRUE, suffix = " lbs")
    )
  )
)
```

## JS Render Functions

### JS cell renderer

`colDef(cell = JS(...))` works correctly with DuckDB — the JS function receives each cell value
as DuckDB returns it.

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  filterable = TRUE,
  columns = list(
    Horsepower = colDef(
      cell = JS("function(cellInfo) {
        const value = cellInfo.value
        const color = value >= 200 ? '#e00' : value >= 150 ? '#f80' : '#080'
        return React.createElement('span', {
          style: { color: color, fontWeight: 'bold' }
        }, value)
      }")
    ),
    Min.Price = colDef(
      format = colFormat(currency = "USD", separators = TRUE, digits = 2)
    )
  )
)
```

### JS style and class functions

`colDef(style = JS(...))` and `colDef(class = JS(...))` work correctly with DuckDB.

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  sortable = TRUE,
  columns = list(
    Horsepower = colDef(
      style = JS("function(rowInfo) {
        const value = rowInfo.value
        if (value >= 200) return { color: '#e00', fontWeight: 'bold' }
        if (value >= 150) return { color: '#f80' }
        return {}
      }")
    ),
    Min.Price = colDef(
      style = JS("function(rowInfo) {
        if (rowInfo.value > 30) return { backgroundColor: '#ffe0e0' }
      }")
    )
  )
)
```

### JS row details

`colDef(details = JS(...))` works correctly with DuckDB.

```{r}
reactable(
  MASS::Cars93[, c("Manufacturer", "Model", "Type", "Min.Price", "Max.Price", "Horsepower")],
  backend = backendDuckDB(),
  defaultPageSize = 10,
  details = JS("function(rowInfo) {
    return React.createElement('div', { style: { padding: '10px', background: '#f8f8f8' } },
      React.createElement('b', null, rowInfo.values.Manufacturer + ' ' + rowInfo.values.Model),
      React.createElement('br'),
      'Type: ' + rowInfo.values.Type,
      React.createElement('br'),
      'Price range: $' + rowInfo.values['Min.Price'] + ' - $' + rowInfo.values['Max.Price']
    )
  }")
)
```

## Theming

### Themed DuckDB table

Themes work normally with the DuckDB backend since they're applied client-side.

```{r}
reactable(
  MASS::Cars93[, c("Manufacturer", "Model", "Type", "Min.Price", "Max.Price", "Horsepower")],
  backend = backendDuckDB(),
  defaultPageSize = 10,
  searchable = TRUE,
  filterable = TRUE,
  highlight = TRUE,
  bordered = TRUE,
  striped = TRUE,
  theme = reactableTheme(
    borderColor = "#dfe2e5",
    stripedColor = "#f6f8fa",
    highlightColor = "#f0f5ff",
    headerStyle = list(
      borderColor = "#dfe2e5",
      backgroundColor = "#f6f8fa"
    ),
    searchInputStyle = list(
      paddingLeft = "1.9rem",
      paddingTop = "0.5rem",
      paddingBottom = "0.5rem"
    )
  )
)
```

## Parquet Sidecar Files

### Explicit Parquet format (small data)

Force Parquet sidecar even for small data. A `.parquet` file should appear in the `_files/lib/` directory.

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(format = "parquet"),
  defaultPageSize = 10,
  sortable = TRUE,
  filterable = TRUE,
  searchable = TRUE
)
```

### Auto Parquet (large data, > 20 MB)

1M rows should auto-switch to Parquet sidecar file.

```{r}
n <- 1000000
parquet_data <- data.frame(
  id = seq_len(n),
  value = rnorm(n),
  category = sample(LETTERS, n, replace = TRUE),
  score = round(runif(n, 0, 100), 2),
  label = paste0("item-", seq_len(n))
)

reactable(
  parquet_data,
  backend = backendDuckDB(),
  defaultPageSize = 20,
  sortable = TRUE,
  filterable = TRUE,
  searchable = TRUE,
  showPageSizeOptions = TRUE
)
```

### Explicit Arrow format (force embedded)

Force embedded Arrow IPC even for large data.

```{r}
n <- 1000000
parquet_data <- data.frame(
  id = seq_len(n),
  value = rnorm(n),
  category = sample(LETTERS, n, replace = TRUE),
  score = round(runif(n, 0, 100), 2),
  label = paste0("item-", seq_len(n))
)

reactable(
  parquet_data,
  backend = backendDuckDB(format = "arrow"),
  defaultPageSize = 20,
  sortable = TRUE,
  searchable = TRUE
)
```

### Parquet with virtual scrolling

Unpaginated virtual scrolling with Parquet sidecar. DuckDB sends all rows in a single
query (pageSize = null), and virtual scrolling handles rendering.

```{r}
n <- 1000000
parquet_data <- data.frame(
  id = seq_len(n),
  value = rnorm(n),
  category = sample(LETTERS, n, replace = TRUE),
  score = round(runif(n, 0, 100), 2),
  label = paste0("item-", seq_len(n))
)

reactable(
  parquet_data,
  backend = backendDuckDB(format = "parquet"),
  pagination = FALSE,
  virtual = TRUE,
  height = 500,
  sortable = TRUE,
  filterable = TRUE,
  searchable = TRUE
)
```

### Parquet in Shiny (client mode)

Test that Parquet sidecar files work in a Shiny app. In Shiny, `backendDuckDB()` defaults
to server mode (data stays in R), so Parquet requires `mode = "client"` to force DuckDB-WASM.
The Parquet file is served via Shiny's htmlDependency resource path, and the locator script
resolves the URL using `document.currentScript.src`.

Test steps:
1. Table should render and be interactive (sort, filter, search).
2. Open browser DevTools Network tab -- requests to the `.parquet` file should use HTTP
   range requests (partial content, status 206 or 200 with Range header).
3. Sorting, filtering should work without full file download.

```{r eval=FALSE}
library(shiny)
library(reactable)

n <- 1000000
parquet_data <- data.frame(
  id = seq_len(n),
  value = rnorm(n),
  category = sample(LETTERS, n, replace = TRUE),
  score = round(runif(n, 0, 100), 2),
  label = paste0("item-", seq_len(n))
)

ui <- fluidPage(
  h3("Parquet sidecar in Shiny (DuckDB client mode, 1M rows)"),
  reactableOutput("table")
)

server <- function(input, output, session) {
  output$table <- renderReactable({
    reactable(
      parquet_data,
      backend = backendDuckDB(mode = "client", format = "parquet"),
      defaultPageSize = 20,
      sortable = TRUE,
      filterable = TRUE,
      searchable = TRUE,
      showPageSizeOptions = TRUE
    )
  })
}

shinyApp(ui, server)
```

### Multiple Parquet tables in Shiny

Test that multiple Parquet-backed tables on the same Shiny page resolve their sidecar
files independently. Each table has its own `parquetId` and locator script, so the
querySelector fallback must use a table-specific selector (not just `parquet-locator`).

Test steps:
1. Both tables should render with correct data.
2. Sorting and filtering should work independently on each table.
3. Check browser console for errors (wrong Parquet URL would cause a fetch failure).

```{r eval=FALSE}
library(shiny)
library(reactable)

ui <- fluidPage(
  h3("Multiple Parquet tables in Shiny"),
  h4("Table 1: mtcars"),
  reactableOutput("table1"),
  h4("Table 2: iris"),
  reactableOutput("table2")
)

server <- function(input, output, session) {
  output$table1 <- renderReactable({
    reactable(
      mtcars,
      backend = backendDuckDB(mode = "client", format = "parquet"),
      defaultPageSize = 10,
      sortable = TRUE,
      filterable = TRUE
    )
  })

  output$table2 <- renderReactable({
    reactable(
      iris,
      backend = backendDuckDB(mode = "client", format = "parquet"),
      defaultPageSize = 10,
      sortable = TRUE,
      filterable = TRUE
    )
  })
}

shinyApp(ui, server)
```

### Arrow IPC in Shiny (client mode)

Test that embedded Arrow IPC works in a Shiny app with DuckDB client mode.
This should work the same as the Parquet example but with data embedded in the HTML.

```{r eval=FALSE}
library(shiny)
library(reactable)

ui <- fluidPage(
  h3("Arrow IPC in Shiny (DuckDB client mode)"),
  reactableOutput("table")
)

server <- function(input, output, session) {
  output$table <- renderReactable({
    reactable(
      MASS::Cars93,
      backend = backendDuckDB(mode = "client", format = "arrow"),
      defaultPageSize = 10,
      sortable = TRUE,
      filterable = TRUE,
      searchable = TRUE
    )
  })
}

shinyApp(ui, server)
```

## Cross-page selection

### Select-all across pages

Select all, then navigate to page 2 -- all rows should still be selected.
Deselect a row on page 2, navigate back to page 1 -- all page 1 rows should
still be selected. Navigate back to page 2 -- the deselected row should still
be unchecked.

With search: type a search term (e.g., "Ford"), select all filtered rows,
clear the search -- only the previously filtered rows should remain selected.
The select-all checkbox should not be checked (only a subset is selected).

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  selection = "multiple",
  searchable = TRUE
)
```

### Select-all + search filter reset

1. Click select-all checkbox (all rows become selected)
2. Type in the search box (e.g., "Ford")
3. Selection should be cleared -- the select-all should not carry over to the
   filtered result set

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  selection = "multiple",
  searchable = TRUE
)
```

### Select-all + column filter reset

1. Click select-all checkbox
2. Type a value in a column filter (e.g., filter Manufacturer to "Ford")
3. Selection should be cleared

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  selection = "multiple",
  filterable = TRUE
)
```

### Select-all + both search and column filter

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  selection = "multiple",
  searchable = TRUE,
  filterable = TRUE
)
```

### Per-row selection across pages

1. Select rows 1 and 3 on page 1
2. Navigate to page 2 -- no rows should be selected
3. Select row 2 on page 2
4. Navigate back to page 1 -- rows 1 and 3 should still be selected
5. Search for "Ford" -- selected Ford rows should still be checked
6. Clear search -- rows 1 and 3 (and row 2 from page 2) should still be selected

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  selection = "multiple",
  searchable = TRUE
)
```

### Default selected rows

Rows 1, 3, and 5 should be pre-selected on load. Navigate to page 2 -- no rows
should be checked. Navigate back to page 1 -- rows 1, 3, and 5 should still be
selected. The select-all checkbox should not be checked (only 3 of 93 are selected).

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  selection = "multiple",
  defaultSelected = c(1, 3, 5),
  searchable = TRUE
)
```

### Default selected with single selection

Row 2 should be pre-selected on load. Clicking another row should deselect row 2.

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  selection = "single",
  defaultSelected = 2
)
```

### Default selected with sorted data

Rows 1 and 2 (by original data order) should be pre-selected. With Horsepower
sorted descending, the selected rows may not appear on the first page. Navigate
to find them -- they should be checked.

```{r}
reactable(
  MASS::Cars93,
  backend = backendDuckDB(),
  defaultPageSize = 10,
  selection = "multiple",
  defaultSelected = c(1, 2),
  defaultSorted = list(Horsepower = "desc"),
  searchable = TRUE
)
```

## Cross-page selection in Shiny

Run this Shiny app to test cross-page selection with `getReactableState("selected")` and
`Reactable.getState().selected`.

Test steps:
1. Select all -- should select all 93 rows. `getReactableState` should show `1:93`.
2. Search for "Ford" -- selected Ford rows should still be checked.
   `getReactableState` should still show all 93 indices.
3. Clear search, filter Manufacturer to "Ford" -- only Ford rows are checked.
4. Click select-all on filtered view -- should select only the filtered rows.
   `getReactableState` should show non-sequential indices (e.g., `14, 15, 16, 17, 18`).
5. Clear filter -- only those Ford rows should remain selected, others unchecked.
6. Sort by a column -- selections should persist.
7. Select rows 1 and 3 on page 1. Navigate to page 2. Click "Check JS getState".
   Both R and JS outputs should show the same selected row indices (R is 1-based, JS is 0-based).
   Before the `selectedRowIndexes` fix, JS would show empty on a different page.

```r
library(shiny)
library(reactable)

ui <- fluidPage(
  h3("Cross-page selection test (DuckDB backend)"),
  actionButton("select_rows", "Select rows 1, 3, 5"),
  actionButton("clear_selection", "Clear selection"),
  actionButton("select_all", "Select all rows"),
  actionButton("check_js", "Check JS getState"),
  reactableOutput("table"),
  h4("R: getReactableState('selected')"),
  verbatimTextOutput("selected_output"),
  h4("JS: Reactable.getState().selected"),
  verbatimTextOutput("js_selected")
)

server <- function(input, output, session) {
  output$table <- renderReactable({
    reactable(
      MASS::Cars93[, 1:8],
      backend = backendDuckDB(),
      selection = "multiple",
      defaultPageSize = 10,
      searchable = TRUE,
      filterable = TRUE,
      elementId = "table"
    )
  })

  observeEvent(input$select_rows, {
    updateReactable("table", selected = c(1, 3, 5))
  })
  observeEvent(input$clear_selection, {
    updateReactable("table", selected = NA)
  })
  observeEvent(input$select_all, {
    updateReactable("table", selected = seq_len(nrow(MASS::Cars93)))
  })

  output$selected_output <- renderPrint({
    selected <- getReactableState("table", "selected")
    cat("Count:", length(selected), "\n")
    cat("Selected:", paste(selected, collapse = ", "), "\n")
  })

  observeEvent(input$check_js, {
    session$sendCustomMessage("check_js_state", list())
  })
  output$js_selected <- renderPrint({
    sel <- input$js_state_selected
    if (is.null(sel) || length(sel) == 0) {
      cat("None")
    } else {
      cat("Count:", length(sel), "\n")
      cat("Selected (0-based):", paste(sel, collapse = ", "), "\n")
    }
  })
}

ui <- tagList(
  ui,
  tags$script(HTML("
    Shiny.addCustomMessageHandler('check_js_state', function(msg) {
      var state = Reactable.getState('table');
      Shiny.setInputValue('js_state_selected', state.selected);
    });
  "))
)

shinyApp(ui, server)
```

### defaultSelected in Shiny

Test that `defaultSelected` highlights the correct rows and that
`getReactableState("selected")` returns the right 1-based indices on load.

Test steps:
1. On load, rows 1, 3, 5 should be checked. Output should show `1, 3, 5`.
2. Navigate to page 2 -- no rows should be checked.
3. Navigate back to page 1 -- rows 1, 3, 5 still checked.
4. Deselect row 3, output should show `1, 5`.

```r
library(shiny)
library(reactable)

ui <- fluidPage(
  h3("defaultSelected test (DuckDB backend)"),
  actionButton("select_rows", "Select rows 2, 4, 6"),
  actionButton("clear_selection", "Clear selection"),
  reactableOutput("table"),
  verbatimTextOutput("selected_output")
)

server <- function(input, output, session) {
  output$table <- renderReactable({
    reactable(
      MASS::Cars93[, 1:8],
      backend = backendDuckDB(),
      selection = "multiple",
      defaultSelected = c(1, 3, 5),
      defaultPageSize = 10,
      searchable = TRUE
    )
  })

  observeEvent(input$select_rows, {
    updateReactable("table", selected = c(2, 4, 6))
  })
  observeEvent(input$clear_selection, {
    updateReactable("table", selected = NA)
  })

  output$selected_output <- renderPrint({
    selected <- getReactableState("table", "selected")
    cat("Count:", length(selected), "\n")
    cat("Selected:", paste(selected, collapse = ", "), "\n")
  })
}

shinyApp(ui, server)
```

### defaultSelected + defaultSorted in Shiny

Test that `defaultSelected` refers to original data row indices, not display
position, even when `defaultSorted` reorders the table.

Test steps:
1. On load, the table is sorted by Horsepower descending. Rows 1 and 2 from the
   original data (Acura Integra and Acura Legend) should be selected, NOT the
   first two rows on screen. Output should show `1, 2`.
2. The selected rows may not be visible on the first page. Navigate to find
   them -- they should be checked.
3. Sort by a different column -- the same rows should stay checked, output
   unchanged.

```r
library(shiny)
library(reactable)

ui <- fluidPage(
  h3("defaultSelected + defaultSorted test (DuckDB backend)"),
  actionButton("select_rows", "Select rows 3, 5, 7"),
  actionButton("clear_selection", "Clear selection"),
  reactableOutput("table"),
  verbatimTextOutput("selected_output")
)

server <- function(input, output, session) {
  output$table <- renderReactable({
    reactable(
      MASS::Cars93[, c("Manufacturer", "Model", "Type", "Min.Price", "Price", "Max.Price", "MPG.city", "Horsepower")],
      backend = backendDuckDB(),
      selection = "multiple",
      defaultSelected = c(1, 2),
      defaultSorted = list(Horsepower = "desc"),
      defaultPageSize = 10,
      searchable = TRUE
    )
  })

  observeEvent(input$select_rows, {
    updateReactable("table", selected = c(3, 5, 7))
  })
  observeEvent(input$clear_selection, {
    updateReactable("table", selected = NA)
  })

  output$selected_output <- renderPrint({
    selected <- getReactableState("table", "selected")
    cat("Count:", length(selected), "\n")
    cat("Selected:", paste(selected, collapse = ", "), "\n")
  })
}

shinyApp(ui, server)
```
