---
title: "DuckDB-WASM Engine — Basic Test"
output:
  html_document:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reactable)
```

## Basic pagination with DuckDB-WASM engine

This test renders a table with the DuckDB-WASM engine, which serializes data as Arrow IPC and
queries it client-side via DuckDB-WASM for pagination.

### MASS::Cars93 (93 rows)

```{r}
reactable(
  MASS::Cars93,
  engine = "duckdb",
  defaultPageSize = 10,
  showPageSizeOptions = TRUE,
  pageSizeOptions = c(10, 25, 50, 100)
)
```

### mtcars (32 rows)

```{r}
reactable(
  mtcars,
  engine = "duckdb",
  defaultPageSize = 10
)
```

### iris (150 rows)

```{r}
reactable(
  iris,
  engine = "duckdb",
  defaultPageSize = 25
)
```

### Small dataset (less than one page)

```{r}
reactable(
  head(mtcars, 5),
  engine = "duckdb",
  defaultPageSize = 10
)
```

### Pre-rendered first page with defaultSorted

```{r}
reactable(
  MASS::Cars93,
  engine = "duckdb",
  defaultPageSize = 10,
  defaultSorted = list(Horsepower = "desc")
)
```

### Large dataset: 100k rows

```{r}
n <- 100000
large_data <- data.frame(
  id = seq_len(n),
  value = rnorm(n),
  category = sample(LETTERS, n, replace = TRUE),
  score = round(runif(n, 0, 100), 2)
)

reactable(
  large_data,
  engine = "duckdb",
  defaultPageSize = 20,
  showPageSizeOptions = TRUE,
  pageSizeOptions = c(20, 50, 100)
)
```

### Large dataset: 1M rows

```{r}
n <- 1000000
big_data <- data.frame(
  id = seq_len(n),
  value = rnorm(n),
  category = sample(LETTERS, n, replace = TRUE),
  score = round(runif(n, 0, 100), 2)
)

reactable(
  big_data,
  engine = "duckdb",
  defaultPageSize = 20,
  showPageSizeOptions = TRUE,
  pageSizeOptions = c(20, 50, 100)
)
```

### Comparison: same data without engine (normal mode)

```{r}
reactable(
  MASS::Cars93,
  defaultPageSize = 10
)
```

## Sorting

### Sortable columns

All columns should be sortable. Click column headers to sort. Sorting should happen server-side
via DuckDB and show correct results across pages.

```{r}
reactable(
  MASS::Cars93,
  engine = "duckdb",
  defaultPageSize = 10,
  sortable = TRUE
)
```

### Default sorted with sortable

Start sorted by Horsepower descending. Clicking other headers should re-sort via DuckDB.

```{r}
reactable(
  MASS::Cars93,
  engine = "duckdb",
  defaultPageSize = 10,
  sortable = TRUE,
  defaultSorted = list(Horsepower = "desc")
)
```

## Column Filters

### Filterable columns

All columns should have filter inputs. Typing in a filter should query DuckDB and update results.
Numeric columns should use "starts with" matching, text columns should use substring matching.

```{r}
reactable(
  MASS::Cars93,
  engine = "duckdb",
  defaultPageSize = 10,
  filterable = TRUE
)
```

### Filterable with sorting

Combining filters and sorting. Filter first, then sort the filtered results.

```{r}
reactable(
  MASS::Cars93,
  engine = "duckdb",
  defaultPageSize = 10,
  filterable = TRUE,
  sortable = TRUE
)
```

## Global Search

### Searchable table

Global search should search across all columns via DuckDB. Page count should update as results
are filtered.

```{r}
reactable(
  MASS::Cars93,
  engine = "duckdb",
  defaultPageSize = 10,
  searchable = TRUE
)
```

### Search with sort and filter combined

All three features working together: search, column filters, and sorting.

```{r}
reactable(
  MASS::Cars93,
  engine = "duckdb",
  defaultPageSize = 10,
  searchable = TRUE,
  filterable = TRUE,
  sortable = TRUE
)
```

## Large Dataset with Sort/Filter/Search

### 100k rows with all features

```{r}
n <- 100000
large_data <- data.frame(
  id = seq_len(n),
  value = rnorm(n),
  category = sample(LETTERS, n, replace = TRUE),
  score = round(runif(n, 0, 100), 2)
)

reactable(
  large_data,
  engine = "duckdb",
  defaultPageSize = 20,
  searchable = TRUE,
  filterable = TRUE,
  sortable = TRUE,
  showPageSizeOptions = TRUE,
  pageSizeOptions = c(20, 50, 100)
)
```

## DuckDB R Server Backend (Shiny only)

The following examples use `server = "duckdb"` which requires a Shiny session. They won't render
in static HTML — run as a Shiny app with `shiny::runApp()` or use the example app at
`inst/examples/shiny-server-data-duckdb.R`.

### Full Shiny app with DuckDB server backend

```{r eval=FALSE}
library(shiny)
library(reactable)

rows <- 100000
dates <- seq.Date(as.Date("2018-01-01"), as.Date("2018-12-01"), "day")
data <- data.frame(
  index = seq_len(rows),
  date = sample(dates, rows, replace = TRUE),
  city = sample(names(precip), rows, replace = TRUE),
  state = sample(rownames(USArrests), rows, replace = TRUE),
  temp = round(runif(rows, 0, 100), 1),
  stringsAsFactors = FALSE
)

ui <- fluidPage(
  titlePanel("DuckDB Server Backend — 100k rows"),
  reactableOutput("tbl")
)

server <- function(input, output) {
  output$tbl <- renderReactable({
    reactable(
      data,
      server = "duckdb",
      defaultPageSize = 20,
      filterable = TRUE,
      searchable = TRUE,
      sortable = TRUE,
      highlight = TRUE,
      showPageSizeOptions = TRUE,
      pageSizeOptions = c(10, 20, 50, 100),
      columns = list(
        index = colDef(name = "ID", width = 80),
        date = colDef(name = "Date"),
        city = colDef(name = "City"),
        state = colDef(
          name = "State",
          html = TRUE,
          cell = JS("function(cell) {
            return '<a href=\"https://wikipedia.org/wiki/' + cell.value + '\">' + cell.value + '</a>'
          }")
        ),
        temp = colDef(
          name = "Temperature",
          format = colFormat(digits = 1)
        )
      )
    )
  })
}

shinyApp(ui, server)
```

### Full Shiny app with 1M rows

```{r eval=FALSE}
library(shiny)
library(reactable)

rows <- 1000000
dates <- seq.Date(as.Date("2018-01-01"), as.Date("2023-12-31"), "day")
data <- data.frame(
  index = seq_len(rows),
  date = sample(dates, rows, replace = TRUE),
  city = sample(names(precip), rows, replace = TRUE),
  state = sample(rownames(USArrests), rows, replace = TRUE),
  temp = round(runif(rows, 0, 100), 1),
  amount = round(runif(rows, 1, 10000), 2),
  stringsAsFactors = FALSE
)

ui <- fluidPage(
  titlePanel("DuckDB Server Backend — 1M rows"),
  reactableOutput("tbl")
)

server <- function(input, output) {
  output$tbl <- renderReactable({
    reactable(
      data,
      server = "duckdb",
      defaultPageSize = 25,
      filterable = TRUE,
      searchable = TRUE,
      sortable = TRUE,
      highlight = TRUE,
      showPageSizeOptions = TRUE,
      pageSizeOptions = c(25, 50, 100, 500),
      columns = list(
        index = colDef(name = "ID", width = 80),
        date = colDef(name = "Date"),
        city = colDef(name = "City"),
        state = colDef(name = "State"),
        temp = colDef(name = "Temperature", format = colFormat(digits = 1)),
        amount = colDef(name = "Amount", format = colFormat(currency = "USD", separators = TRUE))
      )
    )
  })
}

shinyApp(ui, server)
```

### Comparison: DuckDB server vs V8 server vs df server

```{r eval=FALSE}
# All three backends produce the same results for basic sort/filter/search.
# DuckDB is ~100x faster than df for large datasets.

# DuckDB backend (fastest for large data)
reactable(MASS::Cars93, server = "duckdb", filterable = TRUE, searchable = TRUE)

# V8 backend (default, matches client-side behavior exactly)
reactable(MASS::Cars93, server = TRUE, filterable = TRUE, searchable = TRUE)

# Data frame backend (simplest, no external dependencies)
reactable(MASS::Cars93, server = "df", filterable = TRUE, searchable = TRUE)
```
