---
title: "DuckDB-WASM Engine — Basic Test"
output:
  html_document:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reactable)
```

## Basic pagination with DuckDB-WASM engine

This test renders a table with the DuckDB-WASM engine, which serializes data as Arrow IPC and
queries it client-side via DuckDB-WASM for pagination.

### MASS::Cars93 (93 rows)

```{r}
reactable(
  MASS::Cars93,
  engine = "duckdb",
  defaultPageSize = 10,
  showPageSizeOptions = TRUE,
  pageSizeOptions = c(10, 25, 50, 100)
)
```

### mtcars (32 rows)

```{r}
reactable(
  mtcars,
  engine = "duckdb",
  defaultPageSize = 10
)
```

### iris (150 rows)

```{r}
reactable(
  iris,
  engine = "duckdb",
  defaultPageSize = 25
)
```

### Small dataset (less than one page)

```{r}
reactable(
  head(mtcars, 5),
  engine = "duckdb",
  defaultPageSize = 10
)
```

### Pre-rendered first page with defaultSorted

```{r}
reactable(
  MASS::Cars93,
  engine = "duckdb",
  defaultPageSize = 10,
  defaultSorted = list(Horsepower = "desc")
)
```

### Large dataset: 100k rows

```{r}
n <- 100000
large_data <- data.frame(
  id = seq_len(n),
  value = rnorm(n),
  category = sample(LETTERS, n, replace = TRUE),
  score = round(runif(n, 0, 100), 2)
)

reactable(
  large_data,
  engine = "duckdb",
  defaultPageSize = 20,
  showPageSizeOptions = TRUE,
  pageSizeOptions = c(20, 50, 100)
)
```

### Large dataset: 1M rows

```{r}
n <- 1000000
big_data <- data.frame(
  id = seq_len(n),
  value = rnorm(n),
  category = sample(LETTERS, n, replace = TRUE),
  score = round(runif(n, 0, 100), 2)
)

reactable(
  big_data,
  engine = "duckdb",
  defaultPageSize = 20,
  showPageSizeOptions = TRUE,
  pageSizeOptions = c(20, 50, 100)
)
```

### Comparison: same data without engine (normal mode)

```{r}
reactable(
  MASS::Cars93,
  defaultPageSize = 10
)
```
