---
title: "Demo Cookbook"
output:
  html_document:
    toc: true
    toc_float:
      smooth_scroll: false
resource_files:
  - images
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reactable)
```

> A collection of recipes used to create the reactable demos

## Insert links

```{r}
data <- data.frame(
  Address = c("https://google.com", "https://yahoo.com", "https://duckduckgo.com"),
  Site = c("Google", "Yahoo", "DuckDuckGo")
)

reactable(data, columns = list(
  # Using htmltools to render a link
  Address = colDef(cell = function(value) {
    htmltools::tags$a(href = value, target = "_blank", value)
  }),
  # Or using raw HTML
  Site = colDef(html = TRUE, cell = function(value, index) {
    sprintf('<a href="%s" target="_blank">%s</a>', data$Address[index], value)
  })
))
```

## Conditional formatting

### Color scales

To add color scales, you can use R's built-in
[color utilities](https://bookdown.org/rdpeng/exdata/plotting-and-color-in-r.html#color-utilities-in-r)
(or other color manipulation package):

```{r}
data <- iris[10:29, ]
orange_pal <- function(x) rgb(colorRamp(c("#ffe4cc", "#ff9500"))(x), maxColorValue = 255)

reactable(data, columns = list(
  Petal.Length = colDef(style = function(value) {
    normalized <- (value - min(data$Petal.Length)) / (max(data$Petal.Length) - min(data$Petal.Length))
    color <- orange_pal(normalized)
    list(background = color)
  })
))
```

```{r}
dimnames <- list(start(nottem)[1]:end(nottem)[1], month.abb)
temps <- matrix(nottem, ncol = 12, byrow = TRUE, dimnames = dimnames)

# Excel-inspired 3-color scale
GnYlRd <- function(x) rgb(colorRamp(c("#63be7b", "#ffeb84", "#f8696b"))(x), maxColorValue = 255)

reactable(
  temps,
  defaultColDef = colDef(
    style = function(value) {
      if (!is.numeric(value)) return()
      normalized <- (value - min(nottem)) / (max(nottem) - min(nottem))
      color <- GnYlRd(normalized)
      list(background = color)
    },
    format = colFormat(digits = 1),
    minWidth = 50
  ),
  columns = list(
    .rownames = colDef(name = "Year", sortable = TRUE, align = "left")
  ),
  bordered = TRUE
)
```

### Formatting changes

```{r}
stocks <- data.frame(
  Symbol = c("GOOG", "FB", "AMZN", "NFLX", "TSLA"),
  Price = c(1265.13, 187.89, 1761.33, 276.82, 328.13),
  Change = c(4.14, 1.51, -19.45, 5.32, -12.45)
)

reactable(stocks, columns = list(
  Change = colDef(
    cell = function(value) {
      if (value >= 0) paste0("+", value) else value
    },
    style = function(value) {
      color <- if (value > 0) {
        "#008000"
      } else if (value < 0) {
        "#e00000"
      }
      list(fontWeight = 600, color = color)
    }
  )
))
```

### Tags and badges

```{r, echo=FALSE}
set.seed(20)
```

```{r tags, eval=FALSE}
orders <- data.frame(
  Order = 2300:2304,
  Created = seq(as.Date("2019-04-01"), by = "day", length.out = 5),
  Customer = sample(rownames(MASS::painters), 5),
  Status = sample(c("Pending", "Paid", "Canceled"), 5, replace = TRUE)
)

reactable(orders, columns = list(
  Status = colDef(cell = function(value) {
    class <- paste0("tag status-", tolower(value))
    htmltools::div(class = class, value)
  })
))
```

```{css}
.tag {
  display: inline-block;
  padding: 2px 12px;
  border-radius: 15px;
  font-weight: 600;
  font-size: 12px;
}

.status-paid {
  background: hsl(116, 60%, 90%);
  color: hsl(116, 30%, 25%);
}

.status-pending {
  background: hsl(230, 70%, 90%);
  color: hsl(230, 45%, 30%);
}

.status-canceled {
  background: hsl(350, 70%, 90%);
  color: hsl(350, 45%, 30%);
}
```

```{r ref.label="tags", echo=FALSE}
```

## Bar charts

There are different ways to create bar charts, but here's one way
[using HTML and CSS](https://css-tricks.com/making-charts-with-css/):

```{r}
library(htmltools)

data <- MASS::Cars93[20:49, c("Make", "MPG.city", "MPG.highway")]

# Render a bar chart with a label on the left
bar_chart <- function(label, width = "100%", height = "16px", fill = "#00bfc4", background = NULL) {
  bar <- div(style = list(background = fill, width = width, height = height))
  chart <- div(style = list(flexGrow = 1, marginLeft = "8px", background = background), bar)
  div(style = list(display = "flex", alignItems = "center"), label, chart)
}

reactable(data, columns = list(
  MPG.city = colDef(name = "MPG (city)", align = "left", cell = function(value) {
    width <- paste0(value / max(data$MPG.city) * 100, "%")
    bar_chart(value, width = width)
  }),
  MPG.highway = colDef(name = "MPG (highway)", align = "left", cell = function(value) {
    width <- paste0(value / max(data$MPG.highway) * 100, "%")
    bar_chart(value, width = width, fill = "#fc5185", background = "#e1e1e1")
  })
))
```

### Positive and negative values

```{r}
library(htmltools)

# Render a bar chart with positive and negative values
bar_chart_pos_neg <- function(label, value, max_value = 1, height = "16px",
                              pos_fill = "#02aa5c", neg_fill = "#ff121a") {
  neg_chart <- div(style = list(flex = "1 1 0"))
  pos_chart <- div(style = list(flex = "1 1 0"))
  width <- paste0(abs(value / max_value) * 100, "%")

  if (value < 0) {
    bar <- div(style = list(marginLeft = "8px", background = neg_fill, width = width, height = height))
    chart <- div(style = list(display = "flex", alignItems = "center", justifyContent = "flex-end"), label, bar)
    neg_chart <- tagAppendChild(neg_chart, chart)
  } else {
    bar <- div(style = list(marginRight = "8px", background = pos_fill, width = width, height = height))
    chart <- div(style = list(display = "flex", alignItems = "center"), bar, label)
    pos_chart <- tagAppendChild(pos_chart, chart)
  }

  div(style = list(display = "flex"), neg_chart, pos_chart)
}

data <- data.frame(
  company = sprintf("Company%02d", 1:10),
  profit_chg = c(0.2, 0.685, 0.917, 0.284, 0.105, -0.701, -0.528, -0.808, -0.957, -0.11)
)

reactable(data, bordered = TRUE, columns = list(
  company = colDef(name = "Company", minWidth = 100),
  profit_chg = colDef(
    name = "Change in Profit",
    defaultSortOrder = "desc",
    cell = function(value) {
      label <- paste0(round(value * 100), "%")
      bar_chart_pos_neg(label, value)
    },
    align = "center",
    minWidth = 400
  )
))
```

## Embed images

To embed an image, render an [`<img>`](https://developer.mozilla.org/en-US/docs/Web/HTML/Element/img)
element into the table. Make sure to add [`alt` text](https://webaim.org/techniques/alttext/)
for accessibility, if the image is not purely decorative.

```{r}
library(htmltools)

data <- data.frame(
  Animal = c("beaver", "cow", "wolf", "goat"),
  Body = c(1.35, 465, 36.33, 27.66),
  Brain = c(8.1, 423, 119.5, 115)
)

reactable(data, columns = list(
  Animal = colDef(cell = function(value) {
    image <- img(src = sprintf("images/%s.png", value), height = "24px", alt = "")
    tagList(
      div(style = list(display = "inline-block", width = "45px"), image),
      value
    )
  }),
  Body = colDef(name = "Body (kg)"),
  Brain = colDef(name = "Brain (g)")
))
```

## Rating stars

This example uses Font Awesome icons
([via Shiny](https://shiny.rstudio.com/reference/shiny/latest/icon.html))
to render rating stars in a table:

```{r}
library(htmltools)

rating_stars <- function(rating, max_rating = 5) {
  star_icon <- function(empty = FALSE) {
    tagAppendAttributes(shiny::icon("star"),
      style = paste("color:", if (empty) "#edf0f2" else "orange"),
      "aria-hidden" = "true"
    )
  }
  rounded_rating <- floor(rating + 0.5)  # always round up
  stars <- lapply(seq_len(max_rating), function(i) {
    if (i <= rounded_rating) star_icon() else star_icon(empty = TRUE)
  })
  label <- sprintf("%s out of %s", rating, max_rating)
  div(title = label, "aria-label" = label, role = "img", stars)
}

ratings <- data.frame(
  Movie = c("Silent Serpent", "Nowhere to Hyde", "The Ape-Man Goes to Mars", "A Menace in Venice"),
  Rating = c(3.65, 2.35, 4.5, 1.4),
  Votes = c(115, 37, 60, 99)
)

reactable(ratings, columns = list(
  Rating = colDef(cell = function(value) rating_stars(value))
))
```

## Total rows

```{r, message=FALSE}
library(dplyr)
library(htmltools)

data <- MASS::Cars93[18:47, ] %>%
  select(Manufacturer, Model, Type, Sales = Price)

reactable(
  data,
  defaultPageSize = 5,
  columns = list(
    Manufacturer = colDef(footer = "Total"),
    Sales = colDef(footer = sprintf("$%.2f", sum(data$Sales)))
  ),
  defaultColDef = colDef(footerStyle = list(fontWeight = "bold"))
)
```

### Dynamic totals

To update the total when filtering the table, calculate the total in a
JavaScript render function:

```{r}
reactable(
  data,
  searchable = TRUE,
  defaultPageSize = 5,
  minRows = 5,
  columns = list(
    Manufacturer = colDef(footer = "Total"),
    Sales = colDef(
      footer = JS("function(colInfo) {
        var values = colInfo.data.map(function(row) { return row[colInfo.column.id] })
        var total = values.reduce(function(a, b) { return a + b }, 0)
        return '$' + total.toFixed(2)
      }")
    )
  ),
  defaultColDef = colDef(footerStyle = list(fontWeight = "bold"))
)
```

## Nested tables

To create nested tables, use `reactable()` in a row details renderer:

```{r, message=FALSE}
library(dplyr)

data <- MASS::Cars93[18:47, ] %>%
  mutate(ID = as.character(18:47), Date = seq(as.Date("2019-01-01"), by = "day", length.out = 30)) %>%
  select(ID, Date, Manufacturer, Model, Type, Price)

sales_by_mfr <- group_by(data, Manufacturer) %>%
  summarize(Quantity = n(), Sales = sum(Price))

reactable(
  sales_by_mfr,
  details = function(index) {
    sales <- filter(data, Manufacturer == sales_by_mfr$Manufacturer[index]) %>% select(-Manufacturer)
    tbl <- reactable(sales, outlined = TRUE, highlight = TRUE, fullWidth = FALSE)
    htmltools::div(style = list(margin = "12px 45px"), tbl)
  },
  onClick = "expand",
  rowStyle = list(cursor = "pointer")
)
```

## Units on first row only

To show a label on the first row only (even when sorting), use a JavaScript
render function to add the label when the cell's
[`viewIndex`](../custom-rendering.html#cellinfo-properties) is 0.

If the label breaks the alignment of values in the column, realign the values
by adding white space to the cells without units. Two ways to do this are shown below.

```{r units_first_row, eval=FALSE}
data <- MASS::Cars93[40:44, c("Make", "Length", "Luggage.room")]

reactable(data, class = "car-specs", columns = list(
  # Align values using white space (and a monospaced font)
  Length = colDef(
    cell = JS("function(cellInfo) {
      var units = cellInfo.viewIndex === 0 ? '\u2033' : ' '
      return cellInfo.value + units
    }"),
    class = "number"
  ),
  # Align values using a fixed-width container for units
  Luggage.room = colDef(
    name = "Luggage Room",
    cell = JS("function(cellInfo) {
      var units = cellInfo.viewIndex === 0 ? ' ft³' : ''
      return cellInfo.value + '<div class=\"units\">' + units + '</div>'
    }"),
    html = TRUE
  )
))
```

```{css}
.car-specs .number {
  font-family: "Courier New", Courier, monospace;
  white-space: pre;
}

.car-specs .units {
  display: inline-block;
  width: 18px;
}
```

```{r ref.label="units_first_row", echo=FALSE}
```

## Tooltips

To add tooltips to column headers, render the headers with a
[`title`](https://developer.mozilla.org/en-US/docs/Web/HTML/Global_attributes/title)
attribute:

```{r}
data <- mtcars[1:5, c("mpg", "cyl", "disp", "hp")]

with_tooltip <- function(value, tooltip) {
  span(style = "text-decoration: underline; text-decoration-style: dotted;", title = tooltip, value)
}

reactable(
  data,
  columns = list(
    mpg = colDef(header = with_tooltip("mpg", "Miles/(US) gallon")),
    cyl = colDef(header = with_tooltip("cyl", "Number of cylinders")),
    disp = colDef(header = with_tooltip("disp", "Displacement (cu.in.)")),
    hp = colDef(header = with_tooltip("hp", "Gross horsepower"))
  )
)
```

## Highlight cells

```{r}
data <- MASS::road[11:17, ]

reactable(data, defaultColDef = colDef(
  style = function(value, index, name) {
    if (is.numeric(value) && value == max(data[[name]])) {
      list(fontWeight = "bold")
    }
  })
)
```

## Highlight columns

```{r}
reactable(iris[1:5, ], columns = list(
  Petal.Length = colDef(style = list(background = "rgba(0, 0, 0, 0.03)"))
))
```

## Highlight rows

```{r}
reactable(iris[1:5, ], rowStyle = function(index) {
  if (index == 2) list(fontWeight = "bold")
  else if (iris[index, "Petal.Length"] >= 1.5) list(background = "rgba(0, 0, 0, 0.05)")
})
```

## Highlight sorted headers

To style headers on hover, select headers with a `:hover` pseudo-class in CSS:

```{r sort_highlight, eval=FALSE}
reactable(iris[1:5, ], defaultColDef = colDef(headerClass = "sort-header"))
```

```{css}
.sort-header:hover {
  background: rgba(0, 0, 0, 0.03);
}
```

To style sorted headers, select headers with either an `aria-sort="ascending"`
or `aria-sort="descending"` attribute:

```{css}
.sort-header[aria-sort="ascending"],
.sort-header[aria-sort="descending"] {
  background: rgba(0, 0, 0, 0.03);
}
```

```{r ref.label="sort_highlight", echo=FALSE}
```

## Highlight sorted columns

To style sorted columns, use a JavaScript function to style columns based on the
table's sorted state:

```{r}
reactable(
  iris[1:5, ],
  defaultSorted = "Sepal.Width",
  defaultColDef = colDef(
    style = JS("function(rowInfo, colInfo, state) {
      // Highlight sorted columns
      for (var i = 0; i < state.sorted.length; i++) {
        if (state.sorted[i].id === colInfo.id) {
          return { background: 'rgba(0, 0, 0, 0.03)' }
        }
      }
    }")
  )
)
```

## Borders between groups of data

To add borders between groups, use an R or JavaScript function to style rows
based on the previous or next row's data. If the table can be sorted, use a
JavaScript function to style rows only when the groups are sorted.

```{r, message=FALSE}
library(dplyr)

data <- as_tibble(MASS::painters, rownames = "Painter") %>%
  filter(School %in% c("A", "B", "C")) %>%
  mutate(School = recode(School, A = "Renaissance", B = "Mannerist", C = "Seicento")) %>%
  select(Painter, School, everything()) %>%
  group_by(School) %>%
  slice(1:3)

reactable(
  data,
  defaultSorted = list(School = "asc", Drawing = "desc"),
  borderless = TRUE,
  rowStyle = JS("
    function(rowInfo, state) {
      // Add horizontal separators between groups when sorting by school
      var firstSorted = state.sorted[0]
      if (firstSorted && firstSorted.id === 'School') {
        var nextRow = state.pageRows[rowInfo.viewIndex + 1]
        if (nextRow && rowInfo.row['School'] !== nextRow['School']) {
          // Use box-shadow to add a 2px border without taking extra space
          return { boxShadow: 'inset 0 -2px 0 rgba(0, 0, 0, 0.1)' }
        }
      }
    }
  ")
)
```

## Merge cells

You can give the appearance of merged cells by hiding cells based on the
previous row's data. Just like with the example above, you'll need a JavaScript
style function for grouping to work with sorting, filtering, and pagination.

```{r, message=FALSE}
library(dplyr)

data <- as_tibble(MASS::painters, rownames = "Painter") %>%
  filter(School %in% c("A", "B", "C")) %>%
  mutate(School = recode(School, A = "Renaissance", B = "Mannerist", C = "Seicento")) %>%
  select(School, Painter, everything()) %>%
  group_by(School) %>%
  slice(1:3)

reactable(
  data,
  columns = list(
    School = colDef(
      style = JS("function(rowInfo, colInfo, state) {
        var firstSorted = state.sorted[0]
        // Merge cells if unsorted or sorting by school
        if (!firstSorted || firstSorted.id === 'School') {
          var prevRow = state.pageRows[rowInfo.viewIndex - 1]
          if (prevRow && rowInfo.row['School'] === prevRow['School']) {
            return { visibility: 'hidden' }
          }
        }
      }")
    )
  ),
  outlined = TRUE
)
```

## Borders between columns

```{r}
reactable(
  iris[1:5, ],
  columns = list(
    Sepal.Width = colDef(style = list(borderRight = "1px solid rgba(0, 0, 0, 0.1)")),
    Petal.Width = colDef(style = list(borderRight = "1px solid rgba(0, 0, 0, 0.1)"))
  ),
  borderless = TRUE
)
```

## Custom fonts

Tables don't have a default font, and just inherit the font properties
from their parent elements. (This may explain why tables look different in
in R Markdown documents or Shiny apps vs. standalone pages).

To customize the table font, you can set a font on the page, or on the table itself:

```{r}
reactable(
  iris[1:5, ],
  style = list(fontFamily = "Work Sans, sans-serif", fontSize = "14px")
)
```

```{r}
# Add a custom font from Google Fonts
htmltools::tags$link(href = "https://fonts.googleapis.com/css?family=Work+Sans:400,600,700&display=fallback",
                     rel = "stylesheet")
```

> Side note: these docs use a [system font stack](https://css-tricks.com/snippets/css/system-font-stack):
>
> `font-family: -apple-system, BlinkMacSystemFont, "Segoe UI", Helvetica, Arial, sans-serif;`

## Custom sort indicators

To use a custom sort indicator, hide the sort icon with `showSortIcon = FALSE` and
add your own indicator.

For example, changing the sort indicator to a bar using CSS:

```{r bar_sort, eval=FALSE}
reactable(
  iris[1:5, ],
  showSortIcon = FALSE,
  bordered = TRUE,
  defaultSorted = "Petal.Length",
  defaultColDef = colDef(headerClass = "bar-sort-header")
)
```

```{css}
.bar-sort-header[aria-sort="ascending"] {
  box-shadow: inset 0 3px 0 0 rgba(0, 0, 0, 0.6) !important;
}

.bar-sort-header[aria-sort="descending"] {
  box-shadow: inset 0 -3px 0 0 rgba(0, 0, 0, 0.6) !important;
}
```

```{r ref.label="bar_sort", echo=FALSE}
```

```{css echo=FALSE}
/* rmarkdown html documents */
.main-container {
  font-family: -apple-system, BlinkMacSystemFont, "Segoe UI", Helvetica, Arial, sans-serif;
}

.main-container blockquote {
  font-size: inherit;
}

/* pkgdown articles */
.contents {
  font-family: -apple-system, BlinkMacSystemFont, "Segoe UI", Helvetica, Arial, sans-serif;
}

.contents blockquote {
  font-size: inherit;
}

.reactable {
  font-size: 14px;
}
```
