---
title: "Cookbook"
output:
  html_document:
    toc: true
    toc_float:
      smooth_scroll: false
resource_files:
  - images
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reactable)
```

## Insert links

```{r}
data <- data.frame(
  Address = c("https://google.com", "https://yahoo.com", "https://duckduckgo.com"),
  Site = c("Google", "Yahoo", "DuckDuckGo")
)

reactable(data, columns = list(
  # Using htmltools to render a link
  Address = colDef(cell = function(value) {
    htmltools::tags$a(href = value, value)
  }),
  # Or using raw HTML
  Site = colDef(html = TRUE, cell = function(value, index) {
    sprintf('<a href="%s" target="_blank">%s</a>', data$Address[index], value)
  })
))
```

## Conditional coloring

```{r}
data <- iris[1:20, ]
orange_pal <- function(x) rgb(colorRamp(c("#ffe4cc", "#ff9500"))(x), maxColorValue = 255)

reactable(data, highlight = FALSE, columns = list(
  Petal.Width = colDef(style = function(value) {
    color <- orange_pal(value / max(data$Petal.Width))
    list(background = color)
  })
))
```

## Bar charts

```{r}
library(htmltools)

data <- MASS::Cars93[, c("Manufacturer", "Model", "MPG.city")]

# Render a bar chart with a label on the left
bar_chart <- function(label, width = "100%", background = "#00bfc4", height = "16px") {
  bar <- div(style = list(background = background, width = width, height = height))
  chart <- div(style = list(flexGrow = 1, paddingLeft = "8px"), bar)
  div(style = list(display = "flex", alignItems = "center"), label, chart)
}

reactable(data, highlight = FALSE, columns = list(
  MPG.city = colDef(name = "MPG (city)", align = "left", cell = function(value) {
    width <- paste0(value / max(data$MPG.city) * 100, "%")
    bar_chart(value, width)
  })
))
```

## Embed images

```{r}
library(htmltools)

data <- data.frame(
  Animal = c("beaver", "cow", "wolf", "goat"),
  Body = c(1.35, 465, 36.33, 27.66),
  Brain = c(8.1, 423, 119.5, 115)
)

reactable(data, columns = list(
  Animal = colDef(cell = function(value) {
    image <- img(src = sprintf("images/%s.png", value), height = "24px", alt = "")
    tagList(
      div(style = list(display = "inline-block", width = "45px"), image),
      value
    )
  }),
  Body = colDef(name = "Body (kg)"),
  Brain = colDef(name = "Brain (g)")
))
```

## Total rows

```{r, message=FALSE}
library(dplyr)
library(htmltools)

data <- MASS::Cars93[18:47, ] %>%
  select(Manufacturer, Model, Type, Sales = Price)

reactable(
  data,
  defaultPageSize = 5,
  columns = list(
    Manufacturer = colDef(footer = "Total"),
    Sales = colDef(footer = sprintf("$%.2f", sum(data$Sales)))
  ),
  defaultColDef = colDef(footerStyle = list(fontWeight = "bold"))
)
```

### Dynamic totals

To update the total when filtering the table, calculate the total in a
JS renderer instead:

```{r}
reactable(
  data,
  searchable = TRUE,
  defaultPageSize = 5,
  minRows = 5,
  columns = list(
    Manufacturer = colDef(footer = "Total"),
    Sales = colDef(
      footer = JS("function(colInfo) {
        var values = colInfo.data.map(function(row) { return row[colInfo.column.id] })
        var total = values.reduce(function(a, b) { return a + b }, 0)
        return '$' + total.toFixed(2)
      }")
    )
  ),
  defaultColDef = colDef(footerStyle = list(fontWeight = "bold"))
)
```

## Highlight columns

```{r}
reactable(iris[1:5, ], highlight = FALSE, columns = list(
  Petal.Width = colDef(style = list(background = "rgba(0, 0, 0, 0.03)"))
))
```

## Highlight rows

```{r}
reactable(iris[1:5, ], highlight = FALSE, rowStyle = function(index) {
  if (index == 2) list(fontWeight = "bold")
  else if (index == 4) list(background = "rgba(0, 0, 0, 0.05)")
})
```

## Highlight sorted headers

To style headers on hover, select headers with a `:hover` pseudo-class in CSS:

```{r sort_highlight, eval=FALSE}
reactable(iris[1:5, ], defaultColDef = colDef(headerClass = "sort-header"))
```

```{css}
.sort-header:hover {
  background: rgba(0, 0, 0, 0.03);
}
```

To style sorted headers, select headers with either an `aria-sort="ascending"`
or `aria-sort="descending"` attribute:

```{css}
.sort-header[aria-sort="ascending"],
.sort-header[aria-sort="descending"] {
  background: rgba(0, 0, 0, 0.03);
}
```

```{r ref.label="sort_highlight", echo=FALSE}
```

## Highlight sorted columns

Use a JS callback to style columns based on the table's sorted state:

```{r}
reactable(
  iris[1:5, ],
  defaultSorted = "Sepal.Width",
  highlight = FALSE,
  defaultColDef = colDef(
    style = JS("function(rowInfo, colInfo, state) {
      // Highlight sorted columns
      for (var i = 0; i < state.sorted.length; i++) {
        if (state.sorted[i].id === colInfo.id) {
          return { background: 'rgba(0, 0, 0, 0.03)' }
        }
      }
    }")
  )
)
```

## Custom sort indicators

To use a custom sort indicator, hide the sort icon with `showSortIcon = FALSE` and
add your own indicator.

For example, changing the sort indicator to a bar using CSS:

```{r bar_sort, eval=FALSE}
reactable(
  iris[1:5, ],
  showSortIcon = FALSE,
  bordered = TRUE,
  defaultSorted = "Petal.Length",
  defaultColDef = colDef(headerClass = "bar-sort-header")
)
```

```{css}
.bar-sort-header[aria-sort="ascending"] {
  box-shadow: inset 0 3px 0 0 rgba(0, 0, 0, 0.6) !important;
}

.bar-sort-header[aria-sort="descending"] {
  box-shadow: inset 0 -3px 0 0 rgba(0,0,0,0.6) !important;
}
```

```{r ref.label="bar_sort", echo=FALSE}
```

```{css echo=FALSE}
/* rmarkdown html documents */
.main-container {
  font-family: -apple-system, BlinkMacSystemFont, "Segoe UI", Helvetica, Arial, sans-serif;
}

.main-container blockquote {
  font-size: inherit;
}

/* pkgdown articles */
.contents {
  font-family: -apple-system, BlinkMacSystemFont, "Segoe UI", Helvetica, Arial, sans-serif;
}

.contents blockquote {
  font-size: inherit;
}

.reactable {
  font-size: 14px;
}
```
